### 2019-07-02 paper.tex
Initial adaptation from SAP template

\documentclass{llncs}

\usepackage{booktabs}
\usepackage[ruled]{algorithm2e} % For algorithms

\usepackage{siunitx}
\usepackage{graphicx}
\usepackage{multirow}
%\usepackage{subcaption}
\usepackage[caption=false]{subfig}
\usepackage[capitalise]{cleveref}
\usepackage{todonotes}

\begin{document}

\title{Audio Interface of an Active Perception Navigation Aid for People with Visual Impairments}
\titlerunning{Active Perception Interface for People with Visual Impairments}

\author{Jacobus C. Lock\inst{1} \and 
Iain Gilchrist\inst{2} \and
Grzegorz Cielniak\inst{1} \and
Nicola Bellotto\inst{1}}

\institute{University of Lincoln \email{\{jlock, gcielniak, nbellotto\}@lincoln.ac.uk} \url{https://lcas.github.io/ActiVis/} \and
	   University of Bristol \email{i.d.gilchrist@bristol.ac.uk}}

\authorrunning{Jacobus C. Lock et.\ al.}

\maketitle

\begin{abstract}
	Our aim is to build a portable active navigation system for people with visual impairments that uses active perception techniques and a combination of feedback modes to identify a scene and guide a user to their destination.
	Such a system requires a non-visual feedback interface and in this paper we investigate the effectiveness of a spatial audio tone with a varying pitch component, played through bone-conducting headphones, in conveying the pan and tilt angles of a target to the user in a pointing task.
	The secondary goal is to determine how variations to the pitch's rate of change affect the user's performance.
	For this, we conducted a set of experiments with both blindfolded participants and participants with visual impairments and found that, with some limitations, a spatialised audio tone with a varying pitch component can successfully convey a target's pan and tilt angles. 
\keywords{Human-machine interface \and vision impairment \and spatialised sound \and varying pitch, pointing task}
\end{abstract}

\section{Introduction}

In recent years, governments have spearheaded numerous initiatives to support people with disabilities and enable them to play a more active role in modern society.
In the UK, the Royal National Institute of Blind People (RNIB) has prioritised enabling people with visual impairments to use some of the services and products many people take for granted, such as public transport and mobile apps~\cite{rnib-objectives}.
Improvements in modern computing have made it possible for new and innovative solutions to these problems to come to the fore.
In particular, researchers in the active vision field have have made much progress in enabling machines to autonomously manipulate cameras to gather information about an environment for mapping and object and face tracking tasks~\cite{bajcsy2018revisiting}.
There is, however, a significant research question about whether techniques from active machine vision can be applied to humans, i.e.\ can a machine identify a point of interest in a scene and direct a human, instead of an electronic servo, to focus on that point?
If this can be done, it would be beneficial to people with visual impairments and will augment their ability to search for an arbitrary point or object of interest and identify an unknown scene.

To this end, we are developing a mobile navigation system using a Google Project Tango tablet device\footnote{\url{https://en.wikipedia.org/wiki/Tango\_(platform)}}, pictured in \cref{fig:tango-headphone}, that caters to the perception limitations of people with vision impairments by using a non-visual interface.
This system will ultimately allow users with visual impairments to reach a target location in the final leg of their journey, i.e.\ the so-called `last 10 yards' problem~\cite{google2016blind,bellotto2013}. 
The ideas here can also be extended and applied to mainstream navigation tools, such as Google Maps.
A Tango-enabled device comes pre-equipped with powerful image-processing, localisation and depth-perception capabilities and is built on top of a standard Android platform, providing access to the entire set of input/output options that this operating system has to offer.

\begin{figure}
  \centering
  \subfloat[]{\label{fig:tango-headphone}\includegraphics[width=0.5\columnwidth]{figures/tango_headphone.png}}
  ~
  \subfloat[]{\label{fig:participant}\includegraphics[width=0.35\columnwidth]{figures/vi_participant.pdf}}
  \caption{Pictures of the Tango device and bone-conducting headphones the system is based on (left) and both in use in an experiment (right).}
\end{figure}

For this work, we implement and evaluate an audio interface that will be integrated into the guidance system to provide navigation instructions to the user. 
To evaluate this guidance mode, we conducted a set of experiments involving a target-search task where the participants were guided with the audio instructions to point a camera at a set of virtual targets.
The system uses a set of bone-conducting headphones to provide the audio instructions, with the latter being used to avoid interfering with a user's normal hearing function; something people with limited vision typically rely upon. 
However, this type of headphones bypass the external structure of the ear which provide the ability to localise a sound source's elevation.
We therefore use a tone of varying pitch to convey the target's tilt position w.r.t.\ the camera's pointing vector, $C$, pictured in \cref{fig:cam-coords}.
The tone is spatialised to provide the target's position in the pan dimension.
A similar interface can be found in~\cite{durette2008visuo}, where they used a varying pitch component without spatialising the tone. 
However, they did not investigate this interface's target-finding accuracy.
Furthermore, we wish to determine how modifying the pitch's rate of change will affect target-finding performance. 
Therefore, this work contributes to the body of knowledge by determining how accurately a spatialised tone with varying pitch can direct a user towards a target, and how the pitch's rate of change affects this accuracy.

\begin{figure}
  \centering
  \includegraphics[width=0.4\columnwidth]{figures/camera_coordinate.pdf}
  \caption{A diagram showing the device's frame of reference and the target's pan and tilt angles.}\label{fig:cam-coords}
\end{figure}

The remainder of this paper is organised as follows: \cref{sec:lit-review} gives an overview of existing navigation systems and audio interfaces for people with visual impairments.
\cref{sec:portable-navigation} discusses the navigation device's hardware and interface setup and describes how the target's position is conveyed to the user.
The experiment we conducted are explained in \cref{sec:experiments} and the results are presented and discussed in \cref{sec:results}. 
Finally, \cref{sec:conclusion} concludes this paper with a summary of the findings and comments on future work.

\section{Previous Work}\label{sec:lit-review}

Earlier attempts at improving a person with visual impairment's navigation experience involved outfitting the commonly-used walking cane with remote detection systems, such as sonar, radar and RFID tags, to transform a cane into an early-warning system~\cite{ulrich1997,marion2008batcane,faria2010electronic,willis2005}.
Bluetooth tags and a smartphone can be used in a similar manner~\cite{sato2017navcog3}.
However, the expense and effort of maintaining such tags make them less attractive for wide-scale adoption.
Computer vision-based systems provide a good compromise between usability, cost and accuracy.
One solution is to use RGB-D depth sensing cameras, which are becoming increasingly accurate and cost-effective, to build a 3D image map of the environment, and guiding a user through it~\cite{lee2015,rodriguez2012obstacle}.
Alternatively, object recognition techniques can be used to detect various objects and landmarks, such as doors, staircases, etc., and communicate their relative location to the user~\cite{schauerte2012assistive,tian2013b,fiannaca2014headlock}, or use audio instructions to guide a user to fully capture an object of interest~\cite{vazquez2012helping}.

An important feature of user-centric systems is a human-machine interface (HMI) that enables effective two-way communication between the system and the user.
Surveys found that people with visual impairments prefer receiving instructions in the form of speech and haptic feedback cues~\cite{khoo2016multimodal,ross2000wearable,vazquez2012helping}.
However, haptic feedback modes typically have a lower bandwidth when compared to audio feedback and also require the user to wear a special device in order to transmit the signals effectively.
Work has also been done in translating a visual scene into non-visual formats, with so-called sensory substitution systems (e.g. `The Voice'~\cite{meijer2010}) and virtual audio reality (VAR) systems~\cite{frauenberger2003} reporting favourable results.
However, The Voice has a very steep learning curve that has proven to be a significant barrier to entry and with VAR systems, it is not clear how unknown environments, where markers have not yet been encoded, would be handled and described to the user.

Previous experiments have determined that people are able to find the location of a stationary sound source with an error of $\pm$\SI{35}{\degree}, in both the pan and tilt dimensions~\cite{zwiers2001spatial}, for both early-blind and normally-sighted people.
However, in~\cite{lewald2013exceptional,lessard1998early} it was found that the blind have a clear advantage in localisation accuracy over sighted people when presented with more complex tasks, such as targets in motion and narrow-band stimuli, with~\cite{lewald2013exceptional} reporting an average absolute localisation error of around \SI{10}{\degree} in the pan and tilt dimensions.
Researchers have used simulated spatialised audio to inform the user of the sound source's direction~\cite{holland2002audiogps,kammoun2012navigation,rebillat2009smart,menelas2010audio,wilson2007swan,zotkin2004rendering}.
In these works, a sound was played through a set of headphones and the source spatialised with a head-related transfer function (HRTF), tricking the listener into thinking the sound source was located at some arbitrary 3D location.
There are also experimental results about how well users can find targets presented with spatial sound in the tilt and panning dimensions with normal headphones~\cite{katz2011spatial,zwiers2001spatial}, as well as work that has shown similar levels of localisation performance between normal and bone-conducting headphones in the pan dimension~\cite{macdonald2006spatial}.
However, to our knowledge no extensive work or experiments have been done to determine how well users respond to tilt adjustment instructions using a tone with \textit{varying pitch}, in particular when using bone-conducting headphones, and how changes to this tone's behaviour affects performance.

\section{Portable Navigation System}\label{sec:portable-navigation}

\subsection{System Setup}

Our ultimate goal is to deliver a navigation system that can guide a user with visual impairments on the last leg of their journey (e.g.\ to a specific aisle and shelf in a shop) using only a mobile phone.
A large amount of data need to be translated from a visual form into a format that is useful to people with limited or no vision.
We therefore plan to use a combination of voice, audio and vibration cues to translate visual navigation data as effectively as possible and overcome the bandwidth limitation of the human ear. 
However, for this paper, we only considered the spatialised audio mode and its variation in pitch in order to determine its effectiveness in conveying a target's pan and tilt angles.

The system is based on a concept proposed in~\cite{bellotto2013,lock2017portable}, which uses a Google Tango device.
The navigation system is based on a Google Tango tablet, an Android-based device that comes equipped with an RGB-D camera to sense colour and depth and combines an inertial measurement unit (IMU) with powerful and robust landmark recognition and image processing algorithms to localise itself in real-time.
An added benefit of this platform is its familiar, compact form-factor which will help overcome the hurdle of user-acceptance and usability.
The guidance instructions are presented through a set of bone-conducting headphones (see \cref{fig:tango-headphone}) that are placed externally on a user's head so that the system does not occlude the perception of real-world sounds and does not interfere with a user's normal hearing function.

A diagram of the experimental system pipeline is shown in \cref{fig:pipeline}, where the arrows indicate the direction of the flow of information.
When the user taps the Tango's screen, a new virtual target is generated and its coordinates are sent to the audio generation module, along with the device's current position and orientation.
The audio generator then produces a tone based on the difference between the device and target's positions and sends it to the audio output channel, which plays it back to the user.
The WiFi recording module is constantly monitoring the different parameter values of the device and target's positions, as well as the system's output, and records it to a remotely stored datafile.

\begin{figure}
  \centering
  \includegraphics[clip=true, trim=0 120 80 50, width=1.0\columnwidth]{figures/pipeline.pdf}
  \caption{A diagram of the individual system components and their communication pipelines. $F$ indicates a feedback signal and $P$ a pose signal. }\label{fig:pipeline}
\end{figure}

\subsection{Audio Interface}

The audio component is responsible for conveying the 2D position of a target on the vertical plane, shown by \cref{fig:cam-coords}, in terms of pan and tilt angles.
The audio signal is a continuous \SI{10}{\hertz} sinusoidal sound wave that is played to the user through bone-conducting headphones.
We use a sinusoidal wave here because it is relatively simple to manipulate and analyse and will be replaced by a richer, more natural tone at a later stage.

The audio is spatialised in the pan dimension using the OpenAL\footnote{\url{http://openal.org}} sound library's HRTF, while the tilt angle is conveyed by varying the audio tone's pitch.
We use this approach because the external set of bone-conducting headphones plays the sound through a user's cheekbones instead of their outer ears, bypassing the ears' pennae that provide the ability to localise elevated sound sources~\cite{roffler1968factors,algazi2001elevation}.
This makes an HRTF less effective in conveying the target's elevation and therefore needs to be conveyed using another method; by varying the tone's pitch in this case.
The difference between the target's angular position and the device's angular orientation are used to generate the audio signals.

\subsubsection{Pan Direction}

The pan angle ($\phi$) describes the angle which the user needs to rotate the camera vector, $C$, around the $z$-axis shown in \cref{fig:cam-coords}, i.e.\ how far the target is to the left or right of the user.
We use an HRTF to add a spatial element to the audio tone that is played to the user, placing the sound source at the target's location. 
Humans determine a sound source's pan direction by subconsciously analysing the inter-aural time difference (IID) between a sound reaching both of your ears: the greater the IID, the greater the perceived angular distance to the sound source~\cite{wightman1992dominant}.

We implement OpenAL's default HRTF, based on the MIT's KEMAR dataset\footnote{\url{http://sound.media.mit.edu/resources/KEMAR.html}}, to generate a sinusoidal sound wave based on the difference between the user and target's positions.
The library takes position values as input and outputs a tone based on the angle between the two position vectors.

\subsubsection{Tilt Direction}

The system adjusts the tone's pitch ($f$) to convey the target's tilt angle ($\theta$) w.r.t.\ $C$. 
Here, a high pitch means the target is above $C$ (i.e.\ the user should look up) and a low pitch means the target is below $C$ (the user should look down).
This high/low association scheme is chosen because humans naturally associate high-pitched sounds with higher objects and lower-pitched noises with lower objects~\cite{pratt1930spatial,blauert1997spatial}.
We use a logarithmic, octave-based gain function for the pitch, since an increase in octave provides a distinct perceptible change while keeping the timbre roughly constant~\cite{shepard1964circularity}.

We wish to determine how the gradient, $m$, of the pitch gain as a function of $\theta$ (i.e.\ $m = \frac{df}{d\theta}$) affects a user's performance, e.g.\ does a greater $m$ lead to an increased target acquisition rate?
For this we select 3 different pitch settings, referred to henceforth as $m_l$, $m_m$ and $m_h$ for the smallest, middle and highest gain gradients respectively. 
To find these gradients, we set the maximum and minimum limits for $\theta$ and the maximum and minimum frequencies for $f$ for each respective setting.
Furthermore, for the sake of consistency and between the different settings, each gradient generates the same pitch when $C$ is level with the target's tilt.

After practical tests with the device and headphones, we set the neutral, on-target tone to $f_{512} =$ \SI{512}{\hertz} for its audibility.
For $m_l$, we set $f_{max}$ and $f_{min}$ to 1 octave higher and lower than $f_{512}$, giving limits of \SI{1024}{\hertz} and \SI{256}{\hertz} respectively.
For $m_m$, the $f_{max}$ and $f_{min}$ are set to 2 octaves higher and lower than $f_{512}$ (\SI{2048}{\hertz} and \SI{128}{\hertz}) and $m_h$ to 3 octaves higher and lower (\SI{4096}{\hertz} and \SI{64}{\hertz}).
These limits were selected for practical reasons, because the bone conducting headphones have low volume gain at the frequency spectrum extremes, making it difficult to hear. 
The values of $m_l, m_m, m_h$ are visualised in \cref{fig:pitch-gain}.

\begin{figure}
  \centering
  \includegraphics[clip, trim=0 0 0 40, width=1.0\columnwidth]{figures/pitch_gain_functions.png}
  \caption{The pitch gain functions for the 3 different gain settings ($y$-axis in log scale). }\label{fig:pitch-gain}
\end{figure}

We set the range of angles the system can convey to $[-\frac{\pi}{2}; \frac{\pi}{2}]$; anything outside this range implies that the target is behind the user and was not considered in this series of experiments. 
The \SI{0}{\radian} position is directly in front of the user.
With these limits, $f$ can be determined with

\begin{equation}\label{eq:frequency}
  f(\theta) = 2^{b\theta + c},
\end{equation}
where $b$ and $c$ for each setting can be calculated according to the given setting's pitch and angle limits.% and a linear graph equation (i.e.\ $f - f_1 = b(\theta - \theta_1), b=\frac{\Delta f}{\Delta\theta}$).

\section{Target Search Experiment}\label{sec:experiments}

We conducted a set of experiments to evaluate the audio interface's effectiveness at driving a user to point the camera at a target in the pan and tilt dimensions.
Furthermore, we wish to see how changes to $m$ affect a user's performance.

We recruited 2 groups of participants for the experiments.
Group $G1$ consists of 42 young adults (mostly students) with healthy eyesight, with 10 males and 32 females being $20\pm1.5$ years old on average.
This group was blindfolded throughout the experiment and forms the bulk of the participant population.
Group $G2$ is made up of 3 legally blind adults with little to no light perception (3 males, average age of $40\pm8$ years). 
This group makes up less than 10\% of the total experiment population, but provided valuable subjective feedback that will help improve the system before evaluating it with a larger group of people with vision impairments. 
All participants joined the experiment on a voluntary basis with written consent and have no other impairments that affected their interpretation of the audio guidance cues.

For the experiment, the participants were given a Tango tablet device, running an app written for this experiment, and a pair of AfterShockz Sportz3' bone conducting headphones (both pictured in use in \cref{fig:participant}) and were blindfolded where appropriate.
The app generates a virtual target, one at a time at a constant \SI{2}{\metre} from the participant, and uses the device's pose measurements to determine $\phi$ and $\theta$.
These angles are then used to generate and present the user with the audio signal to direct them to the target. 
As the participant rotates the camera, $\phi$ and $\theta$ are adjusted and the audio signal changes to reflect these changes.
When a participant believes that they are pointing at the target, they tap the screen which generates a new target target which they are then directed to. 
Each participant was given 28 targets in total that were randomly generated and equally spread across 4 quadrants to avoid clustering. 
After the 28 targets were found, $m$ was adjusted and the experiment was repeated. 
The order in which the gradient setting was changed for each participant was randomised to minimise the learning effect.
The app streams various data on the current device and target poses to a laptop computer via a WiFi connection in real-time. 
Before the experiment session started, each participant was given some time to familiarise themselves with the device and was given the opportunity to learn what the $f_{512}$ `on-target' and on-centre tone sounds like that will mean that they are pointing at the target.

The measure of performance here is the accuracy with which a user points the camera towards a target.
The accuracy is taken as the angular difference between $C$ and the target's position at the time the participant tapped the screen.
A smaller error indicates more accurate results.
The pan and tilt results are separated to understand the effect of changing the gradient setting has on either dimension's accuracy.
To minimise the speed/accuracy biases in the results, we asked the participants to focus on finding the targets as well as they could without worrying about the time it took.

\section{Results and Discussion}\label{sec:results}

The results for the participants are given by the 2D histograms in \cref{fig:err-results}, where the pan error is on the abscissa and the tilt error on the ordinates.
A set of box-plots of the angular errors' median and standard deviations are given in \cref{fig:err-boxplot} for the $m_l, m_m$ and $m_h$ settings. 
The results for $G1$ are summarised in \cref{tab:results}, while the results for $G2$ are given in \cref{tab:vi-results}.

\begin{figure}
  \centering
  \subfloat[Results for the $m_l$ setting.]{\label{fig:err-results-lo}\includegraphics[clip, trim=450 0 450 110, width=0.8\columnwidth]{figures/err_lo.png}}

  \subfloat[Results for the $m_m$ setting.]{\label{fig:err-results-med}\includegraphics[clip, trim=450 0 450 110, width=0.8\columnwidth]{figures/err_med.png}}

  \subfloat[Results for the $m_h$ setting.]{\label{fig:err-results-hi}\includegraphics[clip, trim=450 0 450 110, width=0.8\columnwidth]{figures/err_hi.png}}

  \caption{Angular error data for the 3 gradient settings. }\label{fig:err-results}
\end{figure}

\begin{figure}
  \centering
  \subfloat[The medians. ]{\label{fig:err-boxplot-median}\includegraphics[clip, trim=20 -70 100 100, width=1.0\columnwidth]{figures/err_boxplot_medians.png}}

  \subfloat[The standard deviations. ]{\label{fig:err-boxplot-std}\includegraphics[clip, trim=20 -70 100 100, width=1.0\columnwidth]{figures/err_boxplot_std.png}}

  \caption{Boxplots for the error medians and standard deviations. }\label{fig:err-boxplot}
\end{figure}

\begin{table}
  \centering
  \caption{A table of the pan and tilt angular error results collected from $G1$. The results include the absolute average error ($\mu_{abs}$), average error ($\mu$) and standard deviations ($\sigma$) given in radians, as well as the Spearman ($R$) correlation scores.}\label{tab:results}
  \begin{tabular}{llcccc}
    \toprule
    \multicolumn{2}{c}{} & $\mu_{abs}$ & $\mu$ & $\sigma$ & $R$ \\\midrule
	 & $m_l$ & 0.23 & -0.02 & 0.32 & 0.75 \\%\cmidrule{2-6}
    Pan  & $m_m$ & 0.21 & -0.01 & 0.28 & 0.77 \\%\cmidrule{2-6}
	 & $m_h$ & 0.22 & -0.05 & 0.31 & 0.75 \\\midrule
	 & $m_l$ & 0.40 & -0.14 & 0.46 & 0.38 \\%\cmidrule{2-6}
    Tilt & $m_m$ & 0.32 & -0.12 & 0.36 & 0.48 \\%\cmidrule{2-6}
	 & $m_h$ & 0.34 & -0.16 & 0.39 & 0.55 \\
    \bottomrule
  \end{tabular}
\end{table}

\begin{table}
  \centering
  \caption{A table summarising the target search results of the participants in $G2$ (P1 - P3). The results include the average error ($\mu$) and standard deviation ($\sigma$), given in radians in the pan and tilt dimensions.}\label{tab:vi-results}
  \begin{tabular}{llcccccc}
    \toprule
    \multicolumn{2}{c}{} & \multicolumn{2}{c}{$m_l$} & \multicolumn{2}{c}{$m_m$} & \multicolumn{2}{c}{$m_h$} \\
    \multicolumn{2}{c}{} & $\mu_l$ & $\sigma_l$ & $\mu_m$ & $\sigma_m$ & $\mu_h$ & $\sigma_h$ \\\midrule
	 & P1 &  0.25 & 0.44 &  0.21 & 0.45 &  0.29 & 0.40 \\% \cline{2-8}
    Pan  & P2 &  0.18 & 0.85 &  0.05 & 0.49 & -0.08 & 0.32 \\% \cline{2-8}
	 & P3 & -0.10 & 0.67 & -0.08 & 0.61 & -0.08 & 0.50 \\ \midrule
	 & P1 & -0.17 & 0.21 & -0.10 & 0.2  & -0.25 & 0.21 \\% \cline{2-8}
    Tilt & P2 & -0.87 & 0.26 & -0.05 & 0.21 & -0.15 & 0.18 \\% \cline{2-8}
	 & P3 & -0.57 & 0.31 & -0.57 & 0.45 & -0.38 & 0.39 \\% \hline
    \bottomrule
  \end{tabular}
\end{table}

\subsection{Pan Results}

\cref{fig:err-results} and the median and mean points from the boxplots in \cref{fig:err-boxplot} show that the data in the pan dimension are approximately normally distributed around zero.
However, a Shapiro-Wilkes test score below the critical threshold do not give enough confidence to consider the data as normally distributed and we therefore analyse the data through non-parametric techniques.

For the data collected from $G1$, there is a strong linear correlation between the participants' guesses and the targets' true pan angles, shown by the high Spearman correlation scores of approximately 0.75 ($p < 0.01$) for all of the datasets, with the $m_m$ setting displaying the marginally best result of 0.77.
The 3 settings have similar average errors and standard deviations, with the $m_m$ setting again producing the marginally best results with the smallest average error and standard deviation.
However, the differences are relatively small (approximately 6\%) and are not significant enough (Friedman test $p = 0.92$) to conclude that the pitch gradient has an effect on the target acquisition accuracy performance in the pan dimension.

The data collected from $G2$ follow a similar trend to $G1$: the error in the pan dimension is centred around \SI{0}{\radian} with similar standard deviations across all 3 settings.
Unfortunately, not enough samples could be collected to draw any final conclusions.
However, these results are in line with results from previous work~\cite{zwiers2001spatial}, indicating that the groups with partial and normal eyesight respond with similar levels of accuracy to spatialised sound when presented with simple tasks.

These results show that both groups managed to find the targets with reasonable level of accuracy and that the pan error is robust to variations in the pitch's rate of change; a useful result in support of the effectiveness of this audio interface.

\subsection{Tilt Results}\label{sec:tilt-results}

The data within the tilt dimension were also found to be non-normal following scores below the critical threshold from the Shapiro-Wilkes test.
The Spearman test for the data reveals a reasonably significant correlation between the participants' guesses and the actual locations of the targets, with correlation scores of approximately 0.38 for $m_l$, 0.48 for $m_m$ and the $m_h$ gradient giving the strongest correlation score with 0.55 ($p < 0.01$ for all 3 settings).
This indicates that the participants generally interpret the cues from the varying the pitch correctly and are mostly successful at pointing the camera in the correct tilt angle.

The average errors of the datasets are relatively close to one another, with the $m_l$ setting producing the largest absolute error and standard deviation at \SI{0.40}{\radian} and \SI{0.46}{\radian} respectively.
The $m_m$ and $m_h$ settings have similar absolute errors of \SI{0.32}{\radian} and \SI{0.34}{\radian} respectively.
This is in line with the correlation scores, with $m_l$ giving the worst result and $m_h$ giving the best results overall, with highest correlation score and relatively low absolute angular errors and standard deviation, marginally beating $m_m$.

Since the data are not normally spread and there is significant levels of noise within the data which may contaminate the mean values, we use the medians to analyse the datasets.
For this, the Friedman test is used on the participants' error medians and it was found that there is a significant difference between the 3 settings' datasets ($p < 0.01$).
A post-hoc analysis with a Wilcoxon signed rank test and a critical threshold with a Bonferroni correction applied ($\alpha=0.016=\frac{0.05}{3}$) shows that both the $m_m$ and $m_h$ settings are significantly different from the $m_l$ setting ($p=0.0014<\alpha$ for both settings), but that the $m_m$ and $m_h$ settings do not produce significantly different results from one another ($p=0.42>\alpha$).
This indicates there is an inflection point where increasing $m$ produces diminishing returns and that we cannot conclude that the $m_h$ setting is better than the $m_m$ gradient.
However, both settings conclusively produce more accurate results than the $m_l$ setting.

From \cref{fig:err-boxplot-median} it can be seen that the median data show a significant skewing to the negative side, indicating a potential bias amongst the experiment's participant-base that must be taken into account.
From the results, it is not completely clear what causes this negative bias in the pitch dimension.
However, a possible explanation is that the floor introduces a position constraint within a user's mind, since the target cannot appear below the ground.
It can potentially appear above the user's head though, giving variable upper an lower limits that depend on the user's height and individual perception.
Future improvements to our audio interface should consider a non-linear increase in pitch as a function of tilt angle instead of the linear one we used in this work to remedy this bias.

The results collected from $G1$ and $G2$ show that a varying pitch tone can convey the tilt angle of a target to a human using bone-conducting headphones with accuracy levels similar to that in literature~\cite{bujacz2011sonification,katz2011spatial,zotkin2004rendering}.
They also highlight a clear and significant difference between the 3 different pitch gain gradients, with the $m_l$ setting giving the lowest acquisition accuracy, while $m_m$ and $m_h$ give similar accuracy levels closest to the target's true tilt.

\subsection{Participants with Visual Impairment's Feedback}

After the experiments with the participants with visual impairments in $G2$ were completed, we took the opportunity to collect their feedback on the audio interface's current implementation and gather suggestions for improvements for future iterations.
When asked about their opinions on current assistive systems available on the market, the general consensus was that these systems are typically prohibitively expensive, too complex and cumbersome to be useful for everyday use.
Indeed, they seemed quite enthusiastic about a guidance system that they can use on a mobile phone.

They were generally satisfied with the guidance interface in its current form and had little problem learning how to use it.
However, they, along with the participants in $G1$, reported auditory fatigue caused by long-term exposure to the sinusoidal audio tone.
We were aware of this issue before the experiments were conducted and have considered replacing the simple sine wave tone with more natural and pleasant sounds.
For example, musical instruments can be used to play different notes and scales to indicate different elevation angles~\cite{brewster1998using}.
Furthermore, they suggested adding some voice component or tone prompt to confirm when they are on target, inform them of other important events or happenings or simply to intermittently assure them that the system is still active.
The former two suggestions were intentionally ignored in order to purely evaluate how accurately the audio tone can guide a user to the target.
Going forward, however, these suggestions will be helpful for a potential commercial release version of the guidance system to make the guidance instructions clearer and to intermittently reassure users that they are still on the right course, for example.

On the hardware, only one participant had reservations regarding the choice of bone-conducting headphones, citing cost concerns. 
However, he indicated that he could justify purchasing a set of bone-conducting headphones if he were to use them throughout the day to avoid cutting off environmental sounds.
None of the participants had any technical issues using the headphones or mobile device and were able to properly use and interpret the guidance instructions.

\section{Conclusion and Future Work}\label{sec:conclusion}

In this work, we present an audio-based guidance interface that uses a spatialised audio tone with varying pitch to direct a user with visual impairments to point a camera toward a target. 
We also present and discuss a set of experiments that were conducted to determine the interface's effectiveness and performance, as well as some subjective feedback that was collected from legally blind participants.

Using a set of bone-conducting headphones, we found that a spatialised audio tone with a varying pitch can successfully convey the pan and tilt angles of a target.
The angular errors made by the participants are in line with those from previous research using similar audio guidance instructions.
We also found that varying the pitch's rate of change influences the accuracy of the system in the tilt dimension without affecting the performance in the pan dimension.
The steeper, $m_h$ and $m_m$ pitch gradients were found to produce the best accuracy results in general.

Feedback from the participants with visual impairments was mostly positive and they agreed with the choice of bone-conduction headphones that do not inhibit normal hearing.
However, they noted that the sine wave audio signal induced audio fatigue after extended exposure, something noted by the other participants as well, and suggested that more full-bodied, natural sounds be used instead.
Using vocal feedback in conjunction with the current audio signal was also suggested.

These results are more interesting from a design point of view than an end-user's. 
Indeed, we have shown that this audio guidance interface is effective at directing an average person, but more data are needed with the intended audience in order to decisively conclude that this interface will be effective in an indoor guidance system for people with vision impairments. 
Nevertheless, these data are a good indication of what to expect when we evaluate this audio interface with a greater number of people with limited vision.
Once we have determined that the intended audience can effectively use this type of guidance instruction, we can start adding more user-oriented features (e.g.\ voice prompts, more pleasant tones, etc.) and integrate them into our guidance system prototype.

\bibliographystyle{splncs04}
\bibliography{bib}

\end{document}

### 2019-07-06 paper.tex
Added intro

@@
 \usepackage{multirow}
-%\usepackage{subcaption}
 \usepackage[caption=false]{subfig}
@@
 \begin{abstract}
-	Our aim is to build a portable active navigation system for people with visual impairments that uses active perception techniques and a combination of feedback modes to identify a scene and guide a user to their destination.
-	Such a system requires a non-visual feedback interface and in this paper we investigate the effectiveness of a spatial audio tone with a varying pitch component, played through bone-conducting headphones, in conveying the pan and tilt angles of a target to the user in a pointing task.
-	The secondary goal is to determine how variations to the pitch's rate of change affect the user's performance.
-	For this, we conducted a set of experiments with both blindfolded participants and participants with visual impairments and found that, with some limitations, a spatialised audio tone with a varying pitch component can successfully convey a target's pan and tilt angles.
+	%Our aim is to build a portable active navigation system for people with visual impairments that uses active perception techniques and a combination of feedback modes to identify a scene and guide a user to their destination.
+	%Such a system requires a non-visual feedback interface and in this paper we investigate the effectiveness of a spatial audio tone with a varying pitch component, played through bone-conducting headphones, in conveying the pan and tilt angles of a target to the user in a pointing task.
+	%The secondary goal is to determine how variations to the pitch's rate of change affect the user's performance.
+	%For this, we conducted a set of experiments with both blindfolded participants and participants with visual impairments and found that, with some limitations, a spatialised audio tone with a varying pitch component can successfully convey a target's pan and tilt angles.
 \keywords{Human-machine interface \and vision impairment \and spatialised sound \and varying pitch, pointing task}
@@
 In recent years, governments have spearheaded numerous initiatives to support people with disabilities and enable them to play a more active role in modern society.
-In the UK, the Royal National Institute of Blind People (RNIB) has prioritised enabling people with visual impairments to use some of the services and products many people take for granted, such as public transport and mobile apps~\cite{rnib-objectives}.
+The UK's Royal National Institute of Blind People (RNIB) for example has prioritised improving access to everyday services and products, such as public transport and mobile apps~\cite{rnib-objectives}.
 Improvements in modern computing have made it possible for new and innovative solutions to these problems to come to the fore.
-In particular, researchers in the active vision field have have made much progress in enabling machines to autonomously manipulate cameras to gather information about an environment for mapping and object and face tracking tasks~\cite{bajcsy2018revisiting}.
+In particular, researchers in the active vision field have have made much progress in enabling machines to autonomously manipulate cameras to gather information about an environment for mapping and object and object finding tasks~\cite{bajcsy2018revisiting}.
 There is, however, a significant research question about whether techniques from active machine vision can be applied to humans, i.e.\ can a machine identify a point of interest in a scene and direct a human, instead of an electronic servo, to focus on that point?
@@
 
-To this end, we are developing a mobile navigation system using a Google Project Tango tablet device\footnote{\url{https://en.wikipedia.org/wiki/Tango\_(platform)}}, pictured in \cref{fig:tango-headphone}, that caters to the perception limitations of people with vision impairments by using a non-visual interface.
-This system will ultimately allow users with visual impairments to reach a target location in the final leg of their journey, i.e.\ the so-called `last 10 yards' problem~\cite{google2016blind,bellotto2013}.
-The ideas here can also be extended and applied to mainstream navigation tools, such as Google Maps.
-A Tango-enabled device comes pre-equipped with powerful image-processing, localisation and depth-perception capabilities and is built on top of a standard Android platform, providing access to the entire set of input/output options that this operating system has to offer.
+The ActiVis project aims to deliver a mobile guidance system based on the Google Project Tango platform\footnote{\url{https://en.wikipedia.org/wiki/Tango\_(platform)}}, pictured in \cref{fig:tango-headphone}, that will ultimately be able to guide a user with vision impairments on the last leg of their journey, i.e.\ the so-called `last 10-yard problem'.
+The entire platform is based on a Google Project Tango device that provides access to powerful real-time localisation (through IMU measurements and landmark tracking) and image-processing facilities and provides access to Android's full range of interface tools and IO options.
+Furthermore, a set of bone-conducting headphones (pictured in \cref{fig:tango-headphone}) that are placed on a user's cheekbones instead of their ears and do not interfere with normal hearing, are used to transmit the audio signals to the user.
+The system will provide real-time guidance instructions for the user to follow and it is important that these instructions are easily understood and are not perceived as a cognitive burden. 
+Humans are naturally able to determine the 3D position of a sound source and by exploiting this ability, the system's guidance instructions can naturally be interpreted without posing a unnecessary burdensome cognitive load.
+A sound source can be spatialised by adjusting a tone's spectral make-up (elevation angle), time delay and level difference (pan angle) and the intensity (distance).
+Here only the pan and elevation positions are transmitted.
+However, since bone-conducting headphones bypass the outer ear structure, their spectral signature cannot properly be interpreted and we therefore convey the target's elevation angle by simply adjusting the tone's pitch.
+
+%For this work, we investigate implementing such an interface that uses audio signals to transmit guidance instructions by exploiting the human auditory and psychoacoustic systems. 
+%Initially the interface is implemented in 2D only (pan and tilt dimensions) to provide the user with audio instructions on where to point the device and our approach was validated through a set of simple pointing tasks.
 
@@
   \centering
-  \subfloat[]{\label{fig:tango-headphone}\includegraphics[width=0.5\columnwidth]{figures/tango_headphone.png}}
-  ~
-  \subfloat[]{\label{fig:participant}\includegraphics[width=0.35\columnwidth]{figures/vi_participant.pdf}}
-  \caption{Pictures of the Tango device and bone-conducting headphones the system is based on (left) and both in use in an experiment (right).}
+  %\subfloat[]{\label{fig:tango-headphone}\includegraphics[width=0.5\columnwidth]{figures/tango_headphone.png}}
+  \includegraphics[width=0.5\columnwidth]{figures/tango_headphone.png}}
+  %~
+  %\subfloat[]{\label{fig:participant}\includegraphics[width=0.35\columnwidth]{figures/vi_participant.pdf}}
+  \caption{Pictures of the Tango device and bone-conducting headphones the system is based on (left) and both in use in an experiment (right).}\label{fig:tango-headphone}
 \end{figure}
 
-For this work, we implement and evaluate an audio interface that will be integrated into the guidance system to provide navigation instructions to the user. 
-To evaluate this guidance mode, we conducted a set of experiments involving a target-search task where the participants were guided with the audio instructions to point a camera at a set of virtual targets.
-The system uses a set of bone-conducting headphones to provide the audio instructions, with the latter being used to avoid interfering with a user's normal hearing function; something people with limited vision typically rely upon.
-However, this type of headphones bypass the external structure of the ear which provide the ability to localise a sound source's elevation.
-We therefore use a tone of varying pitch to convey the target's tilt position w.r.t.\ the camera's pointing vector, $C$, pictured in \cref{fig:cam-coords}.
-The tone is spatialised to provide the target's position in the pan dimension.
-A similar interface can be found in~\cite{durette2008visuo}, where they used a varying pitch component without spatialising the tone. 
-However, they did not investigate this interface's target-finding accuracy.
-Furthermore, we wish to determine how modifying the pitch's rate of change will affect target-finding performance. 
-Therefore, this work contributes to the body of knowledge by determining how accurately a spatialised tone with varying pitch can direct a user towards a target, and how the pitch's rate of change affects this accuracy.
+A similar approach was used in \todo{ref previous varying pitch study}, but the study did not focus on the interface's efficacy.
+Indeed, this work serves as an initial study into such an interface.
+The contribution of this work is an effective approach to addressing the shortcomings of bone-conducting headphones in transmitting spatialised sounds through a generic sound library, as well as initial experiments demonstrating its effectiveness.
 
-\begin{figure}
-  \centering
-  \includegraphics[width=0.4\columnwidth]{figures/camera_coordinate.pdf}
-  \caption{A diagram showing the device's frame of reference and the target's pan and tilt angles.}\label{fig:cam-coords}
-\end{figure}
+The rest of the paper is \todo{lay out rest of paper}
 
-The remainder of this paper is organised as follows: \cref{sec:lit-review} gives an overview of existing navigation systems and audio interfaces for people with visual impairments.
-\cref{sec:portable-navigation} discusses the navigation device's hardware and interface setup and describes how the target's position is conveyed to the user.
-The experiment we conducted are explained in \cref{sec:experiments} and the results are presented and discussed in \cref{sec:results}. 
-Finally, \cref{sec:conclusion} concludes this paper with a summary of the findings and comments on future work.
+\section{Previous Work}
 
-\section{Previous Work}\label{sec:lit-review}
+Work on guiding/directing people (vibration, audio, voice)
+Show preference studies
 
-Earlier attempts at improving a person with visual impairment's navigation experience involved outfitting the commonly-used walking cane with remote detection systems, such as sonar, radar and RFID tags, to transform a cane into an early-warning system~\cite{ulrich1997,marion2008batcane,faria2010electronic,willis2005}.
-Bluetooth tags and a smartphone can be used in a similar manner~\cite{sato2017navcog3}.
-However, the expense and effort of maintaining such tags make them less attractive for wide-scale adoption.
-Computer vision-based systems provide a good compromise between usability, cost and accuracy.
-One solution is to use RGB-D depth sensing cameras, which are becoming increasingly accurate and cost-effective, to build a 3D image map of the environment, and guiding a user through it~\cite{lee2015,rodriguez2012obstacle}.
-Alternatively, object recognition techniques can be used to detect various objects and landmarks, such as doors, staircases, etc., and communicate their relative location to the user~\cite{schauerte2012assistive,tian2013b,fiannaca2014headlock}, or use audio instructions to guide a user to fully capture an object of interest~\cite{vazquez2012helping}.
+Guidance instruction delivery modes using haptic, audio and vocal feedback for non-visual human-machine-interfaces (HMI) have been investigated before.
 
-An important feature of user-centric systems is a human-machine interface (HMI) that enables effective two-way communication between the system and the user.
-Surveys found that people with visual impairments prefer receiving instructions in the form of speech and haptic feedback cues~\cite{khoo2016multimodal,ross2000wearable,vazquez2012helping}.
-However, haptic feedback modes typically have a lower bandwidth when compared to audio feedback and also require the user to wear a special device in order to transmit the signals effectively.
-Work has also been done in translating a visual scene into non-visual formats, with so-called sensory substitution systems (e.g.
+Previous work on HRTF stuff
+
+\section{Interface Implementation}
+
+\subsection{Human Audio Localisation}
+
+Humans localise a sound source in 3 dimensions by considering cues recorded in one ear (monaural cues) and comparing cues derived at both ears (binaural cues).
+The binaural cues include time differences of arrival and intensity differences that help to determine a source's location on the horizontal plane.
+Monaural cues are taken from the interaction of the sound with the human anatomy, e.g.\ head, shoulders, outer ear, before it enters the ear canal.
+The user's anatomy acts as a filter and the direction the sound wave hits the body activates different filter resonances, attenuating or accentuating different frequency bands.
+When the modified sound enters the inner ear, the human brain is able to analyse the frequency response and accurately determine the position of the sound source on the vertical plane. 
+The distance to the source is simply derived as the intensity, or volume, of the source, i.e.\ a louder sound would appear closer to the user than a softer one does.
+
+When an audio signal is transmitted via a set of speakers or headsets, it can be transformed to mimic the characteristics of a natural sound sources before it is transmitted, thereby tricking the brain into believing a sound is located at some arbitrary position.
+Such a transformation can be done using a head-related transfer function (HRTF) and a pair of stereo headsets or speakers.
+An HRTF is a mathematical function that simulates the response signal of a human head and is derived by capturing key characteristics that affect the monaural and binaural responses, such as the user's hearing levels and head size, etc.
+Since hearing responses can be quite unique between user's, the best results would be observed if each user had their own customised HRTF.\
+Given the complicated process involved to capture the required user characteristics, making unique HRTFs is often an untenable solution and using average values (e.g.\ head measurements, height, etc.) have shown to produce acceptable results.\todo{cite HRTF papers}
+
+\subsection{Interface Design}
+
+The guidance target is presented to the user in terms of pan and elevation angles, indicating the angle adjustments that are required to point the device camera at the target location.
+Spatialised audio signals are well-suited to the task, displaying similar levels of performance to vocal feedback, but with less cognitive load and higher resolution.
+However, the audio guidance signals are delivered to the user with a set of bone-conducting headphones that bypass the user's outer-ear structure and since this structure plays an important role in localising a sound source's elevation, an alternative mode of conveying the elevation angle is required.
+For this work, we propose a simple linear adjustment to the signal's pitch as a function of the elevation angle.
+
+\subsubsection{Pan}
+
+Binaural comparison cues are used to
+
+\subsubsection{Elevation}
+
+\subsection{Implementation}
+
+\section{Experiment and Results}
+
+\subsection{Apparatus}
+
+\subsection{Procedure}
+
+\subsection{Results}
+
+\section{Conclusion}
+
+%\section{Previous Work}\label{sec:lit-review}
+
+%Earlier attempts at improving a person with visual impairment's navigation experience involved outfitting the commonly-used walking cane with remote detection systems, such as sonar, radar and RFID tags, to transform a cane into an early-warning system~\cite{ulrich1997,marion2008batcane,faria2010electronic,willis2005}.
+%Bluetooth tags and a smartphone can be used in a similar manner~\cite{sato2017navcog3}.
+%However, the expense and effort of maintaining such tags make them less attractive for wide-scale adoption.
+%Computer vision-based systems provide a good compromise between usability, cost and accuracy.
+%One solution is to use RGB-D depth sensing cameras, which are becoming increasingly accurate and cost-effective, to build a 3D image map of the environment, and guiding a user through it~\cite{lee2015,rodriguez2012obstacle}.
+%Alternatively, object recognition techniques can be used to detect various objects and landmarks, such as doors, staircases, etc., and communicate their relative location to the user~\cite{schauerte2012assistive,tian2013b,fiannaca2014headlock}, or use audio instructions to guide a user to fully capture an object of interest~\cite{vazquez2012helping}.
+
+%An important feature of user-centric systems is a human-machine interface (HMI) that enables effective two-way communication between the system and the user.
+%Surveys found that people with visual impairments prefer receiving instructions in the form of speech and haptic feedback cues~\cite{khoo2016multimodal,ross2000wearable,vazquez2012helping}.
+%However, haptic feedback modes typically have a lower bandwidth when compared to audio feedback and also require the user to wear a special device in order to transmit the signals effectively.
+%Work has also been done in translating a visual scene into non-visual formats, with so-called sensory substitution systems (e.g.
 `The Voice'~\cite{meijer2010}) and virtual audio reality (VAR) systems~\cite{frauenberger2003} reporting favourable results.
-However, The Voice has a very steep learning curve that has proven to be a significant barrier to entry and with VAR systems, it is not clear how unknown environments, where markers have not yet been encoded, would be handled and described to the user.
+%However, The Voice has a very steep learning curve that has proven to be a significant barrier to entry and with VAR systems, it is not clear how unknown environments, where markers have not yet been encoded, would be handled and described to the user.
 
-Previous experiments have determined that people are able to find the location of a stationary sound source with an error of $\pm$\SI{35}{\degree}, in both the pan and tilt dimensions~\cite{zwiers2001spatial}, for both early-blind and normally-sighted people.
-However, in~\cite{lewald2013exceptional,lessard1998early} it was found that the blind have a clear advantage in localisation accuracy over sighted people when presented with more complex tasks, such as targets in motion and narrow-band stimuli, with~\cite{lewald2013exceptional} reporting an average absolute localisation error of around \SI{10}{\degree} in the pan and tilt dimensions.
-Researchers have used simulated spatialised audio to inform the user of the sound source's direction~\cite{holland2002audiogps,kammoun2012navigation,rebillat2009smart,menelas2010audio,wilson2007swan,zotkin2004rendering}.
-In these works, a sound was played through a set of headphones and the source spatialised with a head-related transfer function (HRTF), tricking the listener into thinking the sound source was located at some arbitrary 3D location.
-There are also experimental results about how well users can find targets presented with spatial sound in the tilt and panning dimensions with normal headphones~\cite{katz2011spatial,zwiers2001spatial}, as well as work that has shown similar levels of localisation performance between normal and bone-conducting headphones in the pan dimension~\cite{macdonald2006spatial}.
-However, to our knowledge no extensive work or experiments have been done to determine how well users respond to tilt adjustment instructions using a tone with \textit{varying pitch}, in particular when using bone-conducting headphones, and how changes to this tone's behaviour affects performance.
+%Previous experiments have determined that people are able to find the location of a stationary sound source with an error of $\pm$\SI{35}{\degree}, in both the pan and tilt dimensions~\cite{zwiers2001spatial}, for both early-blind and normally-sighted people.
+%However, in~\cite{lewald2013exceptional,lessard1998early} it was found that the blind have a clear advantage in localisation accuracy over sighted people when presented with more complex tasks, such as targets in motion and narrow-band stimuli, with~\cite{lewald2013exceptional} reporting an average absolute localisation error of around \SI{10}{\degree} in the pan and tilt dimensions.
+%Researchers have used simulated spatialised audio to inform the user of the sound source's direction~\cite{holland2002audiogps,kammoun2012navigation,rebillat2009smart,menelas2010audio,wilson2007swan,zotkin2004rendering}.
+%In these works, a sound was played through a set of headphones and the source spatialised with a head-related transfer function (HRTF), tricking the listener into thinking the sound source was located at some arbitrary 3D location.
+%There are also experimental results about how well users can find targets presented with spatial sound in the tilt and panning dimensions with normal headphones~\cite{katz2011spatial,zwiers2001spatial}, as well as work that has shown similar levels of localisation performance between normal and bone-conducting headphones in the pan dimension~\cite{macdonald2006spatial}.
+%However, to our knowledge no extensive work or experiments have been done to determine how well users respond to tilt adjustment instructions using a tone with \textit{varying pitch}, in particular when using bone-conducting headphones, and how changes to this tone's behaviour affects performance.
 
-\section{Portable Navigation System}\label{sec:portable-navigation}
+%\section{Portable Navigation System}\label{sec:portable-navigation}
 
-\subsection{System Setup}
+%\subsection{System Setup}
 
-Our ultimate goal is to deliver a navigation system that can guide a user with visual impairments on the last leg of their journey (e.g.\ to a specific aisle and shelf in a shop) using only a mobile phone.
-A large amount of data need to be translated from a visual form into a format that is useful to people with limited or no vision.
-We therefore plan to use a combination of voice, audio and vibration cues to translate visual navigation data as effectively as possible and overcome the bandwidth limitation of the human ear. 
-However, for this paper, we only considered the spatialised audio mode and its variation in pitch in order to determine its effectiveness in conveying a target's pan and tilt angles.
+%Our ultimate goal is to deliver a navigation system that can guide a user with visual impairments on the last leg of their journey (e.g.\ to a specific aisle and shelf in a shop) using only a mobile phone.
+%A large amount of data need to be translated from a visual form into a format that is useful to people with limited or no vision.
+%We therefore plan to use a combination of voice, audio and vibration cues to translate visual navigation data as effectively as possible and overcome the bandwidth limitation of the human ear. 
+%However, for this paper, we only considered the spatialised audio mode and its variation in pitch in order to determine its effectiveness in conveying a target's pan and tilt angles.
 
-The system is based on a concept proposed in~\cite{bellotto2013,lock2017portable}, which uses a Google Tango device.
-The navigation system is based on a Google Tango tablet, an Android-based device that comes equipped with an RGB-D camera to sense colour and depth and combines an inertial measurement unit (IMU) with powerful and robust landmark recognition and image processing algorithms to localise itself in real-time.
-An added benefit of this platform is its familiar, compact form-factor which will help overcome the hurdle of user-acceptance and usability.
-The guidance instructions are presented through a set of bone-conducting headphones (see \cref{fig:tango-headphone}) that are placed externally on a user's head so that the system does not occlude the perception of real-world sounds and does not interfere with a user's normal hearing function.
+%The system is based on a concept proposed in~\cite{bellotto2013,lock2017portable}, which uses a Google Tango device.
+%The navigation system is based on a Google Tango tablet, an Android-based device that comes equipped with an RGB-D camera to sense colour and depth and combines an inertial measurement unit (IMU) with powerful and robust landmark recognition and image processing algorithms to localise itself in real-time.
+%An added benefit of this platform is its familiar, compact form-factor which will help overcome the hurdle of user-acceptance and usability.
+%The guidance instructions are presented through a set of bone-conducting headphones (see \cref{fig:tango-headphone}) that are placed externally on a user's head so that the system does not occlude the perception of real-world sounds and does not interfere with a user's normal hearing function.
 
-A diagram of the experimental system pipeline is shown in \cref{fig:pipeline}, where the arrows indicate the direction of the flow of information.
-When the user taps the Tango's screen, a new virtual target is generated and its coordinates are sent to the audio generation module, along with the device's current position and orientation.
-The audio generator then produces a tone based on the difference between the device and target's positions and sends it to the audio output channel, which plays it back to the user.
-The WiFi recording module is constantly monitoring the different parameter values of the device and target's positions, as well as the system's output, and records it to a remotely stored datafile.
+%A diagram of the experimental system pipeline is shown in \cref{fig:pipeline}, where the arrows indicate the direction of the flow of information.
+%When the user taps the Tango's screen, a new virtual target is generated and its coordinates are sent to the audio generation module, along with the device's current position and orientation.
+%The audio generator then produces a tone based on the difference between the device and target's positions and sends it to the audio output channel, which plays it back to the user.
+%The WiFi recording module is constantly monitoring the different parameter values of the device and target's positions, as well as the system's output, and records it to a remotely stored datafile.
 
-\begin{figure}
-  \centering
-  \includegraphics[clip=true, trim=0 120 80 50, width=1.0\columnwidth]{figures/pipeline.pdf}
-  \caption{A diagram of the individual system components and their communication pipelines. $F$ indicates a feedback signal and $P$ a pose signal. }\label{fig:pipeline}
-\end{figure}
+%\begin{figure}
+  %\centering
+  %\includegraphics[clip=true, trim=0 120 80 50, width=1.0\columnwidth]{figures/pipeline.pdf}
+  %\caption{A diagram of the individual system components and their communication pipelines. $F$ indicates a feedback signal and $P$ a pose signal. }\label{fig:pipeline}
+%\end{figure}
 
-\subsection{Audio Interface}
+%\subsection{Audio Interface}
 
-The audio component is responsible for conveying the 2D position of a target on the vertical plane, shown by \cref{fig:cam-coords}, in terms of pan and tilt angles.
-The audio signal is a continuous \SI{10}{\hertz} sinusoidal sound wave that is played to the user through bone-conducting headphones.
-We use a sinusoidal wave here because it is relatively simple to manipulate and analyse and will be replaced by a richer, more natural tone at a later stage.
+%The audio component is responsible for conveying the 2D position of a target on the vertical plane, shown by \cref{fig:cam-coords}, in terms of pan and tilt angles.
+%The audio signal is a continuous \SI{10}{\hertz} sinusoidal sound wave that is played to the user through bone-conducting headphones.
+%We use a sinusoidal wave here because it is relatively simple to manipulate and analyse and will be replaced by a richer, more natural tone at a later stage.
 
-The audio is spatialised in the pan dimension using the OpenAL\footnote{\url{http://openal.org}} sound library's HRTF, while the tilt angle is conveyed by varying the audio tone's pitch.
-We use this approach because the external set of bone-conducting headphones plays the sound through a user's cheekbones instead of their outer ears, bypassing the ears' pennae that provide the ability to localise elevated sound sources~\cite{roffler1968factors,algazi2001elevation}.
-This makes an HRTF less effective in conveying the target's elevation and therefore needs to be conveyed using another method; by varying the tone's pitch in this case.
-The difference between the target's angular position and the device's angular orientation are used to generate the audio signals.
+%The audio is spatialised in the pan dimension using the OpenAL\footnote{\url{http://openal.org}} sound library's HRTF, while the tilt angle is conveyed by varying the audio tone's pitch.
+%We use this approach because the external set of bone-conducting headphones plays the sound through a user's cheekbones instead of their outer ears, bypassing the ears' pennae that provide the ability to localise elevated sound sources~\cite{roffler1968factors,algazi2001elevation}.
+%This makes an HRTF less effective in conveying the target's elevation and therefore needs to be conveyed using another method; by varying the tone's pitch in this case.
+%The difference between the target's angular position and the device's angular orientation are used to generate the audio signals.
 
-\subsubsection{Pan Direction}
+%\subsubsection{Pan Direction}
 
-The pan angle ($\phi$) describes the angle which the user needs to rotate the camera vector, $C$, around the $z$-axis shown in \cref{fig:cam-coords}, i.e.\ how far the target is to the left or right of the user.
-We use an HRTF to add a spatial element to the audio tone that is played to the user, placing the sound source at the target's location. 
-Humans determine a sound source's pan direction by subconsciously analysing the inter-aural time difference (IID) between a sound reaching both of your ears: the greater the IID, the greater the perceived angular distance to the sound source~\cite{wightman1992dominant}.
+%The pan angle ($\phi$) describes the angle which the user needs to rotate the camera vector, $C$, around the $z$-axis shown in \cref{fig:cam-coords}, i.e.\ how far the target is to the left or right of the user.
+%We use an HRTF to add a spatial element to the audio tone that is played to the user, placing the sound source at the target's location. 
+%Humans determine a sound source's pan direction by subconsciously analysing the inter-aural time difference (IID) between a sound reaching both of your ears: the greater the IID, the greater the perceived angular distance to the sound source~\cite{wightman1992dominant}.
 
-We implement OpenAL's default HRTF, based on the MIT's KEMAR dataset\footnote{\url{http://sound.media.mit.edu/resources/KEMAR.html}}, to generate a sinusoidal sound wave based on the difference between the user and target's positions.
-The library takes position values as input and outputs a tone based on the angle between the two position vectors.
+%We implement OpenAL's default HRTF, based on the MIT's KEMAR dataset\footnote{\url{http://sound.media.mit.edu/resources/KEMAR.html}}, to generate a sinusoidal sound wave based on the difference between the user and target's positions.
+%The library takes position values as input and outputs a tone based on the angle between the two position vectors.
 
-\subsubsection{Tilt Direction}
+%\subsubsection{Tilt Direction}
 
-The system adjusts the tone's pitch ($f$) to convey the target's tilt angle ($\theta$) w.r.t.\ $C$. 
-Here, a high pitch means the target is above $C$ (i.e.\ the user should look up) and a low pitch means the target is below $C$ (the user should look down).
-This high/low association scheme is chosen because humans naturally associate high-pitched sounds with higher objects and lower-pitched noises with lower objects~\cite{pratt1930spatial,blauert1997spatial}.
-We use a logarithmic, octave-based gain function for the pitch, since an increase in octave provides a distinct perceptible change while keeping the timbre roughly constant~\cite{shepard1964circularity}.
+%The system adjusts the tone's pitch ($f$) to convey the target's tilt angle ($\theta$) w.r.t.\ $C$. 
+%Here, a high pitch means the target is above $C$ (i.e.\ the user should look up) and a low pitch means the target is below $C$ (the user should look down).
+%This high/low association scheme is chosen because humans naturally associate high-pitched sounds with higher objects and lower-pitched noises with lower objects~\cite{pratt1930spatial,blauert1997spatial}.
+%We use a logarithmic, octave-based gain function for the pitch, since an increase in octave provides a distinct perceptible change while keeping the timbre roughly constant~\cite{shepard1964circularity}.
 
-We wish to determine how the gradient, $m$, of the pitch gain as a function of $\theta$ (i.e.\ $m = \frac{df}{d\theta}$) affects a user's performance, e.g.\ does a greater $m$ lead to an increased target acquisition rate?
-For this we select 3 different pitch settings, referred to henceforth as $m_l$, $m_m$ and $m_h$ for the smallest, middle and highest gain gradients respectively. 
-To find these gradients, we set the maximum and minimum limits for $\theta$ and the maximum and minimum frequencies for $f$ for each respective setting.
-Furthermore, for the sake of consistency and between the different settings, each gradient generates the same pitch when $C$ is level with the target's tilt.
+%We wish to determine how the gradient, $m$, of the pitch gain as a function of $\theta$ (i.e.\ $m = \frac{df}{d\theta}$) affects a user's performance, e.g.\ does a greater $m$ lead to an increased target acquisition rate?
+%For this we select 3 different pitch settings, referred to henceforth as $m_l$, $m_m$ and $m_h$ for the smallest, middle and highest gain gradients respectively. 
+%To find these gradients, we set the maximum and minimum limits for $\theta$ and the maximum and minimum frequencies for $f$ for each respective setting.
+%Furthermore, for the sake of consistency and between the different settings, each gradient generates the same pitch when $C$ is level with the target's tilt.
 
-After practical tests with the device and headphones, we set the neutral, on-target tone to $f_{512} =$ \SI{512}{\hertz} for its audibility.
-For $m_l$, we set $f_{max}$ and $f_{min}$ to 1 octave higher and lower than $f_{512}$, giving limits of \SI{1024}{\hertz} and \SI{256}{\hertz} respectively.
-For $m_m$, the $f_{max}$ and $f_{min}$ are set to 2 octaves higher and lower than $f_{512}$ (\SI{2048}{\hertz} and \SI{128}{\hertz}) and $m_h$ to 3 octaves higher and lower (\SI{4096}{\hertz} and \SI{64}{\hertz}).
-These limits were selected for practical reasons, because the bone conducting headphones have low volume gain at the frequency spectrum extremes, making it difficult to hear. 
-The values of $m_l, m_m, m_h$ are visualised in \cref{fig:pitch-gain}.
+%After practical tests with the device and headphones, we set the neutral, on-target tone to $f_{512} =$ \SI{512}{\hertz} for its audibility.
+%For $m_l$, we set $f_{max}$ and $f_{min}$ to 1 octave higher and lower than $f_{512}$, giving limits of \SI{1024}{\hertz} and \SI{256}{\hertz} respectively.
+%For $m_m$, the $f_{max}$ and $f_{min}$ are set to 2 octaves higher and lower than $f_{512}$ (\SI{2048}{\hertz} and \SI{128}{\hertz}) and $m_h$ to 3 octaves higher and lower (\SI{4096}{\hertz} and \SI{64}{\hertz}).
+%These limits were selected for practical reasons, because the bone conducting headphones have low volume gain at the frequency spectrum extremes, making it difficult to hear. 
+%The values of $m_l, m_m, m_h$ are visualised in \cref{fig:pitch-gain}.
 
-\begin{figure}
-  \centering
-  \includegraphics[clip, trim=0 0 0 40, width=1.0\columnwidth]{figures/pitch_gain_functions.png}
-  \caption{The pitch gain functions for the 3 different gain settings ($y$-axis in log scale). }\label{fig:pitch-gain}
-\end{figure}
+%\begin{figure}
+  %\centering
+  %\includegraphics[clip, trim=0 0 0 40, width=1.0\columnwidth]{figures/pitch_gain_functions.png}
+  %\caption{The pitch gain functions for the 3 different gain settings ($y$-axis in log scale). }\label{fig:pitch-gain}
+%\end{figure}
 
-We set the range of angles the system can convey to $[-\frac{\pi}{2}; \frac{\pi}{2}]$; anything outside this range implies that the target is behind the user and was not considered in this series of experiments.
-The \SI{0}{\radian} position is directly in front of the user.
-With these limits, $f$ can be determined with
+%We set the range of angles the system can convey to $[-\frac{\pi}{2}; \frac{\pi}{2}]$; anything outside this range implies that the target is behind the user and was not considered in this series of experiments.
+%The \SI{0}{\radian} position is directly in front of the user.
+%With these limits, $f$ can be determined with
 
-\begin{equation}\label{eq:frequency}
-  f(\theta) = 2^{b\theta + c},
-\end{equation}
-where $b$ and $c$ for each setting can be calculated according to the given setting's pitch and angle limits.% and a linear graph equation (i.e.\ $f - f_1 = b(\theta - \theta_1), b=\frac{\Delta f}{\Delta\theta}$).
+%\begin{equation}\label{eq:frequency}
+  %f(\theta) = 2^{b\theta + c},
+%\end{equation}
+%where $b$ and $c$ for each setting can be calculated according to the given setting's pitch and angle limits.% and a linear graph equation (i.e.\ $f - f_1 = b(\theta - \theta_1), b=\frac{\Delta f}{\Delta\theta}$).
 
-\section{Target Search Experiment}\label{sec:experiments}
+%\section{Target Search Experiment}\label{sec:experiments}
 
-We conducted a set of experiments to evaluate the audio interface's effectiveness at driving a user to point the camera at a target in the pan and tilt dimensions.
-Furthermore, we wish to see how changes to $m$ affect a user's performance.
+%We conducted a set of experiments to evaluate the audio interface's effectiveness at driving a user to point the camera at a target in the pan and tilt dimensions.
+%Furthermore, we wish to see how changes to $m$ affect a user's performance.
 
-We recruited 2 groups of participants for the experiments.
-Group $G1$ consists of 42 young adults (mostly students) with healthy eyesight, with 10 males and 32 females being $20\pm1.5$ years old on average.
-This group was blindfolded throughout the experiment and forms the bulk of the participant population.
-Group $G2$ is made up of 3 legally blind adults with little to no light perception (3 males, average age of $40\pm8$ years). 
-This group makes up less than 10\% of the total experiment population, but provided valuable subjective feedback that will help improve the system before evaluating it with a larger group of people with vision impairments.
-All participants joined the experiment on a voluntary basis with written consent and have no other impairments that affected their interpretation of the audio guidance cues.
+%We recruited 2 groups of participants for the experiments.
+%Group $G1$ consists of 42 young adults (mostly students) with healthy eyesight, with 10 males and 32 females being $20\pm1.5$ years old on average.
+%This group was blindfolded throughout the experiment and forms the bulk of the participant population.
+%Group $G2$ is made up of 3 legally blind adults with little to no light perception (3 males, average age of $40\pm8$ years). 
+%This group makes up less than 10\% of the total experiment population, but provided valuable subjective feedback that will help improve the system before evaluating it with a larger group of people with vision impairments.
+%All participants joined the experiment on a voluntary basis with written consent and have no other impairments that affected their interpretation of the audio guidance cues.
 
-For the experiment, the participants were given a Tango tablet device, running an app written for this experiment, and a pair of AfterShockz Sportz3' bone conducting headphones (both pictured in use in \cref{fig:participant}) and were blindfolded where appropriate.
-The app generates a virtual target, one at a time at a constant \SI{2}{\metre} from the participant, and uses the device's pose measurements to determine $\phi$ and $\theta$.
-These angles are then used to generate and present the user with the audio signal to direct them to the target. 
-As the participant rotates the camera, $\phi$ and $\theta$ are adjusted and the audio signal changes to reflect these changes.
-When a participant believes that they are pointing at the target, they tap the screen which generates a new target target which they are then directed to. 
-Each participant was given 28 targets in total that were randomly generated and equally spread across 4 quadrants to avoid clustering. 
-After the 28 targets were found, $m$ was adjusted and the experiment was repeated. 
-The order in which the gradient setting was changed for each participant was randomised to minimise the learning effect.
-The app streams various data on the current device and target poses to a laptop computer via a WiFi connection in real-time. 
-Before the experiment session started, each participant was given some time to familiarise themselves with the device and was given the opportunity to learn what the $f_{512}$ `on-target' and on-centre tone sounds like that will mean that they are pointing at the target.
+%For the experiment, the participants were given a Tango tablet device, running an app written for this experiment, and a pair of AfterShockz Sportz3' bone conducting headphones (both pictured in use in \cref{fig:participant}) and were blindfolded where appropriate.
+%The app generates a virtual target, one at a time at a constant \SI{2}{\metre} from the participant, and uses the device's pose measurements to determine $\phi$ and $\theta$.
+%These angles are then used to generate and present the user with the audio signal to direct them to the target. 
+%As the participant rotates the camera, $\phi$ and $\theta$ are adjusted and the audio signal changes to reflect these changes.
+%When a participant believes that they are pointing at the target, they tap the screen which generates a new target target which they are then directed to. 
+%Each participant was given 28 targets in total that were randomly generated and equally spread across 4 quadrants to avoid clustering. 
+%After the 28 targets were found, $m$ was adjusted and the experiment was repeated. 
+%The order in which the gradient setting was changed for each participant was randomised to minimise the learning effect.
+%The app streams various data on the current device and target poses to a laptop computer via a WiFi connection in real-time. 
+%Before the experiment session started, each participant was given some time to familiarise themselves with the device and was given the opportunity to learn what the $f_{512}$ `on-target' and on-centre tone sounds like that will mean that they are pointing at the target.
 
-The measure of performance here is the accuracy with which a user points the camera towards a target.
-The accuracy is taken as the angular difference between $C$ and the target's position at the time the participant tapped the screen.
-A smaller error indicates more accurate results.
-The pan and tilt results are separated to understand the effect of changing the gradient setting has on either dimension's accuracy.
-To minimise the speed/accuracy biases in the results, we asked the participants to focus on finding the targets as well as they could without worrying about the time it took.
+%The measure of performance here is the accuracy with which a user points the camera towards a target.
+%The accuracy is taken as the angular difference between $C$ and the target's position at the time the participant tapped the screen.
+%A smaller error indicates more accurate results.
+%The pan and tilt results are separated to understand the effect of changing the gradient setting has on either dimension's accuracy.
+%To minimise the speed/accuracy biases in the results, we asked the participants to focus on finding the targets as well as they could without worrying about the time it took.
 
-\section{Results and Discussion}\label{sec:results}
+%\section{Results and Discussion}\label{sec:results}
 
-The results for the participants are given by the 2D histograms in \cref{fig:err-results}, where the pan error is on the abscissa and the tilt error on the ordinates.
-A set of box-plots of the angular errors' median and standard deviations are given in \cref{fig:err-boxplot} for the $m_l, m_m$ and $m_h$ settings. 
-The results for $G1$ are summarised in \cref{tab:results}, while the results for $G2$ are given in \cref{tab:vi-results}.
+%The results for the participants are given by the 2D histograms in \cref{fig:err-results}, where the pan error is on the abscissa and the tilt error on the ordinates.
+%A set of box-plots of the angular errors' median and standard deviations are given in \cref{fig:err-boxplot} for the $m_l, m_m$ and $m_h$ settings. 
+%The results for $G1$ are summarised in \cref{tab:results}, while the results for $G2$ are given in \cref{tab:vi-results}.
 
-\begin{figure}
-  \centering
-  \subfloat[Results for the $m_l$ setting.]{\label{fig:err-results-lo}\includegraphics[clip, trim=450 0 450 110, width=0.8\columnwidth]{figures/err_lo.png}}
+%\begin{figure}
+  %\centering
+  %\subfloat[Results for the $m_l$ setting.]{\label{fig:err-results-lo}\includegraphics[clip, trim=450 0 450 110, width=0.8\columnwidth]{figures/err_lo.png}}
 
-  \subfloat[Results for the $m_m$ setting.]{\label{fig:err-results-med}\includegraphics[clip, trim=450 0 450 110, width=0.8\columnwidth]{figures/err_med.png}}
+  %\subfloat[Results for the $m_m$ setting.]{\label{fig:err-results-med}\includegraphics[clip, trim=450 0 450 110, width=0.8\columnwidth]{figures/err_med.png}}
 
-  \subfloat[Results for the $m_h$ setting.]{\label{fig:err-results-hi}\includegraphics[clip, trim=450 0 450 110, width=0.8\columnwidth]{figures/err_hi.png}}
+  %\subfloat[Results for the $m_h$ setting.]{\label{fig:err-results-hi}\includegraphics[clip, trim=450 0 450 110, width=0.8\columnwidth]{figures/err_hi.png}}
 
-  \caption{Angular error data for the 3 gradient settings. }\label{fig:err-results}
-\end{figure}
+  %\caption{Angular error data for the 3 gradient settings. }\label{fig:err-results}
+%\end{figure}
 
-\begin{figure}
-  \centering
-  \subfloat[The medians. ]{\label{fig:err-boxplot-median}\includegraphics[clip, trim=20 -70 100 100, width=1.0\columnwidth]{figures/err_boxplot_medians.png}}
+%\begin{figure}
+  %\centering
+  %\subfloat[The medians. ]{\label{fig:err-boxplot-median}\includegraphics[clip, trim=20 -70 100 100, width=1.0\columnwidth]{figures/err_boxplot_medians.png}}
 
-  \subfloat[The standard deviations. ]{\label{fig:err-boxplot-std}\includegraphics[clip, trim=20 -70 100 100, width=1.0\columnwidth]{figures/err_boxplot_std.png}}
+  %\subfloat[The standard deviations. ]{\label{fig:err-boxplot-std}\includegraphics[clip, trim=20 -70 100 100, width=1.0\columnwidth]{figures/err_boxplot_std.png}}
 
-  \caption{Boxplots for the error medians and standard deviations. }\label{fig:err-boxplot}
-\end{figure}
+  %\caption{Boxplots for the error medians and standard deviations. }\label{fig:err-boxplot}
+%\end{figure}
 
-\begin{table}
-  \centering
-  \caption{A table of the pan and tilt angular error results collected from $G1$.
+%\begin{table}
+  %\centering
+  %\caption{A table of the pan and tilt angular error results collected from $G1$.
 The results include the absolute average error ($\mu_{abs}$), average error ($\mu$) and standard deviations ($\sigma$) given in radians, as well as the Spearman ($R$) correlation scores.}\label{tab:results}
-  \begin{tabular}{llcccc}
-    \toprule
-    \multicolumn{2}{c}{} & $\mu_{abs}$ & $\mu$ & $\sigma$ & $R$ \\\midrule
-	 & $m_l$ & 0.23 & -0.02 & 0.32 & 0.75 \\%\cmidrule{2-6}
-    Pan  & $m_m$ & 0.21 & -0.01 & 0.28 & 0.77 \\%\cmidrule{2-6}
-	 & $m_h$ & 0.22 & -0.05 & 0.31 & 0.75 \\\midrule
-	 & $m_l$ & 0.40 & -0.14 & 0.46 & 0.38 \\%\cmidrule{2-6}
-    Tilt & $m_m$ & 0.32 & -0.12 & 0.36 & 0.48 \\%\cmidrule{2-6}
-	 & $m_h$ & 0.34 & -0.16 & 0.39 & 0.55 \\
-    \bottomrule
-  \end{tabular}
-\end{table}
+  %\begin{tabular}{llcccc}
+    %\toprule
+    %\multicolumn{2}{c}{} & $\mu_{abs}$ & $\mu$ & $\sigma$ & $R$ \\\midrule
+	 %& $m_l$ & 0.23 & -0.02 & 0.32 & 0.75 \\%\cmidrule{2-6}
+    %Pan  & $m_m$ & 0.21 & -0.01 & 0.28 & 0.77 \\%\cmidrule{2-6}
+	 %& $m_h$ & 0.22 & -0.05 & 0.31 & 0.75 \\\midrule
+	 %& $m_l$ & 0.40 & -0.14 & 0.46 & 0.38 \\%\cmidrule{2-6}
+    %Tilt & $m_m$ & 0.32 & -0.12 & 0.36 & 0.48 \\%\cmidrule{2-6}
+	 %& $m_h$ & 0.34 & -0.16 & 0.39 & 0.55 \\
+    %\bottomrule
+  %\end{tabular}
+%\end{table}
 
-\begin{table}
-  \centering
-  \caption{A table summarising the target search results of the participants in $G2$ (P1 - P3).
+%\begin{table}
+  %\centering
+  %\caption{A table summarising the target search results of the participants in $G2$ (P1 - P3).
 The results include the average error ($\mu$) and standard deviation ($\sigma$), given in radians in the pan and tilt dimensions.}\label{tab:vi-results}
-  \begin{tabular}{llcccccc}
-    \toprule
-    \multicolumn{2}{c}{} & \multicolumn{2}{c}{$m_l$} & \multicolumn{2}{c}{$m_m$} & \multicolumn{2}{c}{$m_h$} \\
-    \multicolumn{2}{c}{} & $\mu_l$ & $\sigma_l$ & $\mu_m$ & $\sigma_m$ & $\mu_h$ & $\sigma_h$ \\\midrule
-	 & P1 &  0.25 & 0.44 &  0.21 & 0.45 &  0.29 & 0.40 \\% \cline{2-8}
-    Pan  & P2 &  0.18 & 0.85 &  0.05 & 0.49 & -0.08 & 0.32 \\% \cline{2-8}
-	 & P3 & -0.10 & 0.67 & -0.08 & 0.61 & -0.08 & 0.50 \\ \midrule
-	 & P1 & -0.17 & 0.21 & -0.10 & 0.2  & -0.25 & 0.21 \\% \cline{2-8}
-    Tilt & P2 & -0.87 & 0.26 & -0.05 & 0.21 & -0.15 & 0.18 \\% \cline{2-8}
-	 & P3 & -0.57 & 0.31 & -0.57 & 0.45 & -0.38 & 0.39 \\% \hline
-    \bottomrule
-  \end{tabular}
-\end{table}
+  %\begin{tabular}{llcccccc}
+    %\toprule
+    %\multicolumn{2}{c}{} & \multicolumn{2}{c}{$m_l$} & \multicolumn{2}{c}{$m_m$} & \multicolumn{2}{c}{$m_h$} \\
+    %\multicolumn{2}{c}{} & $\mu_l$ & $\sigma_l$ & $\mu_m$ & $\sigma_m$ & $\mu_h$ & $\sigma_h$ \\\midrule
+	 %& P1 &  0.25 & 0.44 &  0.21 & 0.45 &  0.29 & 0.40 \\% \cline{2-8}
+    %Pan  & P2 &  0.18 & 0.85 &  0.05 & 0.49 & -0.08 & 0.32 \\% \cline{2-8}
+	 %& P3 & -0.10 & 0.67 & -0.08 & 0.61 & -0.08 & 0.50 \\ \midrule
+	 %& P1 & -0.17 & 0.21 & -0.10 & 0.2  & -0.25 & 0.21 \\% \cline{2-8}
+    %Tilt & P2 & -0.87 & 0.26 & -0.05 & 0.21 & -0.15 & 0.18 \\% \cline{2-8}
+	 %& P3 & -0.57 & 0.31 & -0.57 & 0.45 & -0.38 & 0.39 \\% \hline
+    %\bottomrule
+  %\end{tabular}
+%\end{table}
 
-\subsection{Pan Results}
+%\subsection{Pan Results}
 
-\cref{fig:err-results} and the median and mean points from the boxplots in \cref{fig:err-boxplot} show that the data in the pan dimension are approximately normally distributed around zero.
-However, a Shapiro-Wilkes test score below the critical threshold do not give enough confidence to consider the data as normally distributed and we therefore analyse the data through non-parametric techniques.
+%\cref{fig:err-results} and the median and mean points from the boxplots in \cref{fig:err-boxplot} show that the data in the pan dimension are approximately normally distributed around zero.
+%However, a Shapiro-Wilkes test score below the critical threshold do not give enough confidence to consider the data as normally distributed and we therefore analyse the data through non-parametric techniques.
 
-For the data collected from $G1$, there is a strong linear correlation between the participants' guesses and the targets' true pan angles, shown by the high Spearman correlation scores of approximately 0.75 ($p < 0.01$) for all of the datasets, with the $m_m$ setting displaying the marginally best result of 0.77.
-The 3 settings have similar average errors and standard deviations, with the $m_m$ setting again producing the marginally best results with the smallest average error and standard deviation.
-However, the differences are relatively small (approximately 6\%) and are not significant enough (Friedman test $p = 0.92$) to conclude that the pitch gradient has an effect on the target acquisition accuracy performance in the pan dimension.
+%For the data collected from $G1$, there is a strong linear correlation between the participants' guesses and the targets' true pan angles, shown by the high Spearman correlation scores of approximately 0.75 ($p < 0.01$) for all of the datasets, with the $m_m$ setting displaying the marginally best result of 0.77.
+%The 3 settings have similar average errors and standard deviations, with the $m_m$ setting again producing the marginally best results with the smallest average error and standard deviation.
+%However, the differences are relatively small (approximately 6\%) and are not significant enough (Friedman test $p = 0.92$) to conclude that the pitch gradient has an effect on the target acquisition accuracy performance in the pan dimension.
 
-The data collected from $G2$ follow a similar trend to $G1$: the error in the pan dimension is centred around \SI{0}{\radian} with similar standard deviations across all 3 settings.
-Unfortunately, not enough samples could be collected to draw any final conclusions.
-However, these results are in line with results from previous work~\cite{zwiers2001spatial}, indicating that the groups with partial and normal eyesight respond with similar levels of accuracy to spatialised sound when presented with simple tasks.
+%The data collected from $G2$ follow a similar trend to $G1$: the error in the pan dimension is centred around \SI{0}{\radian} with similar standard deviations across all 3 settings.
+%Unfortunately, not enough samples could be collected to draw any final conclusions.
+%However, these results are in line with results from previous work~\cite{zwiers2001spatial}, indicating that the groups with partial and normal eyesight respond with similar levels of accuracy to spatialised sound when presented with simple tasks.
 
-These results show that both groups managed to find the targets with reasonable level of accuracy and that the pan error is robust to variations in the pitch's rate of change; a useful result in support of the effectiveness of this audio interface.
+%These results show that both groups managed to find the targets with reasonable level of accuracy and that the pan error is robust to variations in the pitch's rate of change; a useful result in support of the effectiveness of this audio interface.
 
-\subsection{Tilt Results}\label{sec:tilt-results}
+%\subsection{Tilt Results}\label{sec:tilt-results}
 
-The data within the tilt dimension were also found to be non-normal following scores below the critical threshold from the Shapiro-Wilkes test.
-The Spearman test for the data reveals a reasonably significant correlation between the participants' guesses and the actual locations of the targets, with correlation scores of approximately 0.38 for $m_l$, 0.48 for $m_m$ and the $m_h$ gradient giving the strongest correlation score with 0.55 ($p < 0.01$ for all 3 settings).
-This indicates that the participants generally interpret the cues from the varying the pitch correctly and are mostly successful at pointing the camera in the correct tilt angle.
+%The data within the tilt dimension were also found to be non-normal following scores below the critical threshold from the Shapiro-Wilkes test.
+%The Spearman test for the data reveals a reasonably significant correlation between the participants' guesses and the actual locations of the targets, with correlation scores of approximately 0.38 for $m_l$, 0.48 for $m_m$ and the $m_h$ gradient giving the strongest correlation score with 0.55 ($p < 0.01$ for all 3 settings).
+%This indicates that the participants generally interpret the cues from the varying the pitch correctly and are mostly successful at pointing the camera in the correct tilt angle.
 
-The average errors of the datasets are relatively close to one another, with the $m_l$ setting producing the largest absolute error and standard deviation at \SI{0.40}{\radian} and \SI{0.46}{\radian} respectively.
-The $m_m$ and $m_h$ settings have similar absolute errors of \SI{0.32}{\radian} and \SI{0.34}{\radian} respectively.
-This is in line with the correlation scores, with $m_l$ giving the worst result and $m_h$ giving the best results overall, with highest correlation score and relatively low absolute angular errors and standard deviation, marginally beating $m_m$.
+%The average errors of the datasets are relatively close to one another, with the $m_l$ setting producing the largest absolute error and standard deviation at \SI{0.40}{\radian} and \SI{0.46}{\radian} respectively.
+%The $m_m$ and $m_h$ settings have similar absolute errors of \SI{0.32}{\radian} and \SI{0.34}{\radian} respectively.
+%This is in line with the correlation scores, with $m_l$ giving the worst result and $m_h$ giving the best results overall, with highest correlation score and relatively low absolute angular errors and standard deviation, marginally beating $m_m$.
 
-Since the data are not normally spread and there is significant levels of noise within the data which may contaminate the mean values, we use the medians to analyse the datasets.
-For this, the Friedman test is used on the participants' error medians and it was found that there is a significant difference between the 3 settings' datasets ($p < 0.01$).
-A post-hoc analysis with a Wilcoxon signed rank test and a critical threshold with a Bonferroni correction applied ($\alpha=0.016=\frac{0.05}{3}$) shows that both the $m_m$ and $m_h$ settings are significantly different from the $m_l$ setting ($p=0.0014<\alpha$ for both settings), but that the $m_m$ and $m_h$ settings do not produce significantly different results from one another ($p=0.42>\alpha$).
-This indicates there is an inflection point where increasing $m$ produces diminishing returns and that we cannot conclude that the $m_h$ setting is better than the $m_m$ gradient.
-However, both settings conclusively produce more accurate results than the $m_l$ setting.
+%Since the data are not normally spread and there is significant levels of noise within the data which may contaminate the mean values, we use the medians to analyse the datasets.
+%For this, the Friedman test is used on the participants' error medians and it was found that there is a significant difference between the 3 settings' datasets ($p < 0.01$).
+%A post-hoc analysis with a Wilcoxon signed rank test and a critical threshold with a Bonferroni correction applied ($\alpha=0.016=\frac{0.05}{3}$) shows that both the $m_m$ and $m_h$ settings are significantly different from the $m_l$ setting ($p=0.0014<\alpha$ for both settings), but that the $m_m$ and $m_h$ settings do not produce significantly different results from one another ($p=0.42>\alpha$).
+%This indicates there is an inflection point where increasing $m$ produces diminishing returns and that we cannot conclude that the $m_h$ setting is better than the $m_m$ gradient.
+%However, both settings conclusively produce more accurate results than the $m_l$ setting.
 
-From \cref{fig:err-boxplot-median} it can be seen that the median data show a significant skewing to the negative side, indicating a potential bias amongst the experiment's participant-base that must be taken into account.
-From the results, it is not completely clear what causes this negative bias in the pitch dimension.
-However, a possible explanation is that the floor introduces a position constraint within a user's mind, since the target cannot appear below the ground.
-It can potentially appear above the user's head though, giving variable upper an lower limits that depend on the user's height and individual perception.
-Future improvements to our audio interface should consider a non-linear increase in pitch as a function of tilt angle instead of the linear one we used in this work to remedy this bias.
+%From \cref{fig:err-boxplot-median} it can be seen that the median data show a significant skewing to the negative side, indicating a potential bias amongst the experiment's participant-base that must be taken into account.
+%From the results, it is not completely clear what causes this negative bias in the pitch dimension.
+%However, a possible explanation is that the floor introduces a position constraint within a user's mind, since the target cannot appear below the ground.
+%It can potentially appear above the user's head though, giving variable upper an lower limits that depend on the user's height and individual perception.
+%Future improvements to our audio interface should consider a non-linear increase in pitch as a function of tilt angle instead of the linear one we used in this work to remedy this bias.
 
-The results collected from $G1$ and $G2$ show that a varying pitch tone can convey the tilt angle of a target to a human using bone-conducting headphones with accuracy levels similar to that in literature~\cite{bujacz2011sonification,katz2011spatial,zotkin2004rendering}.
-They also highlight a clear and significant difference between the 3 different pitch gain gradients, with the $m_l$ setting giving the lowest acquisition accuracy, while $m_m$ and $m_h$ give similar accuracy levels closest to the target's true tilt.
+%The results collected from $G1$ and $G2$ show that a varying pitch tone can convey the tilt angle of a target to a human using bone-conducting headphones with accuracy levels similar to that in literature~\cite{bujacz2011sonification,katz2011spatial,zotkin2004rendering}.
+%They also highlight a clear and significant difference between the 3 different pitch gain gradients, with the $m_l$ setting giving the lowest acquisition accuracy, while $m_m$ and $m_h$ give similar accuracy levels closest to the target's true tilt.
 
-\subsection{Participants with Visual Impairment's Feedback}
+%\subsection{Participants with Visual Impairment's Feedback}
 
-After the experiments with the participants with visual impairments in $G2$ were completed, we took the opportunity to collect their feedback on the audio interface's current implementation and gather suggestions for improvements for future iterations.
-When asked about their opinions on current assistive systems available on the market, the general consensus was that these systems are typically prohibitively expensive, too complex and cumbersome to be useful for everyday use.
-Indeed, they seemed quite enthusiastic about a guidance system that they can use on a mobile phone.
+%After the experiments with the participants with visual impairments in $G2$ were completed, we took the opportunity to collect their feedback on the audio interface's current implementation and gather suggestions for improvements for future iterations.
+%When asked about their opinions on current assistive systems available on the market, the general consensus was that these systems are typically prohibitively expensive, too complex and cumbersome to be useful for everyday use.
+%Indeed, they seemed quite enthusiastic about a guidance system that they can use on a mobile phone.
 
-They were generally satisfied with the guidance interface in its current form and had little problem learning how to use it.
-However, they, along with the participants in $G1$, reported auditory fatigue caused by long-term exposure to the sinusoidal audio tone.
-We were aware of this issue before the experiments were conducted and have considered replacing the simple sine wave tone with more natural and pleasant sounds.
-For example, musical instruments can be used to play different notes and scales to indicate different elevation angles~\cite{brewster1998using}.
-Furthermore, they suggested adding some voice component or tone prompt to confirm when they are on target, inform them of other important events or happenings or simply to intermittently assure them that the system is still active.
-The former two suggestions were intentionally ignored in order to purely evaluate how accurately the audio tone can guide a user to the target.
-Going forward, however, these suggestions will be helpful for a potential commercial release version of the guidance system to make the guidance instructions clearer and to intermittently reassure users that they are still on the right course, for example.
+%They were generally satisfied with the guidance interface in its current form and had little problem learning how to use it.
+%However, they, along with the participants in $G1$, reported auditory fatigue caused by long-term exposure to the sinusoidal audio tone.
+%We were aware of this issue before the experiments were conducted and have considered replacing the simple sine wave tone with more natural and pleasant sounds.
+%For example, musical instruments can be used to play different notes and scales to indicate different elevation angles~\cite{brewster1998using}.
+%Furthermore, they suggested adding some voice component or tone prompt to confirm when they are on target, inform them of other important events or happenings or simply to intermittently assure them that the system is still active.
+%The former two suggestions were intentionally ignored in order to purely evaluate how accurately the audio tone can guide a user to the target.
+%Going forward, however, these suggestions will be helpful for a potential commercial release version of the guidance system to make the guidance instructions clearer and to intermittently reassure users that they are still on the right course, for example.
 
-On the hardware, only one participant had reservations regarding the choice of bone-conducting headphones, citing cost concerns. 
-However, he indicated that he could justify purchasing a set of bone-conducting headphones if he were to use them throughout the day to avoid cutting off environmental sounds.
-None of the participants had any technical issues using the headphones or mobile device and were able to properly use and interpret the guidance instructions.
+%On the hardware, only one participant had reservations regarding the choice of bone-conducting headphones, citing cost concerns. 
+%However, he indicated that he could justify purchasing a set of bone-conducting headphones if he were to use them throughout the day to avoid cutting off environmental sounds.
+%None of the participants had any technical issues using the headphones or mobile device and were able to properly use and interpret the guidance instructions.
 
-\section{Conclusion and Future Work}\label{sec:conclusion}
+%\section{Conclusion and Future Work}\label{sec:conclusion}
 
-In this work, we present an audio-based guidance interface that uses a spatialised audio tone with varying pitch to direct a user with visual impairments to point a camera toward a target. 
-We also present and discuss a set of experiments that were conducted to determine the interface's effectiveness and performance, as well as some subjective feedback that was collected from legally blind participants.
+%In this work, we present an audio-based guidance interface that uses a spatialised audio tone with varying pitch to direct a user with visual impairments to point a camera toward a target. 
+%We also present and discuss a set of experiments that were conducted to determine the interface's effectiveness and performance, as well as some subjective feedback that was collected from legally blind participants.
 
-Using a set of bone-conducting headphones, we found that a spatialised audio tone with a varying pitch can successfully convey the pan and tilt angles of a target.
-The angular errors made by the participants are in line with those from previous research using similar audio guidance instructions.
-We also found that varying the pitch's rate of change influences the accuracy of the system in the tilt dimension without affecting the performance in the pan dimension.
-The steeper, $m_h$ and $m_m$ pitch gradients were found to produce the best accuracy results in general.
+%Using a set of bone-conducting headphones, we found that a spatialised audio tone with a varying pitch can successfully convey the pan and tilt angles of a target.
+%The angular errors made by the participants are in line with those from previous research using similar audio guidance instructions.
+%We also found that varying the pitch's rate of change influences the accuracy of the system in the tilt dimension without affecting the performance in the pan dimension.
+%The steeper, $m_h$ and $m_m$ pitch gradients were found to produce the best accuracy results in general.
 
-Feedback from the participants with visual impairments was mostly positive and they agreed with the choice of bone-conduction headphones that do not inhibit normal hearing.
-However, they noted that the sine wave audio signal induced audio fatigue after extended exposure, something noted by the other participants as well, and suggested that more full-bodied, natural sounds be used instead.
-Using vocal feedback in conjunction with the current audio signal was also suggested.
+%Feedback from the participants with visual impairments was mostly positive and they agreed with the choice of bone-conduction headphones that do not inhibit normal hearing.
+%However, they noted that the sine wave audio signal induced audio fatigue after extended exposure, something noted by the other participants as well, and suggested that more full-bodied, natural sounds be used instead.
+%Using vocal feedback in conjunction with the current audio signal was also suggested.
 
-These results are more interesting from a design point of view than an end-user's. 
-Indeed, we have shown that this audio guidance interface is effective at directing an average person, but more data are needed with the intended audience in order to decisively conclude that this interface will be effective in an indoor guidance system for people with vision impairments.
-Nevertheless, these data are a good indication of what to expect when we evaluate this audio interface with a greater number of people with limited vision.
-Once we have determined that the intended audience can effectively use this type of guidance instruction, we can start adding more user-oriented features (e.g.\ voice prompts, more pleasant tones, etc.) and integrate them into our guidance system prototype.
+%These results are more interesting from a design point of view than an end-user's. 
+%Indeed, we have shown that this audio guidance interface is effective at directing an average person, but more data are needed with the intended audience in order to decisively conclude that this interface will be effective in an indoor guidance system for people with vision impairments.
+%Nevertheless, these data are a good indication of what to expect when we evaluate this audio interface with a greater number of people with limited vision.
+%Once we have determined that the intended audience can effectively use this type of guidance instruction, we can start adding more user-oriented features (e.g.\ voice prompts, more pleasant tones, etc.) and integrate them into our guidance system prototype.
 

### 2019-07-08 paper.tex
Added pan description

@@
   \centering
-  %\subfloat[]{\label{fig:tango-headphone}\includegraphics[width=0.5\columnwidth]{figures/tango_headphone.png}}
-  \includegraphics[width=0.5\columnwidth]{figures/tango_headphone.png}}
-  %~
-  %\subfloat[]{\label{fig:participant}\includegraphics[width=0.35\columnwidth]{figures/vi_participant.pdf}}
+  \subfloat[]{\label{fig:tango-headphone}\includegraphics[width=0.4\columnwidth]{figures/tango_headphone.png}}
+  %\includegraphics[width=0.5\columnwidth]{figures/tango_headphone.png}}
+  ~
+  \subfloat[]{\label{fig:participant}\includegraphics[width=0.25\columnwidth]{figures/vi_participant.pdf}}
   \caption{Pictures of the Tango device and bone-conducting headphones the system is based on (left) and both in use in an experiment (right).}\label{fig:tango-headphone}
@@
 
-A similar approach was used in \todo{ref previous varying pitch study}, but the study did not focus on the interface's efficacy.
-Indeed, this work serves as an initial study into such an interface.
-The contribution of this work is an effective approach to addressing the shortcomings of bone-conducting headphones in transmitting spatialised sounds through a generic sound library, as well as initial experiments demonstrating its effectiveness.
+A similar approach was used in\cite{durette2008visuo}, but the study did not focus on or investigate the interface's efficacy.
+Indeed, this work serves as an initial study into an interface that can effectively communicate the pan and elevation angles of a target location through a set of bone-conducting headphones using a standard head-related transfer function (HRTF).
+Furthermore, we demonstrate our approach's effectiveness through a simple experiment.
+%The contribution of this work is an effective approach to addressing the shortcomings of playing spatialised sound through bone-conducting headphones, as well as initial experiments demonstrating its effectivenesst
+%Furthermore, we demonstrate that such an interface is effective at communicating
 
@@
 
-Humans localise a sound source in 3 dimensions by considering cues recorded in one ear (monaural cues) and comparing cues derived at both ears (binaural cues).
+Humans localise a sound source in 3 dimensions by considering cues recorded in one ear (monaural cues) and comparing cues derived at both ears (binaural cues)\cite{blauert1997spatial,blauert1969sound}.
 The binaural cues include time differences of arrival and intensity differences that help to determine a source's location on the horizontal plane.
@@
 
-The guidance target is presented to the user in terms of pan and elevation angles, indicating the angle adjustments that are required to point the device camera at the target location.
-Spatialised audio signals are well-suited to the task, displaying similar levels of performance to vocal feedback, but with less cognitive load and higher resolution.
-However, the audio guidance signals are delivered to the user with a set of bone-conducting headphones that bypass the user's outer-ear structure and since this structure plays an important role in localising a sound source's elevation, an alternative mode of conveying the elevation angle is required.
+The guidance target is presented to the user in terms of pan and elevation angles, indicating the angle adjustments that are required to point the device camera at the target location, as shown in \cref{fig:cam-coords}.
+Spatialised audio signals are well-suited to the task, displaying similar levels of performance to vocal feedback, but with less cognitive load and higher resolution\cite{klatzky2006cognitive}.
+However, the audio guidance signals are delivered to the user with a set of bone-conducting headphones that bypass the user's outer-ear structure and since this structure plays an important role in localising a sound source's elevation\cite{blauert1969sound}, an alternative mode of conveying the elevation angle is required.
 For this work, we propose a simple linear adjustment to the signal's pitch as a function of the elevation angle. 
+The pan angle can be conveyed by transforming the audio signal with an HRTF, and indeed it has been found that this dimension is unaffected by using bone-conducting headphones\cite{schonstein2008comparison,macdonald2006spatial,stanley2006lateralization}.\todo{cite bone cnducting work}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=0.3\columnwidth]{figures/camera_coordinate.pdf}
+  \caption{A diagram showing the device's frame of reference and the target's pan and tilt angles.}\label{fig:cam-coords}
+\end{figure}
+
+%The audio signal used for this work is a simple \SI{10}{\hertz} sinusoidal wave that is easy to manipulate and analyse.
+%People have trouble localising a pure sine wave given its lack of spectral features
+
+%The audio component is responsible for conveying the 2D position of a target on the vertical plane, shown by \cref{fig:cam-coords}, in terms of pan and tilt angles.
+%The audio signal is a continuous \SI{10}{\hertz} sinusoidal sound wave that is played to the user through bone-conducting headphones.
+%We use a sinusoidal wave here because it is relatively simple to manipulate and analyse and will be replaced by a richer, more natural tone at a later stage.
+
+%The audio is spatialised in the pan dimension using the OpenAL\footnote{\url{http://openal.org}} sound library's HRTF, while the tilt angle is conveyed by varying the audio tone's pitch.
+%We use this approach because the external set of bone-conducting headphones plays the sound through a user's cheekbones instead of their outer ears, bypassing the ears' pennae that provide the ability to localise elevated sound sources~\cite{roffler1968factors,algazi2001elevation}.
+%This makes an HRTF less effective in conveying the target's elevation and therefore needs to be conveyed using another method; by varying the tone's pitch in this case.
+%The difference between the target's angular position and the device's angular orientation are used to generate the audio signals.
 
@@
 
-Binaural comparison cues are used to
+The human audition system uses binaural comparison cues, such as the inter-aural time and level differences (ITD and ILD respectively), to localise a sound source on the horizontal plane.
+The ITD is the perceived time delay between the signal reaching both ears, while the ILD is the perceived volume difference in the signal.
+For example, a sound that comes from the individual's right, will hit the right ear first with a slightly higher volume.
+
+In this work, a pure sinusoidal wave oscillating at \SI{10}{\hertz} was used. 
+People typically have trouble localising a pure tone devoid of a rich spectral signature.
+However, the ITD and ILD is independent of the tone's spectral make-up and the elevation angle is given through a different mechanism.
+A pure sine wave is therefore sufficient to convey the target's pan angle.
+To transform the signal, we used OpenAL's default HRTF, based on the MIT's KEMAR dataset\cite{hiebert2005openal,gardner1995hrtf}, that uses the user and targets' positions as input and outputs a transformed audio signal.
+
+%The pan angle ($\phi$) describes the angle which the user needs to rotate the camera vector, $C$, around the $z$-axis shown in \cref{fig:cam-coords}, i.e.\ how far the target is to the left or right of the user.
+%We use an HRTF to add a spatial element to the audio tone that is played to the user, placing the sound source at the target's location. 
+%Humans determine a sound source's pan direction by subconsciously analysing the inter-aural time difference (IID) between a sound reaching both of your ears: the greater the IID, the greater the perceived angular distance to the sound source~\cite{wightman1992dominant}.
+
+%We implement OpenAL's default HRTF, based on the MIT's KEMAR dataset\footnote{\url{http://sound.media.mit.edu/resources/KEMAR.html}}, to generate a sinusoidal sound wave based on the difference between the user and target's positions.
+%The library takes position values as input and outputs a tone based on the angle between the two position vectors.
 
 \subsubsection{Elevation}
+
+%The system adjusts the tone's pitch ($f$) to convey the target's tilt angle ($\theta$) w.r.t.\ $C$. 
+%Here, a high pitch means the target is above $C$ (i.e.\ the user should look up) and a low pitch means the target is below $C$ (the user should look down).
+%This high/low association scheme is chosen because humans naturally associate high-pitched sounds with higher objects and lower-pitched noises with lower objects~\cite{pratt1930spatial,blauert1997spatial}.
+%We use a logarithmic, octave-based gain function for the pitch, since an increase in octave provides a distinct perceptible change while keeping the timbre roughly constant~\cite{shepard1964circularity}.
+
+%We wish to determine how the gradient, $m$, of the pitch gain as a function of $\theta$ (i.e.\ $m = \frac{df}{d\theta}$) affects a user's performance, e.g.\ does a greater $m$ lead to an increased target acquisition rate?
+%For this we select 3 different pitch settings, referred to henceforth as $m_l$, $m_m$ and $m_h$ for the smallest, middle and highest gain gradients respectively. 
+%To find these gradients, we set the maximum and minimum limits for $\theta$ and the maximum and minimum frequencies for $f$ for each respective setting.
+%Furthermore, for the sake of consistency and between the different settings, each gradient generates the same pitch when $C$ is level with the target's tilt.
+
+%After practical tests with the device and headphones, we set the neutral, on-target tone to $f_{512} =$ \SI{512}{\hertz} for its audibility.
+%For $m_l$, we set $f_{max}$ and $f_{min}$ to 1 octave higher and lower than $f_{512}$, giving limits of \SI{1024}{\hertz} and \SI{256}{\hertz} respectively.
+%For $m_m$, the $f_{max}$ and $f_{min}$ are set to 2 octaves higher and lower than $f_{512}$ (\SI{2048}{\hertz} and \SI{128}{\hertz}) and $m_h$ to 3 octaves higher and lower (\SI{4096}{\hertz} and \SI{64}{\hertz}).
+%These limits were selected for practical reasons, because the bone conducting headphones have low volume gain at the frequency spectrum extremes, making it difficult to hear. 
+%The values of $m_l, m_m, m_h$ are visualised in \cref{fig:pitch-gain}.
+
+%\begin{figure}
+  %\centering
+  %\includegraphics[clip, trim=0 0 0 40, width=1.0\columnwidth]{figures/pitch_gain_functions.png}
+  %\caption{The pitch gain functions for the 3 different gain settings ($y$-axis in log scale). }\label{fig:pitch-gain}
+%\end{figure}
+
+%We set the range of angles the system can convey to $[-\frac{\pi}{2}; \frac{\pi}{2}]$; anything outside this range implies that the target is behind the user and was not considered in this series of experiments.
+%The \SI{0}{\radian} position is directly in front of the user.
+%With these limits, $f$ can be determined with
+
+%\begin{equation}\label{eq:frequency}
+  %f(\theta) = 2^{b\theta + c},
+%\end{equation}
+%where $b$ and $c$ for each setting can be calculated according to the given setting's pitch and angle limits.% and a linear graph equation (i.e.\ $f - f_1 = b(\theta - \theta_1), b=\frac{\Delta f}{\Delta\theta}$).
 

### 2019-07-09 paper.tex
Added new elevation and hardware descriptions, as well as more references

@@
 \maketitle
+\setcounter{footnote}{0}
 
@@
 In recent years, governments have spearheaded numerous initiatives to support people with disabilities and enable them to play a more active role in modern society.
-The UK's Royal National Institute of Blind People (RNIB) for example has prioritised improving access to everyday services and products, such as public transport and mobile apps~\cite{rnib-objectives}.
+The UK's Royal National Institute of Blind People (RNIB) for example has prioritised improving access to everyday services and products, such as public transport and mobile apps~\cite{rnib2016uk}.
 Improvements in modern computing have made it possible for new and innovative solutions to these problems to come to the fore.
@@
 
-%For this work, we investigate implementing such an interface that uses audio signals to transmit guidance instructions by exploiting the human auditory and psychoacoustic systems. 
-%Initially the interface is implemented in 2D only (pan and tilt dimensions) to provide the user with audio instructions on where to point the device and our approach was validated through a set of simple pointing tasks.
-
 \begin{figure}
@@
   \subfloat[]{\label{fig:tango-headphone}\includegraphics[width=0.4\columnwidth]{figures/tango_headphone.png}}
-  %\includegraphics[width=0.5\columnwidth]{figures/tango_headphone.png}}
 ~
@@
 
-\section{Interface Implementation}
+\section{Interface Description}
+
+\subsection{Hardware Selection}
+
+Electronic navigation aids and guidance systems for people with vision impairments have had trouble appealing to the visually impaired community at large because of problems that are typical of academic research and prototypes: prohibitive costs and user unfriendly interfaces and hardware requirements, e.g.\ head-mounted cameras and GPS antennae~\cite{golledge2004stated,yusif2016older,arditi2013user}.
+Indeed, the traditional walking cane remains the go-to tool for navigation and obstacle avoidance. 
+With the ActiVis system, we hope to tackle the issue of cost, user-acceptance and usability by using as many as possible `standard' devices and hardware as possible, with the idea that no specialist hardware is required and the user will not stand out if seen walking in public with a phone in their hand and wearing a headset.
+The ActiVis system is based on a concept proposed in~\cite{bellotto2013multimodal,lock2017portable}, which uses a Google Tango device that is an Android-based device that comes equipped with an RGB-D camera to sense colour and depth and combines an inertial measurement unit (IMU) with powerful and robust landmark recognition and image processing algorithms to localise itself in real-time.
+An benefit of such a platform is its familiar, compact form-factor which will help overcome the hurdle of user-acceptance and usability.
+
+A set of bone-conducting headphones was selected as the audio transmission medium.
+These headphones sit on a user's cheekbones and conduct the audio signals through the skull into the inner ear, instead of through the outer ear like typical over-ear headphones do. 
+This has the benefit of not denying the user access to ambient sounds and noise, which could prove dangerous to a person with limited vision that relies on sound to detect oncoming vehicles and people, for example~\cite{lichtenstein2012headphone}.
+Open-back headphones that allow ambient noise through were also considered, but the AfterShockz headphones pictured in \cref{fig:tango-headphone} were ultimately selected since they do not interfere at all with incoming sound (open-back headphones are better than closed back ones in this regard, but still muffle incoming sound) and are also more discreet than the larger over-ear headphone models.
+
+%Our ultimate goal is to deliver a navigation system that can guide a user with visual impairments on the last leg of their journey (e.g.\ to a specific aisle and shelf in a shop) using only a mobile phone.
+%A large amount of data need to be translated from a visual form into a format that is useful to people with limited or no vision.
+%We therefore plan to use a combination of voice, audio and vibration cues to translate visual navigation data as effectively as possible and overcome the bandwidth limitation of the human ear. 
+%However, for this paper, we only considered the spatialised audio mode and its variation in pitch in order to determine its effectiveness in conveying a target's pan and tilt angles.
+
+%The system is based on a concept proposed in~\cite{bellotto2013,lock2017portable}, which uses a Google Tango device.
+%The navigation system is based on a Google Tango tablet, an Android-based device that comes equipped with an RGB-D camera to sense colour and depth and combines an inertial measurement unit (IMU) with powerful and robust landmark recognition and image processing algorithms to localise itself in real-time.
+%An added benefit of this platform is its familiar, compact form-factor which will help overcome the hurdle of user-acceptance and usability.
+%The guidance instructions are presented through a set of bone-conducting headphones (see \cref{fig:tango-headphone}) that are placed externally on a user's head so that the system does not occlude the perception of real-world sounds and does not interfere with a user's normal hearing function.
+
+%A diagram of the experimental system pipeline is shown in \cref{fig:pipeline}, where the arrows indicate the direction of the flow of information.
+%When the user taps the Tango's screen, a new virtual target is generated and its coordinates are sent to the audio generation module, along with the device's current position and orientation.
+%The audio generator then produces a tone based on the difference between the device and target's positions and sends it to the audio output channel, which plays it back to the user.
+%The WiFi recording module is constantly monitoring the different parameter values of the device and target's positions, as well as the system's output, and records it to a remotely stored datafile.
+
+%\begin{figure}
+  %\centering
+  %\includegraphics[clip=true, trim=0 120 80 50, width=1.0\columnwidth]{figures/pipeline.pdf}
+  %\caption{A diagram of the individual system components and their communication pipelines. $F$ indicates a feedback signal and $P$ a pose signal. }\label{fig:pipeline}
+%\end{figure}
 
@@
   \centering
-  \includegraphics[width=0.3\columnwidth]{figures/camera_coordinate.pdf}
-  \caption{A diagram showing the device's frame of reference and the target's pan and tilt angles.}\label{fig:cam-coords}
+  \subfloat[]{\label{fig:cam-coords}\includegraphics[width=0.3\columnwidth]{figures/camera_coordinate.pdf}}
+~
+  \subfloat[]{\label{fig:pitch-gain}\includegraphics[clip, trim=0 0 0 40, width=0.6\columnwidth]{figures/pitch_gain_functions.png}}
+  \caption{The reference system used by the guidance interface showing the camera vector and pan and elevation angles (left) and the pitch gain function used to convey the target's elevation angle (right).
+}
 \end{figure}
-
-%The audio signal used for this work is a simple \SI{10}{\hertz} sinusoidal wave that is easy to manipulate and analyse.
-%People have trouble localising a pure sine wave given its lack of spectral features
-
-%The audio component is responsible for conveying the 2D position of a target on the vertical plane, shown by \cref{fig:cam-coords}, in terms of pan and tilt angles.
-%The audio signal is a continuous \SI{10}{\hertz} sinusoidal sound wave that is played to the user through bone-conducting headphones.
-%We use a sinusoidal wave here because it is relatively simple to manipulate and analyse and will be replaced by a richer, more natural tone at a later stage.
-
-%The audio is spatialised in the pan dimension using the OpenAL\footnote{\url{http://openal.org}} sound library's HRTF, while the tilt angle is conveyed by varying the audio tone's pitch.
-%We use this approach because the external set of bone-conducting headphones plays the sound through a user's cheekbones instead of their outer ears, bypassing the ears' pennae that provide the ability to localise elevated sound sources~\cite{roffler1968factors,algazi2001elevation}.
-%This makes an HRTF less effective in conveying the target's elevation and therefore needs to be conveyed using another method; by varying the tone's pitch in this case.
-%The difference between the target's angular position and the device's angular orientation are used to generate the audio signals.
 
@@
 
-The human audition system uses binaural comparison cues, such as the inter-aural time and level differences (ITD and ILD respectively), to localise a sound source on the horizontal plane.
+The human audition system uses binaural comparison cues, such as the inter-aural time and level differences (ITD and ILD respectively), to localise a sound source on the horizontal plane~\cite{blauert1969sound}.
 The ITD is the perceived time delay between the signal reaching both ears, while the ILD is the perceived volume difference in the signal.
@@
 
-In this work, a pure sinusoidal wave oscillating at \SI{10}{\hertz} was used. 
+In this work, a pure sinusoidal wave refreshed at \SI{10}{\hertz} was used. 
 People typically have trouble localising a pure tone devoid of a rich spectral signature.
@@
 A pure sine wave is therefore sufficient to convey the target's pan angle.
-To transform the signal, we used OpenAL's default HRTF, based on the MIT's KEMAR dataset\cite{hiebert2005openal,gardner1995hrtf}, that uses the user and targets' positions as input and outputs a transformed audio signal.
-
-%The pan angle ($\phi$) describes the angle which the user needs to rotate the camera vector, $C$, around the $z$-axis shown in \cref{fig:cam-coords}, i.e.\ how far the target is to the left or right of the user.
-%We use an HRTF to add a spatial element to the audio tone that is played to the user, placing the sound source at the target's location. 
-%Humans determine a sound source's pan direction by subconsciously analysing the inter-aural time difference (IID) between a sound reaching both of your ears: the greater the IID, the greater the perceived angular distance to the sound source~\cite{wightman1992dominant}.
-
-%We implement OpenAL's default HRTF, based on the MIT's KEMAR dataset\footnote{\url{http://sound.media.mit.edu/resources/KEMAR.html}}, to generate a sinusoidal sound wave based on the difference between the user and target's positions.
-%The library takes position values as input and outputs a tone based on the angle between the two position vectors.
+To transform and spatialise the audio signal, we used OpenAL's default HRTF, based on the MIT's KEMAR dataset~\cite{hiebert2005openal,gardner1995hrtf}, that uses the user and targets' positions as input and outputs a transformed audio signal.
 
@@
 
-%The system adjusts the tone's pitch ($f$) to convey the target's tilt angle ($\theta$) w.r.t.\ $C$. 
-%Here, a high pitch means the target is above $C$ (i.e.\ the user should look up) and a low pitch means the target is below $C$ (the user should look down).
-%This high/low association scheme is chosen because humans naturally associate high-pitched sounds with higher objects and lower-pitched noises with lower objects~\cite{pratt1930spatial,blauert1997spatial}.
-%We use a logarithmic, octave-based gain function for the pitch, since an increase in octave provides a distinct perceptible change while keeping the timbre roughly constant~\cite{shepard1964circularity}.
+Using a generic HRTF, such as OpenAL's KEMAR implementation, with bone-conducting headphones has been shown to have trouble conveying the elevation angle of a sound source accurately~\cite{macdonald2006spatial,schonstein2008comparison}.
+To compensate for this, we convey the target's elevation angle by adjusting the tone's pitch (i.e.\ the sine wave's frequency) as a function of the elevation. 
+When the camera vector is at the target's elevation, the tone is set to neutral pitch.
+Then, when the target is above the camera vector, the pitch is increased and decreased when the target is below the camera vector.
+This high/low association scheme is used given humans' natural association of high-pitched sounds with elevated sound sources and lower-pitched sounds with source's below the individual's earline~\cite{pratt1930spatial,blauert1997spatial}.
+A logarithmic, octave and semi-tone based function is used to adjust the tone's pitch to ensure perceptible changes while keeping the timbre roughly constant~\cite{shepard1964circularity}.
+The pitch is updated at a rate of \SI{10}{\hertz} and changes as the user moves the device.
 
-%We wish to determine how the gradient, $m$, of the pitch gain as a function of $\theta$ (i.e.\ $m = \frac{df}{d\theta}$) affects a user's performance, e.g.\ does a greater $m$ lead to an increased target acquisition rate?
-%For this we select 3 different pitch settings, referred to henceforth as $m_l$, $m_m$ and $m_h$ for the smallest, middle and highest gain gradients respectively. 
-%To find these gradients, we set the maximum and minimum limits for $\theta$ and the maximum and minimum frequencies for $f$ for each respective setting.
-%Furthermore, for the sake of consistency and between the different settings, each gradient generates the same pitch when $C$ is level with the target's tilt.
-
-%After practical tests with the device and headphones, we set the neutral, on-target tone to $f_{512} =$ \SI{512}{\hertz} for its audibility.
-%For $m_l$, we set $f_{max}$ and $f_{min}$ to 1 octave higher and lower than $f_{512}$, giving limits of \SI{1024}{\hertz} and \SI{256}{\hertz} respectively.
-%For $m_m$, the $f_{max}$ and $f_{min}$ are set to 2 octaves higher and lower than $f_{512}$ (\SI{2048}{\hertz} and \SI{128}{\hertz}) and $m_h$ to 3 octaves higher and lower (\SI{4096}{\hertz} and \SI{64}{\hertz}).
-%These limits were selected for practical reasons, because the bone conducting headphones have low volume gain at the frequency spectrum extremes, making it difficult to hear. 
-%The values of $m_l, m_m, m_h$ are visualised in \cref{fig:pitch-gain}.
-
-%\begin{figure}
-  %\centering
-  %\includegraphics[clip, trim=0 0 0 40, width=1.0\columnwidth]{figures/pitch_gain_functions.png}
-  %\caption{The pitch gain functions for the 3 different gain settings ($y$-axis in log scale). }\label{fig:pitch-gain}
-%\end{figure}
-
-%We set the range of angles the system can convey to $[-\frac{\pi}{2}; \frac{\pi}{2}]$; anything outside this range implies that the target is behind the user and was not considered in this series of experiments.
-%The \SI{0}{\radian} position is directly in front of the user.
-%With these limits, $f$ can be determined with
-
-%\begin{equation}\label{eq:frequency}
-  %f(\theta) = 2^{b\theta + c},
-%\end{equation}
-%where $b$ and $c$ for each setting can be calculated according to the given setting's pitch and angle limits.% and a linear graph equation (i.e.\ $f - f_1 = b(\theta - \theta_1), b=\frac{\Delta f}{\Delta\theta}$).
+The pitch is changed as a linear function of the elevation angle and the gradient is determined by setting the angle and pitch limits.
+For this work we only consider a \SI{180}{\degree} field of view in front of the user and limit the elevation angle to a range of $[-\frac{\pi}{2}, \frac{\pi}{2}]$.
+The pitch limits are set at some integer number of octaves above and below the neutral, on-elevation pitch.
+After practical tests with the interface, we set the neutral pitch to \SI{512}{\hertz}, which is comfortably audible and allows for a large number of audible octave limits to be selected.
+For this work, we set the pitch limits to 2 octaves away from the neutral pitch, giving frequency limits of [\SI{128}{\hertz}, \SI{2048}{\hertz}].
+The linear function is visualised in \cref{fig:pitch-gain}.
 

### 2019-07-11 paper.tex
Added experiment description, system implementation

@@
 	%For this, we conducted a set of experiments with both blindfolded participants and participants with visual impairments and found that, with some limitations, a spatialised audio tone with a varying pitch component can successfully convey a target's pan and tilt angles.
+  The ActiVis project is attempting to build a mobile guidance aid to help people with limited vision search for objects within an unknown environment.
+  This system requires an effective non-visual interface to transmit guidance instructions to the user through means of bone-conducting headphones.
+  To this end we propose a audio-based interface that uses a spatialised signal to convey a target's position on the horizontal plane. 
+  The position on the median plan is given by adjusting the tone's pitch to overcome audio localisation issues that arise with bone-conducting headphones. 
+  This interface is validated through a set of experiments. 
 \keywords{Human-machine interface \and vision impairment \and spatialised sound \and varying pitch, pointing task}
@@
 
-\begin{figure}
+\begin{figure}[t]
   \centering
@@
 
-%Our ultimate goal is to deliver a navigation system that can guide a user with visual impairments on the last leg of their journey (e.g.\ to a specific aisle and shelf in a shop) using only a mobile phone.
-%A large amount of data need to be translated from a visual form into a format that is useful to people with limited or no vision.
-%We therefore plan to use a combination of voice, audio and vibration cues to translate visual navigation data as effectively as possible and overcome the bandwidth limitation of the human ear. 
-%However, for this paper, we only considered the spatialised audio mode and its variation in pitch in order to determine its effectiveness in conveying a target's pan and tilt angles.
-
-%The system is based on a concept proposed in~\cite{bellotto2013,lock2017portable}, which uses a Google Tango device.
-%The navigation system is based on a Google Tango tablet, an Android-based device that comes equipped with an RGB-D camera to sense colour and depth and combines an inertial measurement unit (IMU) with powerful and robust landmark recognition and image processing algorithms to localise itself in real-time.
-%An added benefit of this platform is its familiar, compact form-factor which will help overcome the hurdle of user-acceptance and usability.
-%The guidance instructions are presented through a set of bone-conducting headphones (see \cref{fig:tango-headphone}) that are placed externally on a user's head so that the system does not occlude the perception of real-world sounds and does not interfere with a user's normal hearing function.
-
-%A diagram of the experimental system pipeline is shown in \cref{fig:pipeline}, where the arrows indicate the direction of the flow of information.
-%When the user taps the Tango's screen, a new virtual target is generated and its coordinates are sent to the audio generation module, along with the device's current position and orientation.
-%The audio generator then produces a tone based on the difference between the device and target's positions and sends it to the audio output channel, which plays it back to the user.
-%The WiFi recording module is constantly monitoring the different parameter values of the device and target's positions, as well as the system's output, and records it to a remotely stored datafile.
-
-%\begin{figure}
-  %\centering
-  %\includegraphics[clip=true, trim=0 120 80 50, width=1.0\columnwidth]{figures/pipeline.pdf}
-  %\caption{A diagram of the individual system components and their communication pipelines. $F$ indicates a feedback signal and $P$ a pose signal. }\label{fig:pipeline}
-%\end{figure}
-
 \subsection{Human Audio Localisation}
@@
 For this work, we propose a simple linear adjustment to the signal's pitch as a function of the elevation angle. 
-The pan angle can be conveyed by transforming the audio signal with an HRTF, and indeed it has been found that this dimension is unaffected by using bone-conducting headphones\cite{schonstein2008comparison,macdonald2006spatial,stanley2006lateralization}.\todo{cite bone cnducting work}
+The pan angle can be conveyed by transforming the audio signal with an HRTF, and indeed it has been found that this dimension is unaffected by using bone-conducting headphones\cite{schonstein2008comparison,macdonald2006spatial,stanley2006lateralization}.
 
@@
 
+A diagram of the experimental system pipeline is shown in \cref{fig:pipeline}, where the arrows indicate the direction of the flow of information.
+When the user taps the Tango's screen, a new virtual target is generated and its coordinates are sent to the audio generation module, along with the device's current position and orientation.
+The audio generator then produces a tone based on the difference between the device and target's positions and sends it to the audio output channel, which plays it back to the user.
+The WiFi recording module is constantly monitoring the different parameter values of the device and target's positions, as well as the system's output, and records it to a remotely stored datafile.
+
+\begin{figure}
+  \centering
+  \includegraphics[clip=true, trim=0 120 80 50, width=0.8\columnwidth]{figures/pipeline.pdf}
+  \caption{A diagram of the individual system components and their communication pipelines. $F$ indicates a feedback signal and $P$ a pose signal. }\label{fig:pipeline}
+\end{figure}
+
 \section{Experiment and Results}
 
-\subsection{Apparatus}
+To test the interface's effectiveness at directing a user towards a target area, a set of experiments were conducted with a simple pointing task. 
+This pointing task is set up to capture the difference between the targets' actual position and the positions the participants' perceived them to be.
 
-\subsection{Procedure}
+For the experiments, the participants are given a Tango tablet device running an app specifically written for these experiments. 
+This app generates a set of virtual targets and presents them to the participants one at a time. 
+The targets are generated at a constant \SI{2}{\meter} from the participant and their the pan and elevation angles are uniformly generated across 4 quadrants to avoid clustering.
+Each target's angular position is communicated to the participant through the audio interface, the output of which is adjusted in real-time as the participant points the device around. 
+When a participant was confident that they were on-target, i.e.\ hearing the audio front-on at \SI{512}{\hertz}, they tapped the screen, marking the location and generating the next target.
+The targets' positions are all set relative to the Tango device's coordinate system which is tracked using the Tango localisation API.\
+A total of 28 targets were generated per participant.
+
+2 groups of participants were recruited for the experiments. 
+Group \textit{G1} consisted of 10 young adults with normal eyesight that were blindfolded for the experiments, and group \textit{G2} contains 2 people with severe vision impairments. 
+Both groups were given some time with the system before the experiments commenced to familiarise themselves with the system, the audio signal's behaviour and the \SI{512}{\hertz} on-level tone. 
+To minimise the speed/accuracy biases in the results, we asked the participants to focus on finding the targets as well as they could without worrying about the time it took.
 
@@
 %However, to our knowledge no extensive work or experiments have been done to determine how well users respond to tilt adjustment instructions using a tone with \textit{varying pitch}, in particular when using bone-conducting headphones, and how changes to this tone's behaviour affects performance.
-
-%\section{Portable Navigation System}\label{sec:portable-navigation}
-
-%\subsection{System Setup}
-
-%Our ultimate goal is to deliver a navigation system that can guide a user with visual impairments on the last leg of their journey (e.g.\ to a specific aisle and shelf in a shop) using only a mobile phone.
-%A large amount of data need to be translated from a visual form into a format that is useful to people with limited or no vision.
-%We therefore plan to use a combination of voice, audio and vibration cues to translate visual navigation data as effectively as possible and overcome the bandwidth limitation of the human ear. 
-%However, for this paper, we only considered the spatialised audio mode and its variation in pitch in order to determine its effectiveness in conveying a target's pan and tilt angles.
-
-%The system is based on a concept proposed in~\cite{bellotto2013,lock2017portable}, which uses a Google Tango device.
-%The navigation system is based on a Google Tango tablet, an Android-based device that comes equipped with an RGB-D camera to sense colour and depth and combines an inertial measurement unit (IMU) with powerful and robust landmark recognition and image processing algorithms to localise itself in real-time.
-%An added benefit of this platform is its familiar, compact form-factor which will help overcome the hurdle of user-acceptance and usability.
-%The guidance instructions are presented through a set of bone-conducting headphones (see \cref{fig:tango-headphone}) that are placed externally on a user's head so that the system does not occlude the perception of real-world sounds and does not interfere with a user's normal hearing function.
-
-%A diagram of the experimental system pipeline is shown in \cref{fig:pipeline}, where the arrows indicate the direction of the flow of information.
-%When the user taps the Tango's screen, a new virtual target is generated and its coordinates are sent to the audio generation module, along with the device's current position and orientation.
-%The audio generator then produces a tone based on the difference between the device and target's positions and sends it to the audio output channel, which plays it back to the user.
-%The WiFi recording module is constantly monitoring the different parameter values of the device and target's positions, as well as the system's output, and records it to a remotely stored datafile.
-
-%\begin{figure}
-  %\centering
-  %\includegraphics[clip=true, trim=0 120 80 50, width=1.0\columnwidth]{figures/pipeline.pdf}
-  %\caption{A diagram of the individual system components and their communication pipelines. $F$ indicates a feedback signal and $P$ a pose signal. }\label{fig:pipeline}
-%\end{figure}
-
-%\subsection{Audio Interface}
-
-%The audio component is responsible for conveying the 2D position of a target on the vertical plane, shown by \cref{fig:cam-coords}, in terms of pan and tilt angles.
-%The audio signal is a continuous \SI{10}{\hertz} sinusoidal sound wave that is played to the user through bone-conducting headphones.
-%We use a sinusoidal wave here because it is relatively simple to manipulate and analyse and will be replaced by a richer, more natural tone at a later stage.
-
-%The audio is spatialised in the pan dimension using the OpenAL\footnote{\url{http://openal.org}} sound library's HRTF, while the tilt angle is conveyed by varying the audio tone's pitch.
-%We use this approach because the external set of bone-conducting headphones plays the sound through a user's cheekbones instead of their outer ears, bypassing the ears' pennae that provide the ability to localise elevated sound sources~\cite{roffler1968factors,algazi2001elevation}.
-%This makes an HRTF less effective in conveying the target's elevation and therefore needs to be conveyed using another method; by varying the tone's pitch in this case.
-%The difference between the target's angular position and the device's angular orientation are used to generate the audio signals.
-
-%\subsubsection{Pan Direction}
-
-%The pan angle ($\phi$) describes the angle which the user needs to rotate the camera vector, $C$, around the $z$-axis shown in \cref{fig:cam-coords}, i.e.\ how far the target is to the left or right of the user.
-%We use an HRTF to add a spatial element to the audio tone that is played to the user, placing the sound source at the target's location. 
-%Humans determine a sound source's pan direction by subconsciously analysing the inter-aural time difference (IID) between a sound reaching both of your ears: the greater the IID, the greater the perceived angular distance to the sound source~\cite{wightman1992dominant}.
-
-%We implement OpenAL's default HRTF, based on the MIT's KEMAR dataset\footnote{\url{http://sound.media.mit.edu/resources/KEMAR.html}}, to generate a sinusoidal sound wave based on the difference between the user and target's positions.
-%The library takes position values as input and outputs a tone based on the angle between the two position vectors.
-
-%\subsubsection{Tilt Direction}
-
-%The system adjusts the tone's pitch ($f$) to convey the target's tilt angle ($\theta$) w.r.t.\ $C$. 
-%Here, a high pitch means the target is above $C$ (i.e.\ the user should look up) and a low pitch means the target is below $C$ (the user should look down).
-%This high/low association scheme is chosen because humans naturally associate high-pitched sounds with higher objects and lower-pitched noises with lower objects~\cite{pratt1930spatial,blauert1997spatial}.
-%We use a logarithmic, octave-based gain function for the pitch, since an increase in octave provides a distinct perceptible change while keeping the timbre roughly constant~\cite{shepard1964circularity}.
-
-%We wish to determine how the gradient, $m$, of the pitch gain as a function of $\theta$ (i.e.\ $m = \frac{df}{d\theta}$) affects a user's performance, e.g.\ does a greater $m$ lead to an increased target acquisition rate?
-%For this we select 3 different pitch settings, referred to henceforth as $m_l$, $m_m$ and $m_h$ for the smallest, middle and highest gain gradients respectively. 
-%To find these gradients, we set the maximum and minimum limits for $\theta$ and the maximum and minimum frequencies for $f$ for each respective setting.
-%Furthermore, for the sake of consistency and between the different settings, each gradient generates the same pitch when $C$ is level with the target's tilt.
-
-%After practical tests with the device and headphones, we set the neutral, on-target tone to $f_{512} =$ \SI{512}{\hertz} for its audibility.
-%For $m_l$, we set $f_{max}$ and $f_{min}$ to 1 octave higher and lower than $f_{512}$, giving limits of \SI{1024}{\hertz} and \SI{256}{\hertz} respectively.
-%For $m_m$, the $f_{max}$ and $f_{min}$ are set to 2 octaves higher and lower than $f_{512}$ (\SI{2048}{\hertz} and \SI{128}{\hertz}) and $m_h$ to 3 octaves higher and lower (\SI{4096}{\hertz} and \SI{64}{\hertz}).
-%These limits were selected for practical reasons, because the bone conducting headphones have low volume gain at the frequency spectrum extremes, making it difficult to hear. 
-%The values of $m_l, m_m, m_h$ are visualised in \cref{fig:pitch-gain}.
-
-%\begin{figure}
-  %\centering
-  %\includegraphics[clip, trim=0 0 0 40, width=1.0\columnwidth]{figures/pitch_gain_functions.png}
-  %\caption{The pitch gain functions for the 3 different gain settings ($y$-axis in log scale). }\label{fig:pitch-gain}
-%\end{figure}
-
-%We set the range of angles the system can convey to $[-\frac{\pi}{2}; \frac{\pi}{2}]$; anything outside this range implies that the target is behind the user and was not considered in this series of experiments.
-%The \SI{0}{\radian} position is directly in front of the user.
-%With these limits, $f$ can be determined with
-
-%\begin{equation}\label{eq:frequency}
-  %f(\theta) = 2^{b\theta + c},
-%\end{equation}
-%where $b$ and $c$ for each setting can be calculated according to the given setting's pitch and angle limits.% and a linear graph equation (i.e.\ $f - f_1 = b(\theta - \theta_1), b=\frac{\Delta f}{\Delta\theta}$).
-
-%\section{Target Search Experiment}\label{sec:experiments}
-
-%We conducted a set of experiments to evaluate the audio interface's effectiveness at driving a user to point the camera at a target in the pan and tilt dimensions.
-%Furthermore, we wish to see how changes to $m$ affect a user's performance.
-
-%We recruited 2 groups of participants for the experiments.
-%Group $G1$ consists of 42 young adults (mostly students) with healthy eyesight, with 10 males and 32 females being $20\pm1.5$ years old on average.
-%This group was blindfolded throughout the experiment and forms the bulk of the participant population.
-%Group $G2$ is made up of 3 legally blind adults with little to no light perception (3 males, average age of $40\pm8$ years). 
-%This group makes up less than 10\% of the total experiment population, but provided valuable subjective feedback that will help improve the system before evaluating it with a larger group of people with vision impairments.
-%All participants joined the experiment on a voluntary basis with written consent and have no other impairments that affected their interpretation of the audio guidance cues.
-
-%For the experiment, the participants were given a Tango tablet device, running an app written for this experiment, and a pair of AfterShockz Sportz3' bone conducting headphones (both pictured in use in \cref{fig:participant}) and were blindfolded where appropriate.
-%The app generates a virtual target, one at a time at a constant \SI{2}{\metre} from the participant, and uses the device's pose measurements to determine $\phi$ and $\theta$.
-%These angles are then used to generate and present the user with the audio signal to direct them to the target. 
-%As the participant rotates the camera, $\phi$ and $\theta$ are adjusted and the audio signal changes to reflect these changes.
-%When a participant believes that they are pointing at the target, they tap the screen which generates a new target target which they are then directed to. 
-%Each participant was given 28 targets in total that were randomly generated and equally spread across 4 quadrants to avoid clustering. 
-%After the 28 targets were found, $m$ was adjusted and the experiment was repeated. 
-%The order in which the gradient setting was changed for each participant was randomised to minimise the learning effect.
-%The app streams various data on the current device and target poses to a laptop computer via a WiFi connection in real-time. 
-%Before the experiment session started, each participant was given some time to familiarise themselves with the device and was given the opportunity to learn what the $f_{512}$ `on-target' and on-centre tone sounds like that will mean that they are pointing at the target.
-
-%The measure of performance here is the accuracy with which a user points the camera towards a target.
-%The accuracy is taken as the angular difference between $C$ and the target's position at the time the participant tapped the screen.
-%A smaller error indicates more accurate results.
-%The pan and tilt results are separated to understand the effect of changing the gradient setting has on either dimension's accuracy.
-%To minimise the speed/accuracy biases in the results, we asked the participants to focus on finding the targets as well as they could without worrying about the time it took.
 

### 2019-07-12 paper.tex
Added new figures, added results discussion,  added conclusion

@@
   \subfloat[]{\label{fig:participant}\includegraphics[width=0.25\columnwidth]{figures/vi_participant.pdf}}
-  \caption{Pictures of the Tango device and bone-conducting headphones the system is based on (left) and both in use in an experiment (right).}\label{fig:tango-headphone}
+  \caption{Pictures of the Tango device and bone-conducting headphones the system is based on (left) and both in use in an experiment (right).}
 \end{figure}
@@
 Previous work on HRTF stuff
+
+\todo{add previous work}
 
@@
 ~
-  \subfloat[]{\label{fig:pitch-gain}\includegraphics[clip, trim=0 0 0 40, width=0.6\columnwidth]{figures/pitch_gain_functions.png}}
+  \subfloat[]{\label{fig:pitch-gain}\includegraphics[clip, trim=0 0 0 40, width=0.6\columnwidth]{figures/pitch_gain_function.png}}
   \caption{The reference system used by the guidance interface showing the camera vector and pan and elevation angles (left) and the pitch gain function used to convey the target's elevation angle (right).
@@
 2 groups of participants were recruited for the experiments. 
-Group \textit{G1} consisted of 10 young adults with normal eyesight that were blindfolded for the experiments, and group \textit{G2} contains 2 people with severe vision impairments. 
+Group \textit{G1} consisted of 10 young adults with normal eyesight that were blindfolded for the experiments, and group \textit{G2} contains 4 people with severe vision impairments. 
 Both groups were given some time with the system before the experiments commenced to familiarise themselves with the system, the audio signal's behaviour and the \SI{512}{\hertz} on-level tone. 
@@
 
+The data collected during the experiment were grouped together by participant group and analysed.
+Similar to the results from~\cite{macdonald2006spatial,schonstein2008comparison}, we use the absolute mean errors for each dimension which allows us to compare our results to the aforementioned works'.
+These results are presented in the boxplots in \crefrange{fig:sighted-err}{fig:blind-err} and summarised in \cref{tab:results}.
+
+\todo{make plot text larger}
+\todo{should we include blind results?}
+
+\begin{figure}[t]
+  \centering
+  \subfloat[]{\label{fig:sighted-err}\includegraphics[width=0.5\columnwidth]{figures/pan_tilt_err_sight.png}}
+~
+  \subfloat[]{\label{fig:blind-err}\includegraphics[width=0.5\columnwidth]{figures/pan_tilt_err_blind.png}}
+  \caption{Boxplots of the angular errors collected during the pointing experiment for the blindfolded (left) and visually impaired (right) participants. }
+\end{figure}
+
+\begin{table}
+  \centering
+  \caption{}\label{tab:results}
+  \begin{tabular}{p{1cm}cc}
+    \toprule
+	                         &           & Mean absolute error [rad] \\ \midrule
+    \multirow{2}{*}{\textit{G1}} & Pan       & $ 0.25\pm0.24$ \\
+				 & Elevation & $ 0.37\pm0.27$ \\ \midrule
+    \multirow{2}{*}{\textit{G2}} & Pan       & $ 0.48\pm0.33$ \\
+				 & Elevation & $ 0.58\pm0.42$ \\
+    \bottomrule
+  \end{tabular}
+\end{table}
+
+\todo{add correlation plots}
+
+The average pan error from group \textit{G1} falls well within the ranges observed in~\cite{macdonald2006spatial,schonstein2008comparison} of $[0.16, 0.38]$ radians and $[0.17, 0.26]$ radians respectively.
+Indeed, the majority of samples fall within even the more conservative latter range. 
+However, the participants with vision impairments in group \textit{G2}  demonstrate a wider spread in error data and higher average error than \textit{G1} (0.23 vs.
+0.40 radians, Kruskal-Wallis test $p=0.05$).
+\todo{do KW test}
+The results in~\cite{katz2011spatial,zwiers2001spatial} reported a similar trend, but an investigation with a larger sample size for \textit{G2} should be considered before definitive conclusions can be made.
+
+The elevation estimation performance for both groups deteriorated when compared to their pan estimation results, which was somewhat expected given previous research results on human audition.
+\todo{cite vertical vs horizontal performance}
+The mean absolute errors of the participants are \SI{0.37}{\radian} and \SI{0.58}{\radian} for groups \textit{G1} and \textit{G2} respectively.
+Comparing these results to methods that convey elevation through an HRTF, we see an improvement of approximately 46\% and 16\% for each group~\cite{schonstein2008comparison}. 
+There is a lesser improvement (13\%) with \textit{G1} when compared to the individualised and adjusted HRTFs in~\cite{stanley2006lateralization} and a 13\% reduction in performance for group \textit{G2}.
+However, the participants from~\cite{stanley2006lateralization} all had healthy eyesight, making the latter comparison invalid. 
+Interestingly, the elevation errors for group \textit{G2} is centered around the negative angle, indicating a possible underlying bias in the participants or the interface.
+A similar trend was observed in~\cite{stanley2006lateralization} and warrants an investigation with a larger participant pool to determine the underlying bias, if any.
+
 \section{Conclusion}
+
+In this work, we proposed an audio interface that conveys a target's pan and elevation angles through a spatialised audio tone that adjusts the tone's pitch to convey the elevation in compensate for the shortcomings of bone-conducting headsets in transmitting the elevation component of a spatialised signal.
+Results generated from a simple pointing experiment are encouraging and improves performance over purely spatialised signals transmitted via bone-conduction.
+Future work will involve a more in-depth investigation with a larger participant pool to determine the factors that affect performance and enable a more detailed statistical analysis.
+For example, will adjusting the elevation function's gradient make it easier to guess the target's elevation and what is the performance difference between participants with limited and healthy eyesight?
 
@@
 
-%\section{Results and Discussion}\label{sec:results}
-
-%The results for the participants are given by the 2D histograms in \cref{fig:err-results}, where the pan error is on the abscissa and the tilt error on the ordinates.
-%A set of box-plots of the angular errors' median and standard deviations are given in \cref{fig:err-boxplot} for the $m_l, m_m$ and $m_h$ settings. 
-%The results for $G1$ are summarised in \cref{tab:results}, while the results for $G2$ are given in \cref{tab:vi-results}.
-
-%\begin{figure}
-  %\centering
-  %\subfloat[Results for the $m_l$ setting.]{\label{fig:err-results-lo}\includegraphics[clip, trim=450 0 450 110, width=0.8\columnwidth]{figures/err_lo.png}}
-
-  %\subfloat[Results for the $m_m$ setting.]{\label{fig:err-results-med}\includegraphics[clip, trim=450 0 450 110, width=0.8\columnwidth]{figures/err_med.png}}
-
-  %\subfloat[Results for the $m_h$ setting.]{\label{fig:err-results-hi}\includegraphics[clip, trim=450 0 450 110, width=0.8\columnwidth]{figures/err_hi.png}}
-
-  %\caption{Angular error data for the 3 gradient settings. }\label{fig:err-results}
-%\end{figure}
-
-%\begin{figure}
-  %\centering
-  %\subfloat[The medians. ]{\label{fig:err-boxplot-median}\includegraphics[clip, trim=20 -70 100 100, width=1.0\columnwidth]{figures/err_boxplot_medians.png}}
-
-  %\subfloat[The standard deviations. ]{\label{fig:err-boxplot-std}\includegraphics[clip, trim=20 -70 100 100, width=1.0\columnwidth]{figures/err_boxplot_std.png}}
-
-  %\caption{Boxplots for the error medians and standard deviations. }\label{fig:err-boxplot}
-%\end{figure}
-
-%\begin{table}
-  %\centering
-  %\caption{A table of the pan and tilt angular error results collected from $G1$.
-The results include the absolute average error ($\mu_{abs}$), average error ($\mu$) and standard deviations ($\sigma$) given in radians, as well as the Spearman ($R$) correlation scores.}\label{tab:results}
-  %\begin{tabular}{llcccc}
-    %\toprule
-    %\multicolumn{2}{c}{} & $\mu_{abs}$ & $\mu$ & $\sigma$ & $R$ \\\midrule
-	 %& $m_l$ & 0.23 & -0.02 & 0.32 & 0.75 \\%\cmidrule{2-6}
-    %Pan  & $m_m$ & 0.21 & -0.01 & 0.28 & 0.77 \\%\cmidrule{2-6}
-	 %& $m_h$ & 0.22 & -0.05 & 0.31 & 0.75 \\\midrule
-	 %& $m_l$ & 0.40 & -0.14 & 0.46 & 0.38 \\%\cmidrule{2-6}
-    %Tilt & $m_m$ & 0.32 & -0.12 & 0.36 & 0.48 \\%\cmidrule{2-6}
-	 %& $m_h$ & 0.34 & -0.16 & 0.39 & 0.55 \\
-    %\bottomrule
-  %\end{tabular}
-%\end{table}
-
-%\begin{table}
-  %\centering
-  %\caption{A table summarising the target search results of the participants in $G2$ (P1 - P3).
-The results include the average error ($\mu$) and standard deviation ($\sigma$), given in radians in the pan and tilt dimensions.}\label{tab:vi-results}
-  %\begin{tabular}{llcccccc}
-    %\toprule
-    %\multicolumn{2}{c}{} & \multicolumn{2}{c}{$m_l$} & \multicolumn{2}{c}{$m_m$} & \multicolumn{2}{c}{$m_h$} \\
-    %\multicolumn{2}{c}{} & $\mu_l$ & $\sigma_l$ & $\mu_m$ & $\sigma_m$ & $\mu_h$ & $\sigma_h$ \\\midrule
-	 %& P1 &  0.25 & 0.44 &  0.21 & 0.45 &  0.29 & 0.40 \\% \cline{2-8}
-    %Pan  & P2 &  0.18 & 0.85 &  0.05 & 0.49 & -0.08 & 0.32 \\% \cline{2-8}
-	 %& P3 & -0.10 & 0.67 & -0.08 & 0.61 & -0.08 & 0.50 \\ \midrule
-	 %& P1 & -0.17 & 0.21 & -0.10 & 0.2  & -0.25 & 0.21 \\% \cline{2-8}
-    %Tilt & P2 & -0.87 & 0.26 & -0.05 & 0.21 & -0.15 & 0.18 \\% \cline{2-8}
-	 %& P3 & -0.57 & 0.31 & -0.57 & 0.45 & -0.38 & 0.39 \\% \hline
-    %\bottomrule
-  %\end{tabular}
-%\end{table}
-
-%\subsection{Pan Results}
-
-%\cref{fig:err-results} and the median and mean points from the boxplots in \cref{fig:err-boxplot} show that the data in the pan dimension are approximately normally distributed around zero.
-%However, a Shapiro-Wilkes test score below the critical threshold do not give enough confidence to consider the data as normally distributed and we therefore analyse the data through non-parametric techniques.
-
-%For the data collected from $G1$, there is a strong linear correlation between the participants' guesses and the targets' true pan angles, shown by the high Spearman correlation scores of approximately 0.75 ($p < 0.01$) for all of the datasets, with the $m_m$ setting displaying the marginally best result of 0.77.
-%The 3 settings have similar average errors and standard deviations, with the $m_m$ setting again producing the marginally best results with the smallest average error and standard deviation.
-%However, the differences are relatively small (approximately 6\%) and are not significant enough (Friedman test $p = 0.92$) to conclude that the pitch gradient has an effect on the target acquisition accuracy performance in the pan dimension.
-
-%The data collected from $G2$ follow a similar trend to $G1$: the error in the pan dimension is centred around \SI{0}{\radian} with similar standard deviations across all 3 settings.
-%Unfortunately, not enough samples could be collected to draw any final conclusions.
-%However, these results are in line with results from previous work~\cite{zwiers2001spatial}, indicating that the groups with partial and normal eyesight respond with similar levels of accuracy to spatialised sound when presented with simple tasks.
-
-%These results show that both groups managed to find the targets with reasonable level of accuracy and that the pan error is robust to variations in the pitch's rate of change; a useful result in support of the effectiveness of this audio interface.
-
-%\subsection{Tilt Results}\label{sec:tilt-results}
-
-%The data within the tilt dimension were also found to be non-normal following scores below the critical threshold from the Shapiro-Wilkes test.
-%The Spearman test for the data reveals a reasonably significant correlation between the participants' guesses and the actual locations of the targets, with correlation scores of approximately 0.38 for $m_l$, 0.48 for $m_m$ and the $m_h$ gradient giving the strongest correlation score with 0.55 ($p < 0.01$ for all 3 settings).
-%This indicates that the participants generally interpret the cues from the varying the pitch correctly and are mostly successful at pointing the camera in the correct tilt angle.
-
-%The average errors of the datasets are relatively close to one another, with the $m_l$ setting producing the largest absolute error and standard deviation at \SI{0.40}{\radian} and \SI{0.46}{\radian} respectively.
-%The $m_m$ and $m_h$ settings have similar absolute errors of \SI{0.32}{\radian} and \SI{0.34}{\radian} respectively.
-%This is in line with the correlation scores, with $m_l$ giving the worst result and $m_h$ giving the best results overall, with highest correlation score and relatively low absolute angular errors and standard deviation, marginally beating $m_m$.
-
-%Since the data are not normally spread and there is significant levels of noise within the data which may contaminate the mean values, we use the medians to analyse the datasets.
-%For this, the Friedman test is used on the participants' error medians and it was found that there is a significant difference between the 3 settings' datasets ($p < 0.01$).
-%A post-hoc analysis with a Wilcoxon signed rank test and a critical threshold with a Bonferroni correction applied ($\alpha=0.016=\frac{0.05}{3}$) shows that both the $m_m$ and $m_h$ settings are significantly different from the $m_l$ setting ($p=0.0014<\alpha$ for both settings), but that the $m_m$ and $m_h$ settings do not produce significantly different results from one another ($p=0.42>\alpha$).
-%This indicates there is an inflection point where increasing $m$ produces diminishing returns and that we cannot conclude that the $m_h$ setting is better than the $m_m$ gradient.
-%However, both settings conclusively produce more accurate results than the $m_l$ setting.
-
-%From \cref{fig:err-boxplot-median} it can be seen that the median data show a significant skewing to the negative side, indicating a potential bias amongst the experiment's participant-base that must be taken into account.
-%From the results, it is not completely clear what causes this negative bias in the pitch dimension.
-%However, a possible explanation is that the floor introduces a position constraint within a user's mind, since the target cannot appear below the ground.
-%It can potentially appear above the user's head though, giving variable upper an lower limits that depend on the user's height and individual perception.
-%Future improvements to our audio interface should consider a non-linear increase in pitch as a function of tilt angle instead of the linear one we used in this work to remedy this bias.
-
-%The results collected from $G1$ and $G2$ show that a varying pitch tone can convey the tilt angle of a target to a human using bone-conducting headphones with accuracy levels similar to that in literature~\cite{bujacz2011sonification,katz2011spatial,zotkin2004rendering}.
-%They also highlight a clear and significant difference between the 3 different pitch gain gradients, with the $m_l$ setting giving the lowest acquisition accuracy, while $m_m$ and $m_h$ give similar accuracy levels closest to the target's true tilt.
-
-%\subsection{Participants with Visual Impairment's Feedback}
-
-%After the experiments with the participants with visual impairments in $G2$ were completed, we took the opportunity to collect their feedback on the audio interface's current implementation and gather suggestions for improvements for future iterations.
-%When asked about their opinions on current assistive systems available on the market, the general consensus was that these systems are typically prohibitively expensive, too complex and cumbersome to be useful for everyday use.
-%Indeed, they seemed quite enthusiastic about a guidance system that they can use on a mobile phone.
-
-%They were generally satisfied with the guidance interface in its current form and had little problem learning how to use it.
-%However, they, along with the participants in $G1$, reported auditory fatigue caused by long-term exposure to the sinusoidal audio tone.
-%We were aware of this issue before the experiments were conducted and have considered replacing the simple sine wave tone with more natural and pleasant sounds.
-%For example, musical instruments can be used to play different notes and scales to indicate different elevation angles~\cite{brewster1998using}.
-%Furthermore, they suggested adding some voice component or tone prompt to confirm when they are on target, inform them of other important events or happenings or simply to intermittently assure them that the system is still active.
-%The former two suggestions were intentionally ignored in order to purely evaluate how accurately the audio tone can guide a user to the target.
-%Going forward, however, these suggestions will be helpful for a potential commercial release version of the guidance system to make the guidance instructions clearer and to intermittently reassure users that they are still on the right course, for example.
-
-%On the hardware, only one participant had reservations regarding the choice of bone-conducting headphones, citing cost concerns. 
-%However, he indicated that he could justify purchasing a set of bone-conducting headphones if he were to use them throughout the day to avoid cutting off environmental sounds.
-%None of the participants had any technical issues using the headphones or mobile device and were able to properly use and interpret the guidance instructions.
-
 %\section{Conclusion and Future Work}\label{sec:conclusion}

### 2019-07-12 paper.tex
Added extra conclusion line, another result comparrisson

@@
 
+\todo{review title to make focus more on bone conduction and its viability}
 In recent years, governments have spearheaded numerous initiatives to support people with disabilities and enable them to play a more active role in modern society.
@@
 
-The rest of the paper is \todo{lay out rest of paper}
+The rest of the paper starts off by discussing previous relevant works and research in \cref{sec:prev-work}, followed by a discussion on the design and implementation of our interface in \cref{sec:interface-design}.
+This is followed by a description of the experiments that were conducted and a discussion of their results in \cref{sec:experiment-and-results}.
+Finally, the work is concluded with a summary and a short discussion on future research prospects in \cref{sec:conclusion}.
 
-\section{Previous Work}
+\section{Previous Work}\label{sec:prev-work}
 
@@
 
-\section{Interface Description}
+\section{Interface Description}\label{sec:interface-design}
 
@@
 
-\section{Experiment and Results}
+\section{Experiment and Results}\label{sec:experiment-and-results}
 
@@
 The mean absolute errors of the participants are \SI{0.37}{\radian} and \SI{0.58}{\radian} for groups \textit{G1} and \textit{G2} respectively.
+As with the pan estimation errors, the participants from \textit{G1} seemed more adept at estimating the correct elevation.
+However, the difference is not as significant as for the pan, with \textit{G1} displaying a 36\% improvement over \textit{G1} in their elevation estimation vs. 42.5\% in the pan dimension. 
 Comparing these results to methods that convey elevation through an HRTF, we see an improvement of approximately 46\% and 16\% for each group~\cite{schonstein2008comparison}. 
+Indeed, the performance is comparable to that of open-back and expensive in-ear headphones.
 There is a lesser improvement (13\%) with \textit{G1} when compared to the individualised and adjusted HRTFs in~\cite{stanley2006lateralization} and a 13\% reduction in performance for group \textit{G2}.
@@
 
-\section{Conclusion}
+\section{Conclusion}\label{sec:conclusion}
 
@@
 For example, will adjusting the elevation function's gradient make it easier to guess the target's elevation and what is the performance difference between participants with limited and healthy eyesight?
+Furthermore, experiments with a large group of people with vision impairments will provide an opportunity to gauge their opinion on the interface and identify possible avenues of improvements to user friendliness.
 
@@
 
-%\section{Conclusion and Future Work}\label{sec:conclusion}
-
-%In this work, we present an audio-based guidance interface that uses a spatialised audio tone with varying pitch to direct a user with visual impairments to point a camera toward a target. 
-%We also present and discuss a set of experiments that were conducted to determine the interface's effectiveness and performance, as well as some subjective feedback that was collected from legally blind participants.
-
-%Using a set of bone-conducting headphones, we found that a spatialised audio tone with a varying pitch can successfully convey the pan and tilt angles of a target.
-%The angular errors made by the participants are in line with those from previous research using similar audio guidance instructions.
-%We also found that varying the pitch's rate of change influences the accuracy of the system in the tilt dimension without affecting the performance in the pan dimension.
-%The steeper, $m_h$ and $m_m$ pitch gradients were found to produce the best accuracy results in general.
-
-%Feedback from the participants with visual impairments was mostly positive and they agreed with the choice of bone-conduction headphones that do not inhibit normal hearing.
-%However, they noted that the sine wave audio signal induced audio fatigue after extended exposure, something noted by the other participants as well, and suggested that more full-bodied, natural sounds be used instead.
-%Using vocal feedback in conjunction with the current audio signal was also suggested.
-
-%These results are more interesting from a design point of view than an end-user's. 
-%Indeed, we have shown that this audio guidance interface is effective at directing an average person, but more data are needed with the intended audience in order to decisively conclude that this interface will be effective in an indoor guidance system for people with vision impairments.
-%Nevertheless, these data are a good indication of what to expect when we evaluate this audio interface with a greater number of people with limited vision.
-%Once we have determined that the intended audience can effectively use this type of guidance instruction, we can start adding more user-oriented features (e.g.\ voice prompts, more pleasant tones, etc.) and integrate them into our guidance system prototype.
-
 \bibliographystyle{splncs04}

### 2019-07-12 paper.tex
Reading through. Need to add previous work

@@
 \begin{abstract}
-	%Our aim is to build a portable active navigation system for people with visual impairments that uses active perception techniques and a combination of feedback modes to identify a scene and guide a user to their destination.
-	%Such a system requires a non-visual feedback interface and in this paper we investigate the effectiveness of a spatial audio tone with a varying pitch component, played through bone-conducting headphones, in conveying the pan and tilt angles of a target to the user in a pointing task.
-	%The secondary goal is to determine how variations to the pitch's rate of change affect the user's performance.
-	%For this, we conducted a set of experiments with both blindfolded participants and participants with visual impairments and found that, with some limitations, a spatialised audio tone with a varying pitch component can successfully convey a target's pan and tilt angles.
   The ActiVis project is attempting to build a mobile guidance aid to help people with limited vision search for objects within an unknown environment.
@@
 Furthermore, a set of bone-conducting headphones (pictured in \cref{fig:tango-headphone}) that are placed on a user's cheekbones instead of their ears and do not interfere with normal hearing, are used to transmit the audio signals to the user.
-The system will provide real-time guidance instructions for the user to follow and it is important that these instructions are easily understood and are not perceived as a cognitive burden. 
-Humans are naturally able to determine the 3D position of a sound source and by exploiting this ability, the system's guidance instructions can naturally be interpreted without posing a unnecessary burdensome cognitive load.
-A sound source can be spatialised by adjusting a tone's spectral make-up (elevation angle), time delay and level difference (pan angle) and the intensity (distance).
-Here only the pan and elevation positions are transmitted.
-However, since bone-conducting headphones bypass the outer ear structure, their spectral signature cannot properly be interpreted and we therefore convey the target's elevation angle by simply adjusting the tone's pitch.
+The system provides the user with real-time guidance instructions and it is important that these instructions are easily understood and do not add a significant cognitive burden.
 
@@
 
+Humans are naturally able to determine the 3D position of a sound source and by exploiting this ability, our system's guidance instructions can naturally be interpreted without posing a significant cognitive load.
+A sound source can be spatialised by adjusting a tone's spectral make-up (elevation angle), time delay and level difference (pan angle) and the intensity (distance).
+Here only the pan and elevation positions are transmitted.
+However, since bone-conducting headphones bypass the outer ear structure, their spectral signature cannot properly be interpreted and we therefore convey the target's elevation angle by adjusting the tone's pitch.
 A similar approach was used in\cite{durette2008visuo}, but the study did not focus on or investigate the interface's efficacy.
@@
 Furthermore, we demonstrate our approach's effectiveness through a simple experiment.
-%The contribution of this work is an effective approach to addressing the shortcomings of playing spatialised sound through bone-conducting headphones, as well as initial experiments demonstrating its effectivenesst
-%Furthermore, we demonstrate that such an interface is effective at communicating
 
@@
 
-Electronic navigation aids and guidance systems for people with vision impairments have had trouble appealing to the visually impaired community at large because of problems that are typical of academic research and prototypes: prohibitive costs and user unfriendly interfaces and hardware requirements, e.g.\ head-mounted cameras and GPS antennae~\cite{golledge2004stated,yusif2016older,arditi2013user}.
+Electronic navigation aids and guidance systems for people with vision impairments have had trouble appealing to the visually impaired community at large because of problems that are common of academic research and prototypes: prohibitive costs, unfriendly user interfaces and significant hardware requirements, e.g.\ head-mounted cameras and GPS antennae~\cite{golledge2004stated,yusif2016older,arditi2013user}.
 Indeed, the traditional walking cane remains the go-to tool for navigation and obstacle avoidance. 
-With the ActiVis system, we hope to tackle the issue of cost, user-acceptance and usability by using as many as possible `standard' devices and hardware as possible, with the idea that no specialist hardware is required and the user will not stand out if seen walking in public with a phone in their hand and wearing a headset.
+With the ActiVis system, we aim to tackle the issue of cost, user-acceptance and usability by using as many as possible `standard' devices and hardware as possible, with the idea that no specialist hardware is required.
+Following this design target, we base the project on a mobile device and discreet headset with the anticipation that the user will not stand out if seen walking in public with a phone in their hand and wearing a headset.
 The ActiVis system is based on a concept proposed in~\cite{bellotto2013multimodal,lock2017portable}, which uses a Google Tango device that is an Android-based device that comes equipped with an RGB-D camera to sense colour and depth and combines an inertial measurement unit (IMU) with powerful and robust landmark recognition and image processing algorithms to localise itself in real-time.
@@
 
-Humans localise a sound source in 3 dimensions by considering cues recorded in one ear (monaural cues) and comparing cues derived at both ears (binaural cues)\cite{blauert1997spatial,blauert1969sound}.
+Humans localise a sound source in 3 dimensions by considering cues recorded in one ear (monaural cues) and comparing cues received at both ears (binaural cues)~\cite{blauert1997spatial,blauert1969sound}.
 The binaural cues include time differences of arrival and intensity differences that help to determine a source's location on the horizontal plane.
@@
 The user's anatomy acts as a filter and the direction the sound wave hits the body activates different filter resonances, attenuating or accentuating different frequency bands.
-When the modified sound enters the inner ear, the human brain is able to analyse the frequency response and accurately determine the position of the sound source on the vertical plane. 
+When the modified audio signal enters the inner ear canal, the human brain is able to analyse the frequency response and accurately determine the position of the sound source on the median plane. 
 The distance to the source is simply derived as the intensity, or volume, of the source, i.e.\ a louder sound would appear closer to the user than a softer one does.
 
-When an audio signal is transmitted via a set of speakers or headsets, it can be transformed to mimic the characteristics of a natural sound sources before it is transmitted, thereby tricking the brain into believing a sound is located at some arbitrary position.
+When an audio signal is transmitted via a set of speakers or headsets, it can be transformed to mimic the characteristics of a natural sound source before it is transmitted, tricking the brain into believing a sound is located at some arbitrary position.
 Such a transformation can be done using a head-related transfer function (HRTF) and a pair of stereo headsets or speakers.
 An HRTF is a mathematical function that simulates the response signal of a human head and is derived by capturing key characteristics that affect the monaural and binaural responses, such as the user's hearing levels and head size, etc.
-Since hearing responses can be quite unique between user's, the best results would be observed if each user had their own customised HRTF.\
-Given the complicated process involved to capture the required user characteristics, making unique HRTFs is often an untenable solution and using average values (e.g.\ head measurements, height, etc.) have shown to produce acceptable results.\todo{cite HRTF papers}
+Since hearing responses are unique amongst different users, the best results would be observed if each user had their own customised HRTF.\
+However, given the complicated process involved to capture the required user characteristics, making unique HRTFs is often an untenable solution and using average values (e.g.\ head measurements, height, etc.) have shown to produce acceptable results.\todo{cite HRTF papers}
 
@@
 Similar to the results from~\cite{macdonald2006spatial,schonstein2008comparison}, we use the absolute mean errors for each dimension which allows us to compare our results to the aforementioned works'.
-These results are presented in the boxplots in \crefrange{fig:sighted-err}{fig:blind-err} and summarised in \cref{tab:results}.
-
-\todo{make plot text larger}
-\todo{should we include blind results?}
+These results are presented in the boxplots in \cref{fig:results} and summarised in \cref{tab:results}.
 
@@
   \centering
-  \subfloat[]{\label{fig:sighted-err}\includegraphics[width=0.5\columnwidth]{figures/pan_tilt_err_sight.png}}
-~
-  \subfloat[]{\label{fig:blind-err}\includegraphics[width=0.5\columnwidth]{figures/pan_tilt_err_blind.png}}
-  \caption{Boxplots of the angular errors collected during the pointing experiment for the blindfolded (left) and visually impaired (right) participants. }
+  \includegraphics[width=0.7\columnwidth]{figures/pan_tilt_err.png}
+  \caption{Boxplots of the angular errors collected during the pointing experiment. }\label{fig:results}
+  %\subfloat[]{\label{fig:sighted-err}\includegraphics[width=0.5\columnwidth]{figures/pan_tilt_err_sight.png}}
+%~
+  %\subfloat[]{\label{fig:blind-err}\includegraphics[width=0.5\columnwidth]{figures/pan_tilt_err_blind.png}}
 \end{figure}
@@
     \multirow{2}{*}{\textit{G1}} & Pan       & $ 0.25\pm0.24$ \\
-				 & Elevation & $ 0.37\pm0.27$ \\ \midrule
-    \multirow{2}{*}{\textit{G2}} & Pan       & $ 0.48\pm0.33$ \\
-				 & Elevation & $ 0.58\pm0.42$ \\
+				 & Elevation & $ 0.37\pm0.28$ \\ \midrule
+    \multirow{2}{*}{\textit{G2}} & Pan       & $ 0.38\pm0.21$ \\
+				 & Elevation & $ 0.51\pm0.32$ \\
     \bottomrule
@@
 Indeed, the majority of samples fall within even the more conservative latter range. 
-However, the participants with vision impairments in group \textit{G2}  demonstrate a wider spread in error data and higher average error than \textit{G1} (0.23 vs.
-0.40 radians, Kruskal-Wallis test $p=0.05$).
-\todo{do KW test}
+However, the participants with vision impairments in group \textit{G2}  demonstrate a wider spread in error data and higher average error than \textit{G1} (0.25 vs.
+0.37 radians, Kruskal-Wallis test $p<0.001$).
 The results in~\cite{katz2011spatial,zwiers2001spatial} reported a similar trend, but an investigation with a larger sample size for \textit{G2} should be considered before definitive conclusions can be made.
@@
 \todo{cite vertical vs horizontal performance}
-The mean absolute errors of the participants are \SI{0.37}{\radian} and \SI{0.58}{\radian} for groups \textit{G1} and \textit{G2} respectively.
+The mean absolute errors of the participants are \SI{0.37}{\radian} and \SI{0.51}{\radian} for groups \textit{G1} and \textit{G2} respectively.
 As with the pan estimation errors, the participants from \textit{G1} seemed more adept at estimating the correct elevation.
-However, the difference is not as significant as for the pan, with \textit{G1} displaying a 36\% improvement over \textit{G1} in their elevation estimation vs. 42.5\% in the pan dimension. 
-Comparing these results to methods that convey elevation through an HRTF, we see an improvement of approximately 46\% and 16\% for each group~\cite{schonstein2008comparison}. 
+%However, the difference is not as significant as for the pan, with \textit{G1} displaying a 36\% improvement over \textit{G2} in their elevation estimation vs. 42.5\% in the pan dimension. 
+Comparing these results to methods that convey elevation through an HRTF, we see an improvement of approximately 46\% and 26\% for each group~\cite{schonstein2008comparison}. 
 Indeed, the performance is comparable to that of open-back and expensive in-ear headphones.
-There is a lesser improvement (13\%) with \textit{G1} when compared to the individualised and adjusted HRTFs in~\cite{stanley2006lateralization} and a 13\% reduction in performance for group \textit{G2}.
+There is a lesser improvement (11\%) with \textit{G1} when compared to the individualised and adjusted HRTFs in~\cite{stanley2006lateralization} and a 18\% reduction in performance for group \textit{G2}.
 However, the participants from~\cite{stanley2006lateralization} all had healthy eyesight, making the latter comparison invalid. 

### 2019-07-13 paper.tex
Added previous work, added correlation results

@@
 
-Work on guiding/directing people (vibration, audio, voice)
-Show preference studies
+Multiple mobile navigation and travel aids have been devised throughout the years, some as part of a commercial undertaking and many as part of academic research.
+The majority of these systems rely on one or a combination of vocal\todo{cite voice appers}, audio\todo{cite audio papers} and haptic\todo{cite haptic} feedback media to communicate guidance instructions to the user, each with their own sets of features and limitations.
+In general, participants with vision impairments report that they prefer haptic and vocal feedback out of the 3 options. 
+However, haptic feedback systems require significant additional hardware to transmit the guidance instructions.
+Furthermore, haptic and vocal feedback can become a cognitive burden, particularly where high resolution guidance, which can exceed the bandwidth of these particular senses, is required.
+Indeed, participants report that they would prefer control over the vocal feedback channel and trigger guidance instructions, instead of being given constant guidance instructions. 
+Simple audio tones do not suffer from these bandwidth and hardware limitations as badly, but it does have the potential to fatigue the user's senses if the tone is too harsh, psychologically speaking.
 
-Guidance instruction delivery modes using haptic, audio and vocal feedback for non-visual human-machine-interfaces (HMI) have been investigated before.
-
-Previous work on HRTF stuff
-
-\todo{add previous work}
+Researchers have investigated using audio signals that are spatialised with an HRTF, which tricks the brain into thinking that a sound source is located at some arbitrary 3D position~\cite{geronazzo2016interactive,wilson2007swan,katz2010navig,blum2013spatialized}.
+The authors generally report favourable results for this guidance approach when used with standard over-ear headphones or speakers. 
+However, other authors have found that the choice of audio transmission medium can have a significant effect on performance.
+Cheaper headphones and bone-conduction headphones generally report unfavourable results when compared to over-ear or more expensive alternatives~\cite{schonstein2008comparison,macdonald2006spatial,stanley2006lateralization}.
+This seems to be limited to the elevation dimension, however, and can be improved with HRTFs adjusted for the bone-conduction pathway~\cite{stanley2006lateralization}.
+In~\cite{durette2008visuo}, they transmit a target's elevation angle by adjusting the signal's pitch and report favourable results. 
+However, these studies focussed on their system's total effectiveness, rather than the effectiveness of their pitch adjustment mechanism and in this work we endavoiour to investigate this more closeley.
 
@@
 
+\subsection{Procedure}
+
 To test the interface's effectiveness at directing a user towards a target area, a set of experiments were conducted with a simple pointing task. 
@@
 Similar to the results from~\cite{macdonald2006spatial,schonstein2008comparison}, we use the absolute mean errors for each dimension which allows us to compare our results to the aforementioned works'.
-These results are presented in the boxplots in \cref{fig:results} and summarised in \cref{tab:results}.
+These results are presented in the boxplots in \cref{fig:err-results} and summarised in \cref{tab:results}.
 
@@
   \includegraphics[width=0.7\columnwidth]{figures/pan_tilt_err.png}
-  \caption{Boxplots of the angular errors collected during the pointing experiment. }\label{fig:results}
+  \caption{Boxplots of the angular errors collected during the pointing experiment. }\label{fig:err-results}
   %\subfloat[]{\label{fig:sighted-err}\includegraphics[width=0.5\columnwidth]{figures/pan_tilt_err_sight.png}}
@@
   \centering
-  \caption{}\label{tab:results}
-  \begin{tabular}{p{1cm}cc}
+  \caption{A summary of the results collected from the experiment. }\label{tab:results}
+  \begin{tabular}{p{1cm}p{2cm}p{4cm}c}
     \toprule
-	                         &           & Mean absolute error [rad] \\ \midrule
-    \multirow{2}{*}{\textit{G1}} & Pan       & $ 0.25\pm0.24$ \\
-				 & Elevation & $ 0.37\pm0.28$ \\ \midrule
-    \multirow{2}{*}{\textit{G2}} & Pan       & $ 0.38\pm0.21$ \\
-				 & Elevation & $ 0.51\pm0.32$ \\
+    &           & Mean absolute error [rad]  & Pearson Correlation score \\ \midrule
+    \multirow{2}{*}{\textit{G1}} & Pan       & $0.25\pm0.24$ & $0.83\;(p<0.001)$ \\
+				 & Elevation & $0.37\pm0.28$ & $0.44\;(p<0.001)$ \\ \midrule
+    \multirow{2}{*}{\textit{G2}} & Pan       & $0.38\pm0.21$ & $0.35\;(p<0.001)$ \\
+				 & Elevation & $0.51\pm0.32$ & $0.40\;(p=0.001)$ \\
     \bottomrule
@@
 \end{table}
-
-\todo{add correlation plots}
 
@@
 
-The elevation estimation performance for both groups deteriorated when compared to their pan estimation results, which was somewhat expected given previous research results on human audition.
-\todo{cite vertical vs horizontal performance}
+The elevation estimation performance for both groups deteriorated when compared to their pan estimation results, which was somewhat expected given previous research results on human audition~\cite{barfield1997visual}.
 The mean absolute errors of the participants are \SI{0.37}{\radian} and \SI{0.51}{\radian} for groups \textit{G1} and \textit{G2} respectively.
@@
 
+The error data was used to generate scatter plots that show the correlation between the position that the guidance system presented the participants and the position the participants thought the targets were.
+These plots are given in \cref{fig:correlation-results} and show a reasonably positive correlation between the guidance targets and the selected targets, which indicates that the participants correctly interpreted the interface's output for the most part.
+The Pearson correlation scores for each dataset, listed in \cref{tab:results}, confirm this.
+
+\begin{figure}[t]
+  \centering
+  \includegraphics[width=1.0\columnwidth]{figures/correlation.png}
+  \caption{Boxplots of the angular errors collected during the pointing experiment. }\label{fig:correlation-results}
+  %\subfloat[]{\label{fig:sighted-err}\includegraphics[width=0.5\columnwidth]{figures/pan_tilt_err_sight.png}}
+%~
+  %\subfloat[]{\label{fig:blind-err}\includegraphics[width=0.5\columnwidth]{figures/pan_tilt_err_blind.png}}
+\end{figure}
+
 \section{Conclusion}\label{sec:conclusion}
@@
 
-%\section{Previous Work}\label{sec:lit-review}
-
-%Earlier attempts at improving a person with visual impairment's navigation experience involved outfitting the commonly-used walking cane with remote detection systems, such as sonar, radar and RFID tags, to transform a cane into an early-warning system~\cite{ulrich1997,marion2008batcane,faria2010electronic,willis2005}.
-%Bluetooth tags and a smartphone can be used in a similar manner~\cite{sato2017navcog3}.
-%However, the expense and effort of maintaining such tags make them less attractive for wide-scale adoption.
-%Computer vision-based systems provide a good compromise between usability, cost and accuracy.
-%One solution is to use RGB-D depth sensing cameras, which are becoming increasingly accurate and cost-effective, to build a 3D image map of the environment, and guiding a user through it~\cite{lee2015,rodriguez2012obstacle}.
-%Alternatively, object recognition techniques can be used to detect various objects and landmarks, such as doors, staircases, etc., and communicate their relative location to the user~\cite{schauerte2012assistive,tian2013b,fiannaca2014headlock}, or use audio instructions to guide a user to fully capture an object of interest~\cite{vazquez2012helping}.
-
-%An important feature of user-centric systems is a human-machine interface (HMI) that enables effective two-way communication between the system and the user.
-%Surveys found that people with visual impairments prefer receiving instructions in the form of speech and haptic feedback cues~\cite{khoo2016multimodal,ross2000wearable,vazquez2012helping}.
-%However, haptic feedback modes typically have a lower bandwidth when compared to audio feedback and also require the user to wear a special device in order to transmit the signals effectively.
-%Work has also been done in translating a visual scene into non-visual formats, with so-called sensory substitution systems (e.g.
-`The Voice'~\cite{meijer2010}) and virtual audio reality (VAR) systems~\cite{frauenberger2003} reporting favourable results.
-%However, The Voice has a very steep learning curve that has proven to be a significant barrier to entry and with VAR systems, it is not clear how unknown environments, where markers have not yet been encoded, would be handled and described to the user.
-
-%Previous experiments have determined that people are able to find the location of a stationary sound source with an error of $\pm$\SI{35}{\degree}, in both the pan and tilt dimensions~\cite{zwiers2001spatial}, for both early-blind and normally-sighted people.
-%However, in~\cite{lewald2013exceptional,lessard1998early} it was found that the blind have a clear advantage in localisation accuracy over sighted people when presented with more complex tasks, such as targets in motion and narrow-band stimuli, with~\cite{lewald2013exceptional} reporting an average absolute localisation error of around \SI{10}{\degree} in the pan and tilt dimensions.
-%Researchers have used simulated spatialised audio to inform the user of the sound source's direction~\cite{holland2002audiogps,kammoun2012navigation,rebillat2009smart,menelas2010audio,wilson2007swan,zotkin2004rendering}.
-%In these works, a sound was played through a set of headphones and the source spatialised with a head-related transfer function (HRTF), tricking the listener into thinking the sound source was located at some arbitrary 3D location.
-%There are also experimental results about how well users can find targets presented with spatial sound in the tilt and panning dimensions with normal headphones~\cite{katz2011spatial,zwiers2001spatial}, as well as work that has shown similar levels of localisation performance between normal and bone-conducting headphones in the pan dimension~\cite{macdonald2006spatial}.
-%However, to our knowledge no extensive work or experiments have been done to determine how well users respond to tilt adjustment instructions using a tone with \textit{varying pitch}, in particular when using bone-conducting headphones, and how changes to this tone's behaviour affects performance.
-
 \bibliographystyle{splncs04}

### 2019-07-13 paper.tex
More references

@@
 Multiple mobile navigation and travel aids have been devised throughout the years, some as part of a commercial undertaking and many as part of academic research.
-The majority of these systems rely on one or a combination of vocal\todo{cite voice appers}, audio\todo{cite audio papers} and haptic\todo{cite haptic} feedback media to communicate guidance instructions to the user, each with their own sets of features and limitations.
+The majority of these systems rely on one or a combination of vocal~\cite{mocanu2016when,chessa2016integrated,kanwal2015navigation}, audio~\cite{schwarze2015intuitive,rodriguez2012obstacle,katz2010navig} and haptic~\cite{rivera-rubio2015assistive,lee2015rgb,xiao2015assistive} feedback media to communicate guidance instructions to the user, each with their own sets of features and limitations.
 In general, participants with vision impairments report that they prefer haptic and vocal feedback out of the 3 options. 
@@
 However, the participants from~\cite{stanley2006lateralization} all had healthy eyesight, making the latter comparison invalid. 
-Interestingly, the elevation errors for group \textit{G2} is centered around the negative angle, indicating a possible underlying bias in the participants or the interface.
+Interestingly, the elevation errors for group \textit{G2} are centred around the negative angle, indicating a possible underlying bias in the participants or the interface.
 A similar trend was observed in~\cite{stanley2006lateralization} and warrants an investigation with a larger participant pool to determine the underlying bias, if any.

### 2019-07-14 paper.tex
Small revisions

@@
 
-\title{Audio Interface of an Active Perception Navigation Aid for People with Visual Impairments}
-\titlerunning{Active Perception Interface for People with Visual Impairments}
+\title{Viability of Bone-Conduction and Spatialised Audio to Provide Guidance Instructions for People with Visual Impairments}
+\titlerunning{Guiding People with Bone-Conduction and Spatialised Audio}
 
@@
 \begin{abstract}
-  The ActiVis project is attempting to build a mobile guidance aid to help people with limited vision search for objects within an unknown environment.
-  This system requires an effective non-visual interface to transmit guidance instructions to the user through means of bone-conducting headphones.
-  To this end we propose a audio-based interface that uses a spatialised signal to convey a target's position on the horizontal plane. 
-  The position on the median plan is given by adjusting the tone's pitch to overcome audio localisation issues that arise with bone-conducting headphones. 
+  The ActiVis project is attempting to build a mobile guidance aid to help people with limited vision find objects in an unknown environment.
+  This system uses bone-conducting headphones to transmit audio signals to the user and requires an effective non-visual interface.
+  To this end, we propose an audio-based interface that uses a spatialised signal to convey a target's position on the horizontal plane. 
+  The position on the median plan is given by adjusting the tone's pitch to overcome the audio localisation limitations of bone-conducting headphones. 
   This interface is validated through a set of experiments. 
-\keywords{Human-machine interface \and vision impairment \and spatialised sound \and varying pitch, pointing task}
+\keywords{Human-machine interface \and vision impairment \and spatialised sound \and varying pitch \and bone-conduction}
 \end{abstract}
@@
 
-\todo{review title to make focus more on bone conduction and its viability}
 In recent years, governments have spearheaded numerous initiatives to support people with disabilities and enable them to play a more active role in modern society.
@@
 Improvements in modern computing have made it possible for new and innovative solutions to these problems to come to the fore.
-In particular, researchers in the active vision field have have made much progress in enabling machines to autonomously manipulate cameras to gather information about an environment for mapping and object and object finding tasks~\cite{bajcsy2018revisiting}.
+In particular, researchers in the active vision field have have made much progress in enabling machines to autonomously manipulate cameras to gather information about an environment for mapping, as well as object and object finding tasks~\cite{bajcsy2018revisiting}.
 There is, however, a significant research question about whether techniques from active machine vision can be applied to humans, i.e.\ can a machine identify a point of interest in a scene and direct a human, instead of an electronic servo, to focus on that point?
@@
 
-The ActiVis project aims to deliver a mobile guidance system based on the Google Project Tango platform\footnote{\url{https://en.wikipedia.org/wiki/Tango\_(platform)}}, pictured in \cref{fig:tango-headphone}, that will ultimately be able to guide a user with vision impairments on the last leg of their journey, i.e.\ the so-called `last 10-yard problem'.
+The ActiVis project aims to deliver a mobile guidance system based on the Google Project Tango platform~\footnote{\url{https://en.wikipedia.org/wiki/Tango\_(platform)}}, pictured in \cref{fig:tango-headphone}, that will ultimately be able to guide a user with vision impairments on the last leg of their journey, i.e.\ the so-called `last 10-yard problem'.
 The entire platform is based on a Google Project Tango device that provides access to powerful real-time localisation (through IMU measurements and landmark tracking) and image-processing facilities and provides access to Android's full range of interface tools and IO options.
-Furthermore, a set of bone-conducting headphones (pictured in \cref{fig:tango-headphone}) that are placed on a user's cheekbones instead of their ears and do not interfere with normal hearing, are used to transmit the audio signals to the user.
-The system provides the user with real-time guidance instructions and it is important that these instructions are easily understood and do not add a significant cognitive burden.
+Furthermore, a set of bone-conducting headphones (pictured in \cref{fig:tango-headphone}), that are placed on a user's cheekbones instead of their ears and do not interfere with normal hearing, are used to transmit the audio signals to the user.
 
@@
 
-Humans are naturally able to determine the 3D position of a sound source and by exploiting this ability, our system's guidance instructions can naturally be interpreted without posing a significant cognitive load.
-A sound source can be spatialised by adjusting a tone's spectral make-up (elevation angle), time delay and level difference (pan angle) and the intensity (distance).
-Here only the pan and elevation positions are transmitted.
+Humans are able to determine the 3D position of a sound source and by exploiting this natural ability, the real-time guidance instructions can easily be interpreted without posing a significant cognitive load.
+A sound source can be spatialised by adjusting a tone's spectral make-up (elevation angle), time delay and level difference (pan angle) and intensity (distance).
+In this case, only the pan and elevation positions are transmitted.
 However, since bone-conducting headphones bypass the outer ear structure, their spectral signature cannot properly be interpreted and we therefore convey the target's elevation angle by adjusting the tone's pitch.
-A similar approach was used in\cite{durette2008visuo}, but the study did not focus on or investigate the interface's efficacy.
-Indeed, this work serves as an initial study into an interface that can effectively communicate the pan and elevation angles of a target location through a set of bone-conducting headphones using a standard head-related transfer function (HRTF).
-Furthermore, we demonstrate our approach's effectiveness through a simple experiment.
+A similar approach was used in~\cite{durette2008visuo}, but the study did not focus on or investigate the interface's efficacy.
+Indeed, this work serves as an initial study into an interface that can effectively communicate the pan and elevation angles of a target location through a set of bone-conducting headphones using standard audio libraries and audio manipulations.
+Furthermore, we demonstrate this approach's effectiveness through a simple experiment.
 
@@
 The majority of these systems rely on one or a combination of vocal~\cite{mocanu2016when,chessa2016integrated,kanwal2015navigation}, audio~\cite{schwarze2015intuitive,rodriguez2012obstacle,katz2010navig} and haptic~\cite{rivera-rubio2015assistive,lee2015rgb,xiao2015assistive} feedback media to communicate guidance instructions to the user, each with their own sets of features and limitations.
-In general, participants with vision impairments report that they prefer haptic and vocal feedback out of the 3 options. 
-However, haptic feedback systems require significant additional hardware to transmit the guidance instructions.
+In general, participants with vision impairments report that they prefer haptic and vocal feedback out of the 3 options.\todo{ref blind preferences}
+However, haptic feedback systems typically have extra hardware requirements to transmit the guidance instructions.
 Furthermore, haptic and vocal feedback can become a cognitive burden, particularly where high resolution guidance, which can exceed the bandwidth of these particular senses, is required.
@@
 In~\cite{durette2008visuo}, they transmit a target's elevation angle by adjusting the signal's pitch and report favourable results. 
-However, these studies focussed on their system's total effectiveness, rather than the effectiveness of their pitch adjustment mechanism and in this work we endavoiour to investigate this more closeley.
+However, these studies focussed on their system's total effectiveness, rather than the effectiveness of their pitch adjustment mechanism and in this work we endeavour to investigate this more closely.
 
@@
 Indeed, the traditional walking cane remains the go-to tool for navigation and obstacle avoidance. 
-With the ActiVis system, we aim to tackle the issue of cost, user-acceptance and usability by using as many as possible `standard' devices and hardware as possible, with the idea that no specialist hardware is required.
-Following this design target, we base the project on a mobile device and discreet headset with the anticipation that the user will not stand out if seen walking in public with a phone in their hand and wearing a headset.
-The ActiVis system is based on a concept proposed in~\cite{bellotto2013multimodal,lock2017portable}, which uses a Google Tango device that is an Android-based device that comes equipped with an RGB-D camera to sense colour and depth and combines an inertial measurement unit (IMU) with powerful and robust landmark recognition and image processing algorithms to localise itself in real-time.
-An benefit of such a platform is its familiar, compact form-factor which will help overcome the hurdle of user-acceptance and usability.
+With the ActiVis system, we aim to tackle the issue of cost, user-acceptance and usability by using only `standard' devices and hardware as far as possible and avoiding any special requirements.
+Following this design target, we base the project on a mobile device and discreet headset with the expectation that the user will not stand out if seen walking in public with a phone in their hand and wearing a headset.
+The ActiVis system is based on a concept proposed in~\cite{bellotto2013multimodal,lock2017portable}, which uses a Google Tango device that is able to localise itself in real-time and generate and has the benefit of a compact, familiar form-factor which will help overcome the hurdle of user-acceptance and usability.
 
-A set of bone-conducting headphones was selected as the audio transmission medium.
+A set of bone-conducting headphones is used as the audio transmission medium.
 These headphones sit on a user's cheekbones and conduct the audio signals through the skull into the inner ear, instead of through the outer ear like typical over-ear headphones do. 
@@
 Humans localise a sound source in 3 dimensions by considering cues recorded in one ear (monaural cues) and comparing cues received at both ears (binaural cues)~\cite{blauert1997spatial,blauert1969sound}.
-The binaural cues include time differences of arrival and intensity differences that help to determine a source's location on the horizontal plane.
+The binaural cues include ITD and ILD that help to determine a source's location on the horizontal plane.
 Monaural cues are taken from the interaction of the sound with the human anatomy, e.g.\ head, shoulders, outer ear, before it enters the ear canal.
-The user's anatomy acts as a filter and the direction the sound wave hits the body activates different filter resonances, attenuating or accentuating different frequency bands.
+%The user's anatomy acts as a filter and the direction the sound wave hits the body activates different filter resonances, attenuating or accentuating different frequency bands.
 When the modified audio signal enters the inner ear canal, the human brain is able to analyse the frequency response and accurately determine the position of the sound source on the median plane. 
@@
 
-When an audio signal is transmitted via a set of speakers or headsets, it can be transformed to mimic the characteristics of a natural sound source before it is transmitted, tricking the brain into believing a sound is located at some arbitrary position.
-Such a transformation can be done using a head-related transfer function (HRTF) and a pair of stereo headsets or speakers.
+When an audio signal is transmitted via a set of speakers or headsets, it can be transformed with an HRTF to mimic the characteristics of a natural sound source before it is transmitted, tricking the brain into believing a sound is located at some arbitrary position.
 An HRTF is a mathematical function that simulates the response signal of a human head and is derived by capturing key characteristics that affect the monaural and binaural responses, such as the user's hearing levels and head size, etc.
 Since hearing responses are unique amongst different users, the best results would be observed if each user had their own customised HRTF.\
-However, given the complicated process involved to capture the required user characteristics, making unique HRTFs is often an untenable solution and using average values (e.g.\ head measurements, height, etc.) have shown to produce acceptable results.\todo{cite HRTF papers}
+However, given the complicated process involved to capture the required user characteristics, making unique HRTFs is often an untenable solution and using average values (e.g.\ head measurements, height, etc.) have shown to produce acceptable results~\cite{gardner1995hrtf}.
 
@@
 The guidance target is presented to the user in terms of pan and elevation angles, indicating the angle adjustments that are required to point the device camera at the target location, as shown in \cref{fig:cam-coords}.
-Spatialised audio signals are well-suited to the task, displaying similar levels of performance to vocal feedback, but with less cognitive load and higher resolution\cite{klatzky2006cognitive}.
-However, the audio guidance signals are delivered to the user with a set of bone-conducting headphones that bypass the user's outer-ear structure and since this structure plays an important role in localising a sound source's elevation\cite{blauert1969sound}, an alternative mode of conveying the elevation angle is required.
-For this work, we propose a simple linear adjustment to the signal's pitch as a function of the elevation angle. 
+Spatialised audio signals are well-suited to the task, displaying similar levels of performance to vocal feedback, but with less cognitive load and higher resolution~\cite{klatzky2006cognitive}.
+However, the audio guidance signals are delivered to the user with a set of bone-conducting headphones that bypass the user's outer-ear structure and since this structure plays an important role in localising a sound source's elevation\cite{blauert1969sound}, an alternative mode of conveying the elevation angle is required.\todo{check repeating from hardware selection section}
+Therefore, we propose a simple linear adjustment to the signal's pitch as a function of the elevation angle. 
 The pan angle can be conveyed by transforming the audio signal with an HRTF, and indeed it has been found that this dimension is unaffected by using bone-conducting headphones\cite{schonstein2008comparison,macdonald2006spatial,stanley2006lateralization}.
@@
 People typically have trouble localising a pure tone devoid of a rich spectral signature.
-However, the ITD and ILD is independent of the tone's spectral make-up and the elevation angle is given through a different mechanism.
-A pure sine wave is therefore sufficient to convey the target's pan angle.
-To transform and spatialise the audio signal, we used OpenAL's default HRTF, based on the MIT's KEMAR dataset~\cite{hiebert2005openal,gardner1995hrtf}, that uses the user and targets' positions as input and outputs a transformed audio signal.
+However, the ITD and ILD is independent of the tone's spectral make-up and the elevation angle is given through a different mechanism and a pure sine wave is therefore sufficient to convey the target's pan angle.
+To transform and spatialise the audio signal, we used OpenAL's default HRTF, based on the MIT's KEMAR dataset~\cite{hiebert2005openal}, that uses the user and targets' positions as input and outputs a transformed audio signal.
 
@@
 
-Using a generic HRTF, such as OpenAL's KEMAR implementation, with bone-conducting headphones has been shown to have trouble conveying the elevation angle of a sound source accurately~\cite{macdonald2006spatial,schonstein2008comparison}.
+A generic HRTF implementation with bone-conducting headphones has trouble conveying the elevation angle of a sound source accurately~\cite{macdonald2006spatial,schonstein2008comparison}.
 To compensate for this, we convey the target's elevation angle by adjusting the tone's pitch (i.e.\ the sine wave's frequency) as a function of the elevation. 
@@
 
-To test the interface's effectiveness at directing a user towards a target area, a set of experiments were conducted with a simple pointing task. 
-This pointing task is set up to capture the difference between the targets' actual position and the positions the participants' perceived them to be.
-
-For the experiments, the participants are given a Tango tablet device running an app specifically written for these experiments. 
-This app generates a set of virtual targets and presents them to the participants one at a time. 
+To test the interface's effectiveness at directing a user towards a target area, a set of experiments were conducted with a pointing task that is set up to capture the difference between the targets' actual position and the positions the participants' perceived them to be.
+The participants are given a Tango tablet device running an app written for the experiment that generates a set of virtual targets and presents them to each participants one at a time. 
 The targets are generated at a constant \SI{2}{\meter} from the participant and their the pan and elevation angles are uniformly generated across 4 quadrants to avoid clustering.
@@
 When a participant was confident that they were on-target, i.e.\ hearing the audio front-on at \SI{512}{\hertz}, they tapped the screen, marking the location and generating the next target.
-The targets' positions are all set relative to the Tango device's coordinate system which is tracked using the Tango localisation API.\
+The targets' positions are all set relative to the Tango device's coordinate system which is tracked using the Tango hardware and localisation API.\
 A total of 28 targets were generated per participant.
@@
   \caption{Boxplots of the angular errors collected during the pointing experiment. }\label{fig:err-results}
-  %\subfloat[]{\label{fig:sighted-err}\includegraphics[width=0.5\columnwidth]{figures/pan_tilt_err_sight.png}}
-%~
-  %\subfloat[]{\label{fig:blind-err}\includegraphics[width=0.5\columnwidth]{figures/pan_tilt_err_blind.png}}
 \end{figure}
@@
   \caption{Boxplots of the angular errors collected during the pointing experiment. }\label{fig:correlation-results}
-  %\subfloat[]{\label{fig:sighted-err}\includegraphics[width=0.5\columnwidth]{figures/pan_tilt_err_sight.png}}
-%~
-  %\subfloat[]{\label{fig:blind-err}\includegraphics[width=0.5\columnwidth]{figures/pan_tilt_err_blind.png}}
 \end{figure}
@@
 Results generated from a simple pointing experiment are encouraging and improves performance over purely spatialised signals transmitted via bone-conduction.
+Furthermore, the limitations of the bone-conducting headphones were largely negated by conveying the elevation angle of the target by adjusting the tone's pitch, with pan and elevation errors comparable to that from normal over-ear headphones.
+
 Future work will involve a more in-depth investigation with a larger participant pool to determine the factors that affect performance and enable a more detailed statistical analysis.

### 2019-07-14 paper.tex
Modifed bib

@@
 Results generated from a simple pointing experiment are encouraging and improves performance over purely spatialised signals transmitted via bone-conduction.
-Furthermore, the limitations of the bone-conducting headphones were largely negated by conveying the elevation angle of the target by adjusting the tone's pitch, with pan and elevation errors comparable to that from normal over-ear headphones.
+Furthermore, the limitations of these headphones were largely negated by conveying the elevation angle of the target by adjusting the tone's pitch, with pan and elevation errors comparable to that from normal over-ear headphones.
 

### 2019-07-15 paper.tex
revised up to sec 2

@@
 
+\usepackage{soul}
+\DeclareRobustCommand{\tofix}[1]{{\sethlcolor{yellow}\hl{[#1]}}}
+
 \begin{document}
 
-\title{Viability of Bone-Conduction and Spatialised Audio to Provide Guidance Instructions for People with Visual Impairments}
+\title{Bone-Conduction Audio Interface to\\Guide People with Visual Impairments}
 \titlerunning{Guiding People with Bone-Conduction and Spatialised Audio}
@@
 
-\institute{University of Lincoln \email{\{jlock, gcielniak, nbellotto\}@lincoln.ac.uk} \url{https://lcas.github.io/ActiVis/} \and
+\institute{University of Lincoln \email{\{jlock, gcielniak, nbellotto\}@lincoln.ac.uk} \and
 	   University of Bristol \email{i.d.gilchrist@bristol.ac.uk}}
@@
 \begin{abstract}
-  The ActiVis project is attempting to build a mobile guidance aid to help people with limited vision find objects in an unknown environment.
-  This system uses bone-conducting headphones to transmit audio signals to the user and requires an effective non-visual interface.
-  To this end, we propose an audio-based interface that uses a spatialised signal to convey a target's position on the horizontal plane. 
-  The position on the median plan is given by adjusting the tone's pitch to overcome the audio localisation limitations of bone-conducting headphones. 
-  This interface is validated through a set of experiments. 
+  The ActiVis\footnote{\url{https://lcas.github.io/ActiVis/}} project's aim is to build a mobile guidance aid to help people with limited vision find objects in an unknown environment.
+  This system uses bone-conduction headphones to transmit audio signals to the user and requires an effective non-visual interface.
+  To this end, we propose a new audio-based interface that uses a spatialised signal to convey a target's position on the horizontal plane. 
+  The vertical position on the median plan is given by adjusting the tone's pitch to overcome the audio localisation limitations of bone-conduction headphones. 
+  This interface is validated through a set of experiments with blindfolded and visually impaired users. 
 \keywords{Human-machine interface \and vision impairment \and spatialised sound \and varying pitch \and bone-conduction}
@@
 Improvements in modern computing have made it possible for new and innovative solutions to these problems to come to the fore.
-In particular, researchers in the active vision field have have made much progress in enabling machines to autonomously manipulate cameras to gather information about an environment for mapping, as well as object and object finding tasks~\cite{bajcsy2018revisiting}.
+In particular, researchers in the active vision field have have made much progress in enabling machines to autonomously manipulate cameras to gather information about an environment for mapping and object finding tasks~\cite{bajcsy2018revisiting,lock2019active}.
 There is, however, a significant research question about whether techniques from active machine vision can be applied to humans, i.e.\ can a machine identify a point of interest in a scene and direct a human, instead of an electronic servo, to focus on that point?
@@
 
-The ActiVis project aims to deliver a mobile guidance system based on the Google Project Tango platform~\footnote{\url{https://en.wikipedia.org/wiki/Tango\_(platform)}}, pictured in \cref{fig:tango-headphone}, that will ultimately be able to guide a user with vision impairments on the last leg of their journey, i.e.\ the so-called `last 10-yard problem'.
-The entire platform is based on a Google Project Tango device that provides access to powerful real-time localisation (through IMU measurements and landmark tracking) and image-processing facilities and provides access to Android's full range of interface tools and IO options.
-Furthermore, a set of bone-conducting headphones (pictured in \cref{fig:tango-headphone}), that are placed on a user's cheekbones instead of their ears and do not interfere with normal hearing, are used to transmit the audio signals to the user.
+The ActiVis project aims to deliver a mobile guidance system, pictured in \cref{fig:tango-headphone}, that will ultimately be able to guide a user with vision impairments on the last leg of their journey, i.e.\ the so-called `last 10-yard problem'.
+The prototype platform is based on a Google Project Tango\footnote{\url{https://en.wikipedia.org/wiki/Tango\_(platform)}} device that provides access to powerful real-time localisation (through IMU measurements and landmark tracking) and image-processing facilities.
+It also provides access to Android's full range of interface tools and I/O options.
+Furthermore, a set of bone-conduction headphones (pictured in \cref{fig:tango-headphone}), which are placed on a user's cheekbones instead of their ears and do not interfere with normal hearing, are used to transmit the audio signals to the user.
 
@@
   \subfloat[]{\label{fig:participant}\includegraphics[width=0.25\columnwidth]{figures/vi_participant.pdf}}
-  \caption{Pictures of the Tango device and bone-conducting headphones the system is based on (left) and both in use in an experiment (right).}
+  \caption{Pictures of the Tango device and bone-conduction headphones the system is based on (left) and both in use in an experiment (right).\tofix{add a picture of the Tango phone}}
 \end{figure}
@@
 Humans are able to determine the 3D position of a sound source and by exploiting this natural ability, the real-time guidance instructions can easily be interpreted without posing a significant cognitive load.
-A sound source can be spatialised by adjusting a tone's spectral make-up (elevation angle), time delay and level difference (pan angle) and intensity (distance).
-In this case, only the pan and elevation positions are transmitted.
-However, since bone-conducting headphones bypass the outer ear structure, their spectral signature cannot properly be interpreted and we therefore convey the target's elevation angle by adjusting the tone's pitch.
+A sound source can be spatialised by adjusting a tone's spectral make-up (elevation angle), time delay and level difference (pan angle), and intensity (distance).
+In our case, only the pan and elevation positions are transmitted for an object pointing task.
+However, since bone-conduction headphones bypass the outer ear structure, their spectral signature cannot properly be interpreted and we therefore convey the target's elevation angle by adjusting the tone's pitch.
 A similar approach was used in~\cite{durette2008visuo}, but the study did not focus on or investigate the interface's efficacy.
-Indeed, this work serves as an initial study into an interface that can effectively communicate the pan and elevation angles of a target location through a set of bone-conducting headphones using standard audio libraries and audio manipulations.
+Indeed, the work presented in this paper serves as an initial study into an interface that can effectively communicate the pan and elevation angles of a target location through a set of bone-conduction headphones using standard audio libraries and audio manipulations.
 Furthermore, we demonstrate this approach's effectiveness through a simple experiment.
 
-The rest of the paper starts off by discussing previous relevant works and research in \cref{sec:prev-work}, followed by a discussion on the design and implementation of our interface in \cref{sec:interface-design}.
+The rest of the paper starts by discussing previous relevant works and research in \cref{sec:prev-work}, followed by a discussion on the design and implementation of our interface in \cref{sec:interface-design}.
 This is followed by a description of the experiments that were conducted and a discussion of their results in \cref{sec:experiment-and-results}.
-Finally, the work is concluded with a summary and a short discussion on future research prospects in \cref{sec:conclusion}.
+Finally, the paper concludes with a summary and a short discussion on future research prospects in \cref{sec:conclusion}.
 
@@
 The majority of these systems rely on one or a combination of vocal~\cite{mocanu2016when,chessa2016integrated,kanwal2015navigation}, audio~\cite{schwarze2015intuitive,rodriguez2012obstacle,katz2010navig} and haptic~\cite{rivera-rubio2015assistive,lee2015rgb,xiao2015assistive} feedback media to communicate guidance instructions to the user, each with their own sets of features and limitations.
-In general, participants with vision impairments report that they prefer haptic and vocal feedback out of the 3 options.\todo{ref blind preferences}
+In general, participants with vision impairments report that they prefer haptic and vocal feedback out of the three options. \todo{ref blind preferences}
 However, haptic feedback systems typically have extra hardware requirements to transmit the guidance instructions.
 Furthermore, haptic and vocal feedback can become a cognitive burden, particularly where high resolution guidance, which can exceed the bandwidth of these particular senses, is required.
-Indeed, participants report that they would prefer control over the vocal feedback channel and trigger guidance instructions, instead of being given constant guidance instructions. 
-Simple audio tones do not suffer from these bandwidth and hardware limitations as badly, but it does have the potential to fatigue the user's senses if the tone is too harsh, psychologically speaking.
+Indeed, participants report that they would prefer control over the vocal feedback channel and trigger guidance instructions, instead of being given constant guidance instructions.\tofix{reference?}
+Simple audio tones are less affected by these bandwidth and hardware limitations, but they can potentially fatigue the user's senses if too unpleasant.
 
-Researchers have investigated using audio signals that are spatialised with an HRTF, which tricks the brain into thinking that a sound source is located at some arbitrary 3D position~\cite{geronazzo2016interactive,wilson2007swan,katz2010navig,blum2013spatialized}.
+Researchers have investigated using audio signals that are spatialised with a head-related transfer function (HRTF), simulating a sound source located at some arbitrary 3D position~\cite{geronazzo2016interactive,wilson2007swan,katz2010navig,blum2013spatialized}.
 The authors generally report favourable results for this guidance approach when used with standard over-ear headphones or speakers. 
@@
 In~\cite{durette2008visuo}, they transmit a target's elevation angle by adjusting the signal's pitch and report favourable results. 
-However, these studies focussed on their system's total effectiveness, rather than the effectiveness of their pitch adjustment mechanism and in this work we endeavour to investigate this more closely.
+However, these studies focussed on their system's total effectiveness, rather than the effectiveness of the pitch adjustment mechanism alone.
+In the next sections we want to investigate this more closely.
 
@@
 
-A set of bone-conducting headphones is used as the audio transmission medium.
+A set of bone-conduction headphones is used as the audio transmission medium.
 These headphones sit on a user's cheekbones and conduct the audio signals through the skull into the inner ear, instead of through the outer ear like typical over-ear headphones do. 
@@
 Spatialised audio signals are well-suited to the task, displaying similar levels of performance to vocal feedback, but with less cognitive load and higher resolution~\cite{klatzky2006cognitive}.
-However, the audio guidance signals are delivered to the user with a set of bone-conducting headphones that bypass the user's outer-ear structure and since this structure plays an important role in localising a sound source's elevation\cite{blauert1969sound}, an alternative mode of conveying the elevation angle is required.\todo{check repeating from hardware selection section}
+However, the audio guidance signals are delivered to the user with a set of bone-conduction headphones that bypass the user's outer-ear structure and since this structure plays an important role in localising a sound source's elevation\cite{blauert1969sound}, an alternative mode of conveying the elevation angle is required.\todo{check repeating from hardware selection section}
 Therefore, we propose a simple linear adjustment to the signal's pitch as a function of the elevation angle. 
-The pan angle can be conveyed by transforming the audio signal with an HRTF, and indeed it has been found that this dimension is unaffected by using bone-conducting headphones\cite{schonstein2008comparison,macdonald2006spatial,stanley2006lateralization}.
+The pan angle can be conveyed by transforming the audio signal with an HRTF, and indeed it has been found that this dimension is unaffected by using bone-conduction headphones\cite{schonstein2008comparison,macdonald2006spatial,stanley2006lateralization}.
 
@@
 
-A generic HRTF implementation with bone-conducting headphones has trouble conveying the elevation angle of a sound source accurately~\cite{macdonald2006spatial,schonstein2008comparison}.
+A generic HRTF implementation with bone-conduction headphones has trouble conveying the elevation angle of a sound source accurately~\cite{macdonald2006spatial,schonstein2008comparison}.
 To compensate for this, we convey the target's elevation angle by adjusting the tone's pitch (i.e.\ the sine wave's frequency) as a function of the elevation. 
@@
 
-In this work, we proposed an audio interface that conveys a target's pan and elevation angles through a spatialised audio tone that adjusts the tone's pitch to convey the elevation in compensate for the shortcomings of bone-conducting headsets in transmitting the elevation component of a spatialised signal.
+In this work, we proposed an audio interface that conveys a target's pan and elevation angles through a spatialised audio tone that adjusts the tone's pitch to convey the elevation in compensate for the shortcomings of bone-conduction headsets in transmitting the elevation component of a spatialised signal.
 Results generated from a simple pointing experiment are encouraging and improves performance over purely spatialised signals transmitted via bone-conduction.

### 2019-07-15 paper.tex
Added reference, removed wrong statement. Fixed todo

@@
 
-\title{Bone-Conduction Audio Interface to\\Guide People with Visual Impairments}
-\titlerunning{Guiding People with Bone-Conduction and Spatialised Audio}
+%\title{Bone-Conduction Audio Interface to\\Guide People with Visual Impairments}
+\title{Bone-Conduction and Spatialised Audio to Guide People with Visual Impairments}
 
@@
 The majority of these systems rely on one or a combination of vocal~\cite{mocanu2016when,chessa2016integrated,kanwal2015navigation}, audio~\cite{schwarze2015intuitive,rodriguez2012obstacle,katz2010navig} and haptic~\cite{rivera-rubio2015assistive,lee2015rgb,xiao2015assistive} feedback media to communicate guidance instructions to the user, each with their own sets of features and limitations.
-In general, participants with vision impairments report that they prefer haptic and vocal feedback out of the three options. \todo{ref blind preferences}
+In general, participants with vision impairments report that they prefer haptic and vocal feedback out of the three options~\cite{arditi2013user}.
 However, haptic feedback systems typically have extra hardware requirements to transmit the guidance instructions.
@@
 Humans localise a sound source in 3 dimensions by considering cues recorded in one ear (monaural cues) and comparing cues received at both ears (binaural cues)~\cite{blauert1997spatial,blauert1969sound}.
-The binaural cues include ITD and ILD that help to determine a source's location on the horizontal plane.
+The binaural cues include inter-aural time and level differences (ITD and ILD respectively) that help to determine a source's location on the horizontal plane.
 Monaural cues are taken from the interaction of the sound with the human anatomy, e.g.\ head, shoulders, outer ear, before it enters the ear canal.
@@
 Spatialised audio signals are well-suited to the task, displaying similar levels of performance to vocal feedback, but with less cognitive load and higher resolution~\cite{klatzky2006cognitive}.
-However, the audio guidance signals are delivered to the user with a set of bone-conduction headphones that bypass the user's outer-ear structure and since this structure plays an important role in localising a sound source's elevation\cite{blauert1969sound}, an alternative mode of conveying the elevation angle is required.\todo{check repeating from hardware selection section}
-Therefore, we propose a simple linear adjustment to the signal's pitch as a function of the elevation angle. 
-The pan angle can be conveyed by transforming the audio signal with an HRTF, and indeed it has been found that this dimension is unaffected by using bone-conduction headphones\cite{schonstein2008comparison,macdonald2006spatial,stanley2006lateralization}.
+However, given the previously discussed limitations of bone-conduction and spatialised audio, we propose a simple linear adjustment to the signal's pitch as a function of the elevation angle. 
+The pan angle can be conveyed by transforming the audio signal with an HRTF, and indeed it has been found that this dimension is unaffected by using bone-conducting headphones\cite{schonstein2008comparison,macdonald2006spatial,stanley2006lateralization}.
 
@@
 
-The human audition system uses binaural comparison cues, such as the inter-aural time and level differences (ITD and ILD respectively), to localise a sound source on the horizontal plane~\cite{blauert1969sound}.
+The human audition system uses binaural comparison cues, such as ITD and ILD, to localise a sound source on the horizontal plane~\cite{blauert1969sound}.
 The ITD is the perceived time delay between the signal reaching both ears, while the ILD is the perceived volume difference in the signal.
@@
 However, the participants from~\cite{stanley2006lateralization} all had healthy eyesight, making the latter comparison invalid. 
-Interestingly, the elevation errors for group \textit{G2} are centred around the negative angle, indicating a possible underlying bias in the participants or the interface.
+Interestingly, the elevation errors for group \textit{G2} are centred around the negative angle, indicating a possible underlying bias in the participants or the interface\todo{show}.
 A similar trend was observed in~\cite{stanley2006lateralization} and warrants an investigation with a larger participant pool to determine the underlying bias, if any.
@@
 These plots are given in \cref{fig:correlation-results} and show a reasonably positive correlation between the guidance targets and the selected targets, which indicates that the participants correctly interpreted the interface's output for the most part.
-The Pearson correlation scores for each dataset, listed in \cref{tab:results}, confirm this.
+The Pearson correlation scores for each dataset, listed in \cref{tab:results}, confirm this, as well as the negative angle bias in the elevation dimensions.
 
@@
 Furthermore, experiments with a large group of people with vision impairments will provide an opportunity to gauge their opinion on the interface and identify possible avenues of improvements to user friendliness.
+Incorporating a progressive co-adaptation mechanism into the system, as suggested in~\cite{gallina2015progressive}, can also be considered to improve the overall user-friendliness and guidance performance.
 

### 2019-07-15 paper.tex
Adjusted table spacing

@@
 The ActiVis project aims to deliver a mobile guidance system, pictured in \cref{fig:tango-headphone}, that will ultimately be able to guide a user with vision impairments on the last leg of their journey, i.e.\ the so-called `last 10-yard problem'.
-The prototype platform is based on a Google Project Tango\footnote{\url{https://en.wikipedia.org/wiki/Tango\_(platform)}} device that provides access to powerful real-time localisation (through IMU measurements and landmark tracking) and image-processing facilities.
+The prototype platform is based on a Google Project Tango\footnote{\url{https://en.wikipedia.org/wiki/Tango\_(platform)}} device that embeds a color camera and provides access to powerful real-time localisation (through IMU measurements and landmark tracking) and image-processing facilities.
 It also provides access to Android's full range of interface tools and I/O options.
@@
 A sound source can be spatialised by adjusting a tone's spectral make-up (elevation angle), time delay and level difference (pan angle), and intensity (distance).
-In our case, only the pan and elevation positions are transmitted for an object pointing task.
+In our case, only the pan and elevation positions are transmitted to the user to point the camera towards a target object or visual feature.
 However, since bone-conduction headphones bypass the outer ear structure, their spectral signature cannot properly be interpreted and we therefore convey the target's elevation angle by adjusting the tone's pitch.
@@
 
-Electronic navigation aids and guidance systems for people with vision impairments have had trouble appealing to the visually impaired community at large because of problems that are common of academic research and prototypes: prohibitive costs, unfriendly user interfaces and significant hardware requirements, e.g.\ head-mounted cameras and GPS antennae~\cite{golledge2004stated,yusif2016older,arditi2013user}.
-Indeed, the traditional walking cane remains the go-to tool for navigation and obstacle avoidance. 
-With the ActiVis system, we aim to tackle the issue of cost, user-acceptance and usability by using only `standard' devices and hardware as far as possible and avoiding any special requirements.
-Following this design target, we base the project on a mobile device and discreet headset with the expectation that the user will not stand out if seen walking in public with a phone in their hand and wearing a headset.
-The ActiVis system is based on a concept proposed in~\cite{bellotto2013multimodal,lock2017portable}, which uses a Google Tango device that is able to localise itself in real-time and generate and has the benefit of a compact, familiar form-factor which will help overcome the hurdle of user-acceptance and usability.
+Electronic navigation aids and guidance systems for people with vision impairments often have acceptability and usability problems due to limitations in academic research and prototypes: prohibitive costs, unfriendly user interfaces, significant hardware requirements (e.g.\ head-mounted cameras and GPS antennae)~\cite{golledge2004stated,yusif2016older,arditi2013user}.
+Indeed, the traditional walking cane remains the standard tool for navigation and obstacle avoidance. 
+With the ActiVis system, we aim to tackle the issue of cost, user-acceptance and usability by using only off-the-shelf devices and hardware, as far as possible, and avoiding any special requirements.
+Following this design principle, we adopted a hand-held mobile device and a discreet headset to avoid the person standing out in public places while using the system.
+The ActiVis system is based on a concept proposed in~\cite{lock2017portable,lock2019active}, using a Google Tango device that is able to localise itself in real-time and has the benefit of a compact, familiar form-factor, which will help to overcome the hurdle of user-acceptance and usability.
 
 A set of bone-conduction headphones is used as the audio transmission medium.
-These headphones sit on a user's cheekbones and conduct the audio signals through the skull into the inner ear, instead of through the outer ear like typical over-ear headphones do. 
-This has the benefit of not denying the user access to ambient sounds and noise, which could prove dangerous to a person with limited vision that relies on sound to detect oncoming vehicles and people, for example~\cite{lichtenstein2012headphone}.
-Open-back headphones that allow ambient noise through were also considered, but the AfterShockz headphones pictured in \cref{fig:tango-headphone} were ultimately selected since they do not interfere at all with incoming sound (open-back headphones are better than closed back ones in this regard, but still muffle incoming sound) and are also more discreet than the larger over-ear headphone models.
+These headphones sit on a user's cheekbones and conduct the audio signals through the skull into the inner ear, instead of through the outer part like typical over-ear headphones. 
+This has the benefit of allowing the user access to ambient sounds and noise, so a person with limited vision can still relies on sound to detect, for example, oncoming vehicles and people~\cite{lichtenstein2012headphone}.
+Alternative solutions, such as open-back headphones, were also considered.
+These allow ambient noise through, but they still filter the incoming sound.
+The bone-conduction AfterShockz headphones (\cref{fig:tango-headphone}) were ultimately selected since they do not interfere on other sounds and are also more discreet than the larger over-ear headphones.
 
@@
 When the modified audio signal enters the inner ear canal, the human brain is able to analyse the frequency response and accurately determine the position of the sound source on the median plane. 
-The distance to the source is simply derived as the intensity, or volume, of the source, i.e.\ a louder sound would appear closer to the user than a softer one does.
+The distance to the source is simply derived as the intensity, or volume, of the source, i.e.\ a louder sound would appear closer to the user than a softer one.
 
 When an audio signal is transmitted via a set of speakers or headsets, it can be transformed with an HRTF to mimic the characteristics of a natural sound source before it is transmitted, tricking the brain into believing a sound is located at some arbitrary position.
-An HRTF is a mathematical function that simulates the response signal of a human head and is derived by capturing key characteristics that affect the monaural and binaural responses, such as the user's hearing levels and head size, etc.
+An HRTF is a mathematical function that simulates the response signal of a human head and is derived by capturing key characteristics that affect the monaural and binaural responses, such as the user's hearing levels and head size.
 Since hearing responses are unique amongst different users, the best results would be observed if each user had their own customised HRTF.\
@@
 
-The guidance target is presented to the user in terms of pan and elevation angles, indicating the angle adjustments that are required to point the device camera at the target location, as shown in \cref{fig:cam-coords}.
+The guidance information is presented to the user in terms of pan and elevation angles, indicating the angular adjustments required to point the device camera at the target location, as shown in \cref{fig:cam-coords}.
 Spatialised audio signals are well-suited to the task, displaying similar levels of performance to vocal feedback, but with less cognitive load and higher resolution~\cite{klatzky2006cognitive}.
 However, given the previously discussed limitations of bone-conduction and spatialised audio, we propose a simple linear adjustment to the signal's pitch as a function of the elevation angle. 
-The pan angle can be conveyed by transforming the audio signal with an HRTF, and indeed it has been found that this dimension is unaffected by using bone-conducting headphones\cite{schonstein2008comparison,macdonald2006spatial,stanley2006lateralization}.
+The pan angle, instead, can be conveyed by transforming the audio signal with an HRTF, and indeed it has been found that this dimension is unaffected by using bone-conducting headphones\cite{schonstein2008comparison,macdonald2006spatial,stanley2006lateralization}.
 
@@
 The ITD is the perceived time delay between the signal reaching both ears, while the ILD is the perceived volume difference in the signal.
-For example, a sound that comes from the individual's right, will hit the right ear first with a slightly higher volume.
+For example, a sound that comes from the individual's right will hit the right ear first with a slightly higher volume.
 
-In this work, a pure sinusoidal wave refreshed at \SI{10}{\hertz} was used. 
-People typically have trouble localising a pure tone devoid of a rich spectral signature.
-However, the ITD and ILD is independent of the tone's spectral make-up and the elevation angle is given through a different mechanism and a pure sine wave is therefore sufficient to convey the target's pan angle.
+In this work, a pure \SI{10}{\hertz} sinusoidal wave was used. 
+People typically have trouble localising a pure tone without a sufficiently rich spectral signature.
+However, the ITD and ILD are independent of the tone's spectral make-up, while the elevation angle is given through a different mechanism.
+Therefore, a pure sine wave is suitable to convey the target's pan angle.
 To transform and spatialise the audio signal, we used OpenAL's default HRTF, based on the MIT's KEMAR dataset~\cite{hiebert2005openal}, that uses the user and targets' positions as input and outputs a transformed audio signal.

### 2019-07-15 paper.tex
Updated figures and numbers

@@
   \caption{A summary of the results collected from the experiment. }\label{tab:results}
-  \begin{tabular}{p{1cm}p{2cm}p{4cm}c}
+  \begin{tabular}{p{1cm}p{2cm}p{2.2cm}@{\hskip 1cm}p{3cm}}
     \toprule
-    &           & Mean absolute error [rad]  & Pearson Correlation score \\ \midrule
-    \multirow{2}{*}{\textit{G1}} & Pan       & $0.25\pm0.24$ & $0.83\;(p<0.001)$ \\
-				 & Elevation & $0.37\pm0.28$ & $0.44\;(p<0.001)$ \\ \midrule
-    \multirow{2}{*}{\textit{G2}} & Pan       & $0.38\pm0.21$ & $0.35\;(p<0.001)$ \\
-				 & Elevation & $0.51\pm0.32$ & $0.40\;(p=0.001)$ \\
+    &           & Mean absolute error [rad]  & Spearman Correlation score \\ \midrule
+    \multirow{2}{*}{\textit{G1}} & Pan       & $0.23\pm0.24$ & $0.88\;(p<0.001)$ \\
+				 & Elevation & $0.25\pm0.28$ & $0.64\;(p<0.001)$ \\ \midrule
+    \multirow{2}{*}{\textit{G2}} & Pan       & $0.45\pm0.21$ & $0.41\;(p<0.001)$ \\
+				 & Elevation & $0.36\pm0.32$ & $0.61\;(p=0.001)$ \\
     \bottomrule
@@
 Indeed, the majority of samples fall within even the more conservative latter range. 
-However, the participants with vision impairments in group \textit{G2}  demonstrate a wider spread in error data and higher average error than \textit{G1} (0.25 vs.
-0.37 radians, Kruskal-Wallis test $p<0.001$).
+However, the participants with vision impairments in group \textit{G2}  demonstrate a wider spread in error data and higher average error than \textit{G1} (0.23 vs.
+0.45 radians, Kruskal-Wallis test $p<0.001$).
 The results in~\cite{katz2011spatial,zwiers2001spatial} reported a similar trend, but an investigation with a larger sample size for \textit{G2} should be considered before definitive conclusions can be made.
@@
 The elevation estimation performance for both groups deteriorated when compared to their pan estimation results, which was somewhat expected given previous research results on human audition~\cite{barfield1997visual}.
-The mean absolute errors of the participants are \SI{0.37}{\radian} and \SI{0.51}{\radian} for groups \textit{G1} and \textit{G2} respectively.
+The mean absolute errors of the participants are \SI{0.25}{\radian} and \SI{0.36}{\radian} for groups \textit{G1} and \textit{G2} respectively.
 As with the pan estimation errors, the participants from \textit{G1} seemed more adept at estimating the correct elevation.
 %However, the difference is not as significant as for the pan, with \textit{G1} displaying a 36\% improvement over \textit{G2} in their elevation estimation vs. 42.5\% in the pan dimension. 
-Comparing these results to methods that convey elevation through an HRTF, we see an improvement of approximately 46\% and 26\% for each group~\cite{schonstein2008comparison}. 
+Comparing these results to methods that convey elevation through an HRTF, we see an improvement of approximately 63\% and 47\% for each group~\cite{schonstein2008comparison}. 
 Indeed, the performance is comparable to that of open-back and expensive in-ear headphones.
-There is a lesser improvement (11\%) with \textit{G1} when compared to the individualised and adjusted HRTFs in~\cite{stanley2006lateralization} and a 18\% reduction in performance for group \textit{G2}.
+There is a more modest improvement of 43\% in \textit{G1} and 16\% for \textit{G2} when compared to the individualised and adjusted HRTFs in~\cite{stanley2006lateralization}.
 However, the participants from~\cite{stanley2006lateralization} all had healthy eyesight, making the latter comparison invalid. 
@@
 These plots are given in \cref{fig:correlation-results} and show a reasonably positive correlation between the guidance targets and the selected targets, which indicates that the participants correctly interpreted the interface's output for the most part.
-The Pearson correlation scores for each dataset, listed in \cref{tab:results}, confirm this, as well as the negative angle bias in the elevation dimensions.
+The Spearman correlation scores for each dataset, listed in \cref{tab:results}, confirm this, as well as the negative angle bias in the elevation dimensions.
 
@@
   \includegraphics[width=1.0\columnwidth]{figures/correlation.png}
-  \caption{Boxplots of the angular errors collected during the pointing experiment. }\label{fig:correlation-results}
+  \caption{Scatter plots that show the targets' actual positions and the participants' guesses, as well as the line of best fit showing the relationship between the data.
+}\label{fig:correlation-results}
 \end{figure}

### 2019-07-15 paper.tex
Changes to figures, added new signed error figure

@@
 Similar to the results from~\cite{macdonald2006spatial,schonstein2008comparison}, we use the absolute mean errors for each dimension which allows us to compare our results to the aforementioned works'.
-These results are presented in the boxplots in \cref{fig:err-results} and summarised in \cref{tab:results}.
+These results are presented in the boxplots in \cref{fig:err-results-fig} and summarised in \cref{tab:results}.
 
-\begin{figure}[t]
+\begin{figure}
   \centering
-  \includegraphics[width=0.7\columnwidth]{figures/pan_tilt_err.png}
-  \caption{Boxplots of the angular errors collected during the pointing experiment. }\label{fig:err-results}
+  \subfloat[]{\label{fig:err-results-abs}\includegraphics[width=0.7\columnwidth]{figures/pan_tilt_err_abs.png}}
+\quad
+  \subfloat[]{\label{fig:err-results}\includegraphics[width=0.7\columnwidth]{figures/pan_tilt_err.png}}
+  \caption{Boxplots of the absolute (top) and signed (bottom) angular errors collected during the pointing experiment. }\label{fig:err-results-fig}
 \end{figure}
@@
   \caption{A summary of the results collected from the experiment. }\label{tab:results}
-  \begin{tabular}{p{1cm}p{2cm}p{2.2cm}@{\hskip 1cm}p{3cm}}
+  \begin{tabular}{p{1cm}p{2cm}p{2.2cm}@{\hskip 1cm}p{3.2cm}}
     \toprule
@@
     \multirow{2}{*}{\textit{G2}} & Pan       & $0.45\pm0.21$ & $0.41\;(p<0.001)$ \\
-				 & Elevation & $0.36\pm0.32$ & $0.61\;(p=0.001)$ \\
+				 & Elevation & $0.36\pm0.32$ & $0.61\;(p<0.001)$ \\
     \bottomrule
@@
 However, the participants from~\cite{stanley2006lateralization} all had healthy eyesight, making the latter comparison invalid. 
-Interestingly, the elevation errors for group \textit{G2} are centred around the negative angle, indicating a possible underlying bias in the participants or the interface\todo{show}.
+Interestingly, the elevation errors for group \textit{G2} are centred around the negative angle, as seen in \cref{fig:err-results}, indicating a possible underlying bias in the participants or the interface.
 A similar trend was observed in~\cite{stanley2006lateralization} and warrants an investigation with a larger participant pool to determine the underlying bias, if any.

### 2019-07-15 paper.tex
revised up to sec 3

@@
 %\title{Bone-Conduction Audio Interface to\\Guide People with Visual Impairments}
-\title{Bone-Conduction and Spatialised Audio to Guide People with Visual Impairments}
+\title{Bone-Conduction Audio Interface to\\Guide People with Visual Impairments}
 
@@
   \caption{The reference system used by the guidance interface showing the camera vector and pan and elevation angles (left) and the pitch gain function used to convey the target's elevation angle (right).
-}
+Note the logarithmic scale of the frequency axis.\tofix{can you add more horizontal lines on the graph?}}
 \end{figure}
@@
 Therefore, a pure sine wave is suitable to convey the target's pan angle.
-To transform and spatialise the audio signal, we used OpenAL's default HRTF, based on the MIT's KEMAR dataset~\cite{hiebert2005openal}, that uses the user and targets' positions as input and outputs a transformed audio signal.
+To transform and spatialise the audio signal, we used OpenAL's default HRTF, based on the MIT's KEMAR dataset~\cite{hiebert2005openal}, which uses the person and targets' positions as input, and outputs a transformed audio signal.
 
@@
 
-A generic HRTF implementation with bone-conduction headphones has trouble conveying the elevation angle of a sound source accurately~\cite{macdonald2006spatial,schonstein2008comparison}.
-To compensate for this, we convey the target's elevation angle by adjusting the tone's pitch (i.e.\ the sine wave's frequency) as a function of the elevation. 
-When the camera vector is at the target's elevation, the tone is set to neutral pitch.
-Then, when the target is above the camera vector, the pitch is increased and decreased when the target is below the camera vector.
-This high/low association scheme is used given humans' natural association of high-pitched sounds with elevated sound sources and lower-pitched sounds with source's below the individual's earline~\cite{pratt1930spatial,blauert1997spatial}.
-A logarithmic, octave and semi-tone based function is used to adjust the tone's pitch to ensure perceptible changes while keeping the timbre roughly constant~\cite{shepard1964circularity}.
+A generic HRTF implementation with bone-conduction headphones is not very effective in conveying the elevation angle of a sound source~\cite{macdonald2006spatial,schonstein2008comparison}.
+To compensate for this, we communicate the target's elevation angle by adjusting the tone's pitch (i.e.\ the sine wave's frequency) as a function of the elevation. 
+When the camera vector is at the right target's elevation, the tone pitch is set to neutral.
+When the target is above or below the camera vector, the pitch is increased or decreased, respectively.
+This high/low association scheme is motivated by humans' natural association of high-pitched sounds with elevated sound sources, and low-pitched sounds with source's below the individual's earline~\cite{pratt1930spatial,blauert1997spatial}.
+An octave- and semitone-based function is used to adjust the tone's pitch to ensure perceptible changes, while keeping the timbre roughly constant~\cite{shepard1964circularity}.
 The pitch is updated at a rate of \SI{10}{\hertz} and changes as the user moves the device.
@@
 The pitch is changed as a linear function of the elevation angle and the gradient is determined by setting the angle and pitch limits.
-For this work we only consider a \SI{180}{\degree} field of view in front of the user and limit the elevation angle to a range of $[-\frac{\pi}{2}, \frac{\pi}{2}]$.
+For this work, we only consider a \SI{180}{\degree} field of view in front of the user and limit the elevation angle to a range of \SI{\pm90}{\degree}, or $[-\frac{\pi}{2}, \frac{\pi}{2}]$.
 The pitch limits are set at some integer number of octaves above and below the neutral, on-elevation pitch.
-After practical tests with the interface, we set the neutral pitch to \SI{512}{\hertz}, which is comfortably audible and allows for a large number of audible octave limits to be selected.
-For this work, we set the pitch limits to 2 octaves away from the neutral pitch, giving frequency limits of [\SI{128}{\hertz}, \SI{2048}{\hertz}].
+After practical tests with the interface, we set the neutral pitch to \SI{512}{\hertz}, which is comfortably audible and allows for a large number of suitable octave limits to be selected.
+We set the pitch limits to 2 octaves away from the neutral pitch, giving frequency limits of [\SI{128}{\hertz}, \SI{2048}{\hertz}].
 The linear function is visualised in \cref{fig:pitch-gain}.
+
+\begin{figure}[t]
+  \centering
+  \includegraphics[clip=true, trim=0 120 80 50, width=0.9\columnwidth]{figures/pipeline.pdf}
+  \caption{A diagram of the individual system components and their communication pipelines. $F$ indicates a feedback signal and $P$ a pose signal. }\label{fig:pipeline}
+\end{figure}
 
@@
 
-A diagram of the experimental system pipeline is shown in \cref{fig:pipeline}, where the arrows indicate the direction of the flow of information.
+A diagram of the experimental system pipeline is shown in \cref{fig:pipeline}, where the arrows indicate the direction of the information flow.
 When the user taps the Tango's screen, a new virtual target is generated and its coordinates are sent to the audio generation module, along with the device's current position and orientation.
-The audio generator then produces a tone based on the difference between the device and target's positions and sends it to the audio output channel, which plays it back to the user.
-The WiFi recording module is constantly monitoring the different parameter values of the device and target's positions, as well as the system's output, and records it to a remotely stored datafile.
+The audio generator then produces a tone based on the difference between the device and the target's positions. The tone is sent to the audio output channel, which plays it back to the user.
+A WiFi recording module is constantly monitoring the different values of the device's parameters and of the target's position, as well as the system's output, recording everything to a remotely stored datafile.
 
-\begin{figure}
-  \centering
-  \includegraphics[clip=true, trim=0 120 80 50, width=0.8\columnwidth]{figures/pipeline.pdf}
-  \caption{A diagram of the individual system components and their communication pipelines. $F$ indicates a feedback signal and $P$ a pose signal. }\label{fig:pipeline}
-\end{figure}
-
-\section{Experiment and Results}\label{sec:experiment-and-results}
+\section{Experiments and Results}\label{sec:experiment-and-results}
 

### 2019-07-15 paper.tex
Added Asus photo

@@
 
-%\title{Bone-Conduction Audio Interface to\\Guide People with Visual Impairments}
 \title{Bone-Conduction Audio Interface to\\Guide People with Visual Impairments}
@@
 
-The ActiVis project aims to deliver a mobile guidance system, pictured in \cref{fig:tango-headphone}, that will ultimately be able to guide a user with vision impairments on the last leg of their journey, i.e.\ the so-called `last 10-yard problem'.
-The prototype platform is based on a Google Project Tango\footnote{\url{https://en.wikipedia.org/wiki/Tango\_(platform)}} device that embeds a color camera and provides access to powerful real-time localisation (through IMU measurements and landmark tracking) and image-processing facilities.
+The ActiVis project aims to deliver a mobile guidance system, pictured in \cref{fig:tango} and \cref{fig:tango-headphone}, that will ultimately be able to guide a user with vision impairments on the last leg of their journey, i.e.\ the so-called `last 10-yard problem'.
+The prototype platform is based on a Google Project Tango\footnote{\url{https://en.wikipedia.org/wiki/Tango\_(platform)}} device that embeds a colour camera and provides access to powerful real-time localisation (through IMU measurements and landmark tracking) and image-processing facilities.
 It also provides access to Android's full range of interface tools and I/O options.
@@
   \centering
+  \subfloat[]{\label{fig:tango}\includegraphics[width=0.4\columnwidth]{figures/tango_asus.jpeg}}
+~
   \subfloat[]{\label{fig:tango-headphone}\includegraphics[width=0.4\columnwidth]{figures/tango_headphone.png}}
 ~
-  \subfloat[]{\label{fig:participant}\includegraphics[width=0.25\columnwidth]{figures/vi_participant.pdf}}
-  \caption{Pictures of the Tango device and bone-conduction headphones the system is based on (left) and both in use in an experiment (right).\tofix{add a picture of the Tango phone}}
+  \subfloat[]{\label{fig:participant}\includegraphics[width=0.2\columnwidth]{figures/vi_participant.pdf}}
+  \caption{Pictures of the latest Tango mobile phone model (left), the Tango tablet and bone-conducting headphones that were used in this work (centre) and both in use in an experiment (right).}
 \end{figure}
@@
 Furthermore, haptic and vocal feedback can become a cognitive burden, particularly where high resolution guidance, which can exceed the bandwidth of these particular senses, is required.
-Indeed, participants report that they would prefer control over the vocal feedback channel and trigger guidance instructions, instead of being given constant guidance instructions.\tofix{reference?}
+Indeed, participants report that they would prefer control over the vocal feedback channel and trigger guidance instructions, instead of being given constant guidance instructions~\cite{arditi2013user}.
 Simple audio tones are less affected by these bandwidth and hardware limitations, but they can potentially fatigue the user's senses if too unpleasant.
@@
   \caption{The reference system used by the guidance interface showing the camera vector and pan and elevation angles (left) and the pitch gain function used to convey the target's elevation angle (right).
-Note the logarithmic scale of the frequency axis.\tofix{can you add more horizontal lines on the graph?}}
+Note the logarithmic scale of the frequency axis.}
 \end{figure}

### 2019-07-15 paper.tex

@@
   This system uses bone-conduction headphones to transmit audio signals to the user and requires an effective non-visual interface.
-  To this end, we propose a new audio-based interface that uses a spatialised signal to convey a target's position on the horizontal plane. 
-  The vertical position on the median plan is given by adjusting the tone's pitch to overcome the audio localisation limitations of bone-conduction headphones. 
+  To this end, we propose a new audio-based interface that uses a spatialised signal to convey a target's direction on the horizontal plane. 
+  The vertical direction on the median plan is given by adjusting the tone's pitch to overcome the audio localisation limitations of bone-conduction headphones. 
   This interface is validated through a set of experiments with blindfolded and visually impaired users. 
-\keywords{Human-machine interface \and vision impairment \and spatialised sound \and varying pitch \and bone-conduction}
+\keywords{Human-machine interface \and visual impairment \and spatialised sound \and varying pitch \and bone-conduction}
 \end{abstract}
@@
   \subfloat[]{\label{fig:participant}\includegraphics[width=0.2\columnwidth]{figures/vi_participant.pdf}}
-  \caption{Pictures of the latest Tango mobile phone model (left), the Tango tablet and bone-conducting headphones that were used in this work (centre) and both in use in an experiment (right).}
+  \caption{Pictures of the latest Tango mobile phone (left), the former Tango tablet and the bone-conducting headphones used in this work (centre), and a visually impaired user during an experiment (right).}
 \end{figure}
@@
 The majority of these systems rely on one or a combination of vocal~\cite{mocanu2016when,chessa2016integrated,kanwal2015navigation}, audio~\cite{schwarze2015intuitive,rodriguez2012obstacle,katz2010navig} and haptic~\cite{rivera-rubio2015assistive,lee2015rgb,xiao2015assistive} feedback media to communicate guidance instructions to the user, each with their own sets of features and limitations.
-In general, participants with vision impairments report that they prefer haptic and vocal feedback out of the three options~\cite{arditi2013user}.
+In general, participants with visual impairments report that they prefer haptic and vocal feedback out of the three options~\cite{arditi2013user}.
 However, haptic feedback systems typically have extra hardware requirements to transmit the guidance instructions.
@@
 
-Electronic navigation aids and guidance systems for people with vision impairments often have acceptability and usability problems due to limitations in academic research and prototypes: prohibitive costs, unfriendly user interfaces, significant hardware requirements (e.g.\ head-mounted cameras and GPS antennae)~\cite{golledge2004stated,yusif2016older,arditi2013user}.
+Electronic navigation aids and guidance systems for people with visual impairments often have acceptability and usability problems due to limitations in academic research and prototypes: prohibitive costs, unfriendly user interfaces, significant hardware requirements (e.g.\ head-mounted cameras and GPS antennae)~\cite{golledge2004stated,yusif2016older,arditi2013user}.
 Indeed, the traditional walking cane remains the standard tool for navigation and obstacle avoidance. 
@@
 
-To test the interface's effectiveness at directing a user towards a target area, a set of experiments were conducted with a pointing task that is set up to capture the difference between the targets' actual position and the positions the participants' perceived them to be.
-The participants are given a Tango tablet device running an app written for the experiment that generates a set of virtual targets and presents them to each participants one at a time. 
-The targets are generated at a constant \SI{2}{\meter} from the participant and their the pan and elevation angles are uniformly generated across 4 quadrants to avoid clustering.
+To test the interface's effectiveness at guiding the user in a pointing task, a set of experiments were conducted to capture the difference between the targets' actual direction and the directions the participants' perceived them to be.
+The participants are given a Tango device running an app written for the experiment that generates a set of virtual targets and presents them to each participants, one at a time. 
+The targets are generated at a constant distance from the participant and their pan and elevation angles are uniformly generated across the four quadrants of the pan-elevation plane to avoid clustering.
 Each target's angular position is communicated to the participant through the audio interface, the output of which is adjusted in real-time as the participant points the device around. 
-When a participant was confident that they were on-target, i.e.\ hearing the audio front-on at \SI{512}{\hertz}, they tapped the screen, marking the location and generating the next target.
-The targets' positions are all set relative to the Tango device's coordinate system which is tracked using the Tango hardware and localisation API.\
+When a participant was confident that the device was on-target, i.e.~hearing the audio front-on at \SI{512}{\hertz}, they tapped the screen, marking the location and generating the next target.
+The targets' positions are all set relative to the device's coordinate system, which is tracked using the Tango hardware and localisation API.\
 A total of 28 targets were generated per participant.
 
-2 groups of participants were recruited for the experiments. 
-Group \textit{G1} consisted of 10 young adults with normal eyesight that were blindfolded for the experiments, and group \textit{G2} contains 4 people with severe vision impairments. 
-Both groups were given some time with the system before the experiments commenced to familiarise themselves with the system, the audio signal's behaviour and the \SI{512}{\hertz} on-level tone. 
-To minimise the speed/accuracy biases in the results, we asked the participants to focus on finding the targets as well as they could without worrying about the time it took.
+Two groups of participants were recruited for the experiments. 
+Group \textit{G1} consisted of 10 young adults with normal eyesight who were blindfolded for the experiments, and group \textit{G2} contained 4 people with severe visual impairments. 
+Both groups were given some time before the experiments to familiarise with the system, the audio signal's behaviour and the \SI{512}{\hertz} on-level tone. 
+To minimise the speed/accuracy biases in the results, we asked the participants to focus on finding the targets, without worrying about the time it took.
 
@@
 
-The data collected during the experiment were grouped together by participant group and analysed.
-Similar to the results from~\cite{macdonald2006spatial,schonstein2008comparison}, we use the absolute mean errors for each dimension which allows us to compare our results to the aforementioned works'.
+The data collected during the experiment were clustered together by participant groups and analysed.
+Similar to the results from~\cite{macdonald2006spatial,schonstein2008comparison}, we use the absolute mean errors for each dimension, which allows us to compare our results to the aforementioned works'.
 These results are presented in the boxplots in \cref{fig:err-results-fig} and summarised in \cref{tab:results}.
@@
 The average pan error from group \textit{G1} falls well within the ranges observed in~\cite{macdonald2006spatial,schonstein2008comparison} of $[0.16, 0.38]$ radians and $[0.17, 0.26]$ radians respectively.
-Indeed, the majority of samples fall within even the more conservative latter range. 
-However, the participants with vision impairments in group \textit{G2}  demonstrate a wider spread in error data and higher average error than \textit{G1} (0.23 vs.
+Indeed, the majority of samples fall even within the latter, more conservative range. 
+However, the participants with visual impairments in group \textit{G2}  demonstrate a wider spread in error data and higher average error than \textit{G1} (0.23 vs.
 0.45 radians, Kruskal-Wallis test $p<0.001$).
-The results in~\cite{katz2011spatial,zwiers2001spatial} reported a similar trend, but an investigation with a larger sample size for \textit{G2} should be considered before definitive conclusions can be made.
+The results in~\cite{katz2011spatial,zwiers2001spatial} reported a similar trend, although an investigation with a larger sample size for \textit{G2} should be considered before definitive conclusions can be made.
 
-The elevation estimation performance for both groups deteriorated when compared to their pan estimation results, which was somewhat expected given previous research results on human audition~\cite{barfield1997visual}.
-The mean absolute errors of the participants are \SI{0.25}{\radian} and \SI{0.36}{\radian} for groups \textit{G1} and \textit{G2} respectively.
+The elevation estimation performance for both groups deteriorated when compared to their pan estimation results, which was somewhat expected given previous experimental results on human audition~\cite{barfield1997visual}.
+The mean absolute errors of the participants are \SI{0.25}{\radian} and \SI{0.36}{\radian} for groups \textit{G1} and \textit{G2}, respectively.
 As with the pan estimation errors, the participants from \textit{G1} seemed more adept at estimating the correct elevation.
@@
 There is a more modest improvement of 43\% in \textit{G1} and 16\% for \textit{G2} when compared to the individualised and adjusted HRTFs in~\cite{stanley2006lateralization}.
-However, the participants from~\cite{stanley2006lateralization} all had healthy eyesight, making the latter comparison invalid. 
-Interestingly, the elevation errors for group \textit{G2} are centred around the negative angle, as seen in \cref{fig:err-results}, indicating a possible underlying bias in the participants or the interface.
-A similar trend was observed in~\cite{stanley2006lateralization} and warrants an investigation with a larger participant pool to determine the underlying bias, if any.
+However, all the participants in~\cite{stanley2006lateralization} had healthy eyesight, making the latter comparison invalid. 
+Interestingly, the elevation errors for group \textit{G2} are centred around the negative angle, as shown in \cref{fig:err-results}, indicating a possible underlying bias either in the participants or the interface.
+A similar trend was observed in~\cite{stanley2006lateralization}. Further investigation with a larger participant pool is needed to determine this underlying bias, if any.
 
-The error data was used to generate scatter plots that show the correlation between the position that the guidance system presented the participants and the position the participants thought the targets were.
-These plots are given in \cref{fig:correlation-results} and show a reasonably positive correlation between the guidance targets and the selected targets, which indicates that the participants correctly interpreted the interface's output for the most part.
+The error data was used to generate scatter plots that show the correlation between the target's direction presented by the guidance system to the participants and the one they thought the targets were.
+These plots are given in \cref{fig:correlation-results} and show a reasonably positive correlation between the guidance and the selected targets, indicating that the participants correctly interpreted the interface's output in most of the cases.
 The Spearman correlation scores for each dataset, listed in \cref{tab:results}, confirm this, as well as the negative angle bias in the elevation dimensions.
@@
 For example, will adjusting the elevation function's gradient make it easier to guess the target's elevation and what is the performance difference between participants with limited and healthy eyesight?
-Furthermore, experiments with a large group of people with vision impairments will provide an opportunity to gauge their opinion on the interface and identify possible avenues of improvements to user friendliness.
+Furthermore, experiments with a large group of people with visual impairments will provide an opportunity to gauge their opinion on the interface and identify possible avenues of improvements to user friendliness.
 Incorporating a progressive co-adaptation mechanism into the system, as suggested in~\cite{gallina2015progressive}, can also be considered to improve the overall user-friendliness and guidance performance.

### 2019-07-15 paper.tex
Proofread to before experiments

--- 2019-07-15 paper.tex
@@
   The vertical position on the median plan is given by adjusting the tone's pitch to overcome the audio localisation limitations of bone-conduction headphones. 
-  This interface is validated through a set of experiments with blindfolded and visually impaired users. 
+  This interface is validated through a set of experiments with blindfolded and visually impaired participants. 
 \keywords{Human-machine interface \and vision impairment \and spatialised sound \and varying pitch \and bone-conduction}
@@
 
-The ActiVis project aims to deliver a mobile guidance system, pictured in \cref{fig:tango} and \cref{fig:tango-headphone}, that will ultimately be able to guide a user with vision impairments on the last leg of their journey, i.e.\ the so-called `last 10-yard problem'.
-The prototype platform is based on a Google Project Tango\footnote{\url{https://en.wikipedia.org/wiki/Tango\_(platform)}} device that embeds a colour camera and provides access to powerful real-time localisation (through IMU measurements and landmark tracking) and image-processing facilities.
+The ActiVis project aims to deliver a mobile guidance system that will ultimately be able to guide a user with vision impairments on the last leg of their journey, i.e.\ the so-called `last 10-yard problem'.
+The prototype platform is based on a Google Project Tango\footnote{\url{https://en.wikipedia.org/wiki/Tango\_(platform)}} device, pictured in \cref{fig:tango} and \cref{fig:tango-headphone}, that embeds a colour camera and provides access to powerful real-time localisation (through IMU measurements and landmark tracking) and image-processing facilities.
 It also provides access to Android's full range of interface tools and I/O options. 
@@
 
-Electronic navigation aids and guidance systems for people with vision impairments often have acceptability and usability problems due to limitations in academic research and prototypes: prohibitive costs, unfriendly user interfaces, significant hardware requirements (e.g.\ head-mounted cameras and GPS antennae)~\cite{golledge2004stated,yusif2016older,arditi2013user}.
+Electronic navigation aids and guidance systems for people with vision impairments often have acceptability and usability problems due to limitations in academic research and prototypes, such as prohibitive costs, unfriendly user interfaces and significant hardware requirements (e.g.\ head-mounted cameras and GPS antennae)~\cite{golledge2004stated,yusif2016older,arditi2013user}.
 Indeed, the traditional walking cane remains the standard tool for navigation and obstacle avoidance. 
@@
 Following this design principle, we adopted a hand-held mobile device and a discreet headset to avoid the person standing out in public places while using the system.
-The ActiVis system is based on a concept proposed in~\cite{lock2017portable,lock2019active}, using a Google Tango device that is able to localise itself in real-time and has the benefit of a compact, familiar form-factor, which will help to overcome the hurdle of user-acceptance and usability.
+The ActiVis system is based on a concept proposed in~\cite{lock2017portable,lock2019active} that uses a Google Tango device that is able to localise itself in real-time and has the benefit of a compact, familiar form-factor, which will help to overcome the hurdle of user-acceptance and usability.
 
@@
 These allow ambient noise through, but they still filter the incoming sound.
-The bone-conduction AfterShockz headphones (\cref{fig:tango-headphone}) were ultimately selected since they do not interfere on other sounds and are also more discreet than the larger over-ear headphones.
+The AfterShockz headphones (\cref{fig:tango-headphone}) were ultimately selected since they do not interfere on other sounds and are also more discreet than the larger over-ear headphones.
 
@@
 
-When an audio signal is transmitted via a set of speakers or headsets, it can be transformed with an HRTF to mimic the characteristics of a natural sound source before it is transmitted, tricking the brain into believing a sound is located at some arbitrary position.
+When an audio signal is transmitted via a set of speakers or headphones, it can be transformed with an HRTF to mimic the characteristics of a natural sound source before it is transmitted, tricking the brain into believing a sound is located at some arbitrary position.
 An HRTF is a mathematical function that simulates the response signal of a human head and is derived by capturing key characteristics that affect the monaural and binaural responses, such as the user's hearing levels and head size.
@@
 
-In this work, a pure \SI{10}{\hertz} sinusoidal wave was used. 
+In this work, a pure sinusoidal wave was used. 
 People typically have trouble localising a pure tone without a sufficiently rich spectral signature.
@@
 To compensate for this, we communicate the target's elevation angle by adjusting the tone's pitch (i.e.\ the sine wave's frequency) as a function of the elevation. 
-When the camera vector is at the right target's elevation, the tone pitch is set to neutral.
+When the camera vector is at the correct elevation, the tone pitch is set to neutral.
 When the target is above or below the camera vector, the pitch is increased or decreased, respectively.

### 2019-07-15 paper.tex

--- 2019-07-15 paper.tex
@@
   \subfloat[]{\label{fig:participant}\includegraphics[width=0.2\columnwidth]{figures/vi_participant.pdf}}
-  \caption{Pictures of the latest Tango mobile phone (left), the former Tango tablet and the bone-conducting headphones used in this work (centre), and a visually impaired user during an experiment (right).}
+  \caption{Pictures of the latest Tango mobile phone (left), the former Tango tablet and the bone-conducting headphones used in this work (centre), and a participant with visual impairments during an experiment (right).}
 \end{figure}

### 2019-07-15 paper.tex

@@
   This system uses bone-conduction headphones to transmit audio signals to the user and requires an effective non-visual interface.
-  To this end, we propose a new audio-based interface that uses a spatialised signal to convey a target's direction on the horizontal plane. 
-  The vertical direction on the median plan is given by adjusting the tone's pitch to overcome the audio localisation limitations of bone-conduction headphones. 
-  This interface is validated through a set of experiments with blindfolded and visually impaired users. 
-\keywords{Human-machine interface \and visual impairment \and spatialised sound \and varying pitch \and bone-conduction}
+  To this end, we propose a new audio-based interface that uses a spatialised signal to convey a target's position on the horizontal plane. 
+  The vertical position on the median plan is given by adjusting the tone's pitch to overcome the audio localisation limitations of bone-conduction headphones. 
+  This interface is validated through a set of experiments with blindfolded and visually impaired participants. 
+\keywords{Human-machine interface \and vision impairment \and spatialised sound \and varying pitch \and bone-conduction}
 \end{abstract}
@@
 
-Electronic navigation aids and guidance systems for people with visual impairments often have acceptability and usability problems due to limitations in academic research and prototypes: prohibitive costs, unfriendly user interfaces, significant hardware requirements (e.g.\ head-mounted cameras and GPS antennae)~\cite{golledge2004stated,yusif2016older,arditi2013user}.
+Electronic navigation aids and guidance systems for people with vision impairments often have acceptability and usability problems due to limitations in academic research and prototypes, such as prohibitive costs, unfriendly user interfaces and significant hardware requirements (e.g.\ head-mounted cameras and GPS antennae)~\cite{golledge2004stated,yusif2016older,arditi2013user}.
 Indeed, the traditional walking cane remains the standard tool for navigation and obstacle avoidance. 

### 2019-07-15 paper.tex
Fixed legacy sentance

@@
 However, all the participants in~\cite{stanley2006lateralization} had healthy eyesight, making the latter comparison invalid. 
-Interestingly, the elevation errors for group \textit{G2} are centred around the negative angle, as shown in \cref{fig:err-results}, indicating a possible underlying bias either in the participants or the interface.
+Interestingly, the elevation errors for both groups are centred around the negative angle, as shown in \cref{fig:err-results}, indicating a possible underlying bias either in the participants or the interface.
 A similar trend was observed in~\cite{stanley2006lateralization}. Further investigation with a larger participant pool is needed to determine this underlying bias, if any.

### 2019-07-15 paper.tex
Fixed clunky sentence in conclusion

@@
 
-In this work, we proposed an audio interface that conveys a target's pan and elevation angles through a spatialised audio tone that adjusts the tone's pitch to convey the elevation in compensate for the shortcomings of bone-conduction headsets in transmitting the elevation component of a spatialised signal.
+In this work, we proposed an audio interface that conveys a target's pan and elevation angles through a spatialised audio tone.
+To compensate for the shortcomings of bone-conduction headsets and standard HRTFs, the elevation is given by adjusting the tone's pitch.
 Results generated from a simple pointing experiment are encouraging and improves performance over purely spatialised signals transmitted via bone-conduction.

### 2019-07-15 paper.tex

--- 2019-07-15 paper.tex
@@
 
-In this work, we proposed an audio interface that conveys a target's pan and elevation angles through a spatialised audio tone that adjusts the tone's pitch to convey the elevation in compensate for the shortcomings of bone-conduction headsets in transmitting the elevation component of a spatialised signal.
-Results generated from a simple pointing experiment are encouraging and improves performance over purely spatialised signals transmitted via bone-conduction.
-Furthermore, the limitations of these headphones were largely negated by conveying the elevation angle of the target by adjusting the tone's pitch, with pan and elevation errors comparable to that from normal over-ear headphones.
+In this work, we proposed a new audio interface that conveys a target's pan and elevation angles through a spatialised audio tone.
+The tone's pitch is adjusted to convey the elevation and compensate for the shortcomings of bone-conduction headsets in transmitting the elevation component of a spatialised signal.
+The results obtained from a simple pointing experiment are encouraging as they show improved performance over purely spatialised signals transmitted via bone-conduction headphones.
+Furthermore, the limitations of the latter were largely overcome by communicating the target's elevation with the tone's pitch, achieving pan and elevation errors comparable to that of normal over-ear headphones.
 
-Future work will involve a more in-depth investigation with a larger participant pool to determine the factors that affect performance and enable a more detailed statistical analysis.
-For example, will adjusting the elevation function's gradient make it easier to guess the target's elevation and what is the performance difference between participants with limited and healthy eyesight?
-Furthermore, experiments with a large group of people with visual impairments will provide an opportunity to gauge their opinion on the interface and identify possible avenues of improvements to user friendliness.
-Incorporating a progressive co-adaptation mechanism into the system, as suggested in~\cite{gallina2015progressive}, can also be considered to improve the overall user-friendliness and guidance performance.
+Future work should consider a more in-depth investigation with a larger participant pool and a more detailed statistical analysis.
+Interesting questions include whether adjusting the gradient of the elevation function would make it easier for the user to guess the actual target's elevation, and to what degree the performance of participants with limited and healthy eyesight differs.
+Furthermore, experiments with a large group of people with visual impairments would provide an opportunity to gauge their opinion on the interface and identify possible avenues of improvements.
+Finally, it would be interesting to incorporate a co-adaptation mechanism between human and assistive device, as suggested in~\cite{gallina2015progressive}, to improve its usability and performance over time.
 

### 2019-07-15 paper.tex

@@
 ~
-  \subfloat[]{\label{fig:tango-headphone}\includegraphics[width=0.4\columnwidth]{figures/tango_headphone.png}}
+  \subfloat[]{\label{fig:tango-headphone}\includegraphics[width=0.35\columnwidth]{figures/tango_headphone.png}}
 ~

### 2019-09-02 paper.tex
added acknowledgements
Moved picture
Minor text changes to correct grammar and legibility. No major corrections suggested by reviewrs

@@
 \usepackage[capitalise]{cleveref}
-\usepackage{todonotes}
-
-\usepackage{soul}
-\DeclareRobustCommand{\tofix}[1]{{\sethlcolor{yellow}\hl{[#1]}}}
 
@@
   This system uses bone-conduction headphones to transmit audio signals to the user and requires an effective non-visual interface.
-  To this end, we propose a new audio-based interface that uses a spatialised signal to convey a target's position on the horizontal plane. 
-  The vertical position on the median plan is given by adjusting the tone's pitch to overcome the audio localisation limitations of bone-conduction headphones. 
-  This interface is validated through a set of experiments with blindfolded and visually impaired participants. 
+  To this end, we propose an audio-based interface that uses a spatialised signal to convey a target's position on the horizontal plane. 
+  The vertical position on the median plan is given by adjusting the tone's pitch to overcome the audio localisation limitations of bone-conducting headphones. 
+  This interface was validated through a set of experiments with blindfolded and visually impaired participants. 
 \keywords{Human-machine interface \and vision impairment \and spatialised sound \and varying pitch \and bone-conduction}
@@
 In recent years, governments have spearheaded numerous initiatives to support people with disabilities and enable them to play a more active role in modern society.
-The UK's Royal National Institute of Blind People (RNIB) for example has prioritised improving access to everyday services and products, such as public transport and mobile apps~\cite{rnib2016uk}.
+The UK's Royal National Institute of Blind People for example has prioritised improving access to everyday services and products, such as public transport and mobile apps~\cite{rnib2016uk}.
 Improvements in modern computing have made it possible for new and innovative solutions to these problems to come to the fore.
@@
 
-The ActiVis project aims to deliver a mobile guidance system that will ultimately be able to guide a user with vision impairments on the last leg of their journey, i.e.\ the so-called `last 10-yard problem'.
-The prototype platform is based on a Google Project Tango\footnote{\url{https://en.wikipedia.org/wiki/Tango\_(platform)}} device, pictured in \cref{fig:tango} and \cref{fig:tango-headphone}, that embeds a colour camera and provides access to powerful real-time localisation (through IMU measurements and landmark tracking) and image-processing facilities.
+The ActiVis project aims to deliver a mobile guidance system that will ultimately be able to guide a user with vision impairments on the last leg of their journey, addressing the so-called `last 10-yards problem'.
+The prototype platform is based on a Google Project Tango\footnote{\url{https://en.wikipedia.org/wiki/Tango\_(platform)}} device, pictured in \cref{fig:tango} and \cref{fig:tango-headphone}, provides access to powerful real-time localisation (through IMU measurements and landmark tracking with embedded RGB-D camera) and image-processing facilities.
 It also provides access to Android's full range of interface tools and I/O options. 
-Furthermore, a set of bone-conduction headphones (pictured in \cref{fig:tango-headphone}), which are placed on a user's cheekbones instead of their ears and do not interfere with normal hearing, are used to transmit the audio signals to the user.
+Furthermore, a set of bone-conducting headphones (pictured in \cref{fig:tango-headphone}), which are placed on a user's cheekbones instead of their ears and do not interfere with normal hearing, are used to transmit the audio signals to the user.
 
@@
 
-Humans are able to determine the 3D position of a sound source and by exploiting this natural ability, the real-time guidance instructions can easily be interpreted without posing a significant cognitive load.
+Humans are able to determine the 3D position of a sound source and by exploiting this natural ability, the system's real-time guidance instructions can be interpreted without posing a significant cognitive load.
 A sound source can be spatialised by adjusting a tone's spectral make-up (elevation angle), time delay and level difference (pan angle), and intensity (distance).
 In our case, only the pan and elevation positions are transmitted to the user to point the camera towards a target object or visual feature.
-However, since bone-conduction headphones bypass the outer ear structure, their spectral signature cannot properly be interpreted and we therefore convey the target's elevation angle by adjusting the tone's pitch.
-A similar approach was used in~\cite{durette2008visuo}, but the study did not focus on or investigate the interface's efficacy.
+However, since bone-conducting headphones bypass the outer-ear structure, their spectral signature cannot properly be interpreted and we therefore convey the target's elevation angle by adjusting the tone's pitch.
+A similar approach was used with success in~\cite{durette2008visuo}, but the study did not focus on or investigate the interface's efficacy.
 Indeed, the work presented in this paper serves as an initial study into an interface that can effectively communicate the pan and elevation angles of a target location through a set of bone-conduction headphones using standard audio libraries and audio manipulations.
@@
 
-The rest of the paper starts by discussing previous relevant works and research in \cref{sec:prev-work}, followed by a discussion on the design and implementation of our interface in \cref{sec:interface-design}.
+The paper starts by discussing previous relevant works and research in \cref{sec:prev-work}, followed by a discussion on the design and implementation of our interface in \cref{sec:interface-design}.
 This is followed by a description of the experiments that were conducted and a discussion of their results in \cref{sec:experiment-and-results}.
@@
 Multiple mobile navigation and travel aids have been devised throughout the years, some as part of a commercial undertaking and many as part of academic research.
-The majority of these systems rely on one or a combination of vocal~\cite{mocanu2016when,chessa2016integrated,kanwal2015navigation}, audio~\cite{schwarze2015intuitive,rodriguez2012obstacle,katz2010navig} and haptic~\cite{rivera-rubio2015assistive,lee2015rgb,xiao2015assistive} feedback media to communicate guidance instructions to the user, each with their own sets of features and limitations.
-In general, participants with visual impairments report that they prefer haptic and vocal feedback out of the three options~\cite{arditi2013user}.
-However, haptic feedback systems typically have extra hardware requirements to transmit the guidance instructions.
+The majority of these systems rely on one or a combination of vocal~\cite{mocanu2016when,chessa2016integrated,kanwal2015navigation}, audio~\cite{schwarze2015intuitive,rodriguez2012obstacle,katz2010navig} and haptic~\cite{rivera-rubio2015assistive,lee2015rgb,xiao2015assistive} feedback media to communicate guidance instructions to the user, each with their own set of features and limitations.
+In general, participants with visual impairments report that they prefer haptic and vocal feedback out of the three aforementioned options~\cite{arditi2013user}.
+However, haptic feedback systems typically have extra hardware requirements to reliably transmit guidance signals.
 Furthermore, haptic and vocal feedback can become a cognitive burden, particularly where high resolution guidance, which can exceed the bandwidth of these particular senses, is required.
-Indeed, participants report that they would prefer control over the vocal feedback channel and trigger guidance instructions, instead of being given constant guidance instructions~\cite{arditi2013user}.
-Simple audio tones are less affected by these bandwidth and hardware limitations, but they can potentially fatigue the user's senses if too unpleasant.
+Indeed, participants report that they would prefer control over the vocal feedback channel and manually trigger guidance instructions instead of being given constant guidance instructions~\cite{arditi2013user}.
+Simple audio tones are less affected by bandwidth and hardware limitations, but they can potentially fatigue the user if they are too unpleasant.
 
@@
 However, other authors have found that the choice of audio transmission medium can have a significant effect on performance.
-Cheaper headphones and bone-conduction headphones generally report unfavourable results when compared to over-ear or more expensive alternatives~\cite{schonstein2008comparison,macdonald2006spatial,stanley2006lateralization}.
+Cheaper headphones and bone-conducting headphones generally report unfavourable results when compared to over-ear or more expensive alternatives~\cite{schonstein2008comparison,macdonald2006spatial,stanley2006lateralization}.
 This seems to be limited to the elevation dimension, however, and can be improved with HRTFs adjusted for the bone-conduction pathway~\cite{stanley2006lateralization}.
@@
 However, these studies focussed on their system's total effectiveness, rather than the effectiveness of the pitch adjustment mechanism alone. 
-In the next sections we want to investigate this more closely.
+In the next sections we investigate this more closely.
 
@@
 
-Electronic navigation aids and guidance systems for people with vision impairments often have acceptability and usability problems due to limitations in academic research and prototypes, such as prohibitive costs, unfriendly user interfaces and significant hardware requirements (e.g.\ head-mounted cameras and GPS antennae)~\cite{golledge2004stated,yusif2016older,arditi2013user}.
+Electronic navigation aids and guidance systems for people with vision impairments often have acceptability and usability problems due to limitations in academic research and prototypes, such as prohibitive costs, unfriendly user interfaces and significant hardware requirements, e.g.\ head-mounted cameras and GPS antennae~\cite{golledge2004stated,yusif2016older,arditi2013user}.
 Indeed, the traditional walking cane remains the standard tool for navigation and obstacle avoidance. 
-With the ActiVis system, we aim to tackle the issue of cost, user-acceptance and usability by using only off-the-shelf devices and hardware, as far as possible, and avoiding any special requirements.
+With the ActiVis system, we aim to tackle the issue of cost, user-acceptance and usability by using only off-the-shelf devices and hardware as far as possible and avoiding any special requirements.
 Following this design principle, we adopted a hand-held mobile device and a discreet headset to avoid the person standing out in public places while using the system.
-The ActiVis system is based on a concept proposed in~\cite{lock2017portable,lock2019active} that uses a Google Tango device that is able to localise itself in real-time and has the benefit of a compact, familiar form-factor, which will help to overcome the hurdle of user-acceptance and usability.
+The ActiVis system is based on a concept proposed in~\cite{lock2017portable,lock2019active} that uses a Google Tango device that is able to localise itself in real-time and has the benefit of a compact, familiar form-factor which will help overcome the hurdle of user-acceptance and usability.
 
-A set of bone-conduction headphones is used as the audio transmission medium.
-These headphones sit on a user's cheekbones and conduct the audio signals through the skull into the inner ear, instead of through the outer part like typical over-ear headphones. 
-This has the benefit of allowing the user access to ambient sounds and noise, so a person with limited vision can still relies on sound to detect, for example, oncoming vehicles and people~\cite{lichtenstein2012headphone}.
+A set of bone-conducting headphones is used as the audio transmission medium.
+These headphones sit on a user's cheekbones and conduct the audio signals through the skull into the inner ear instead of through the outer part like typical over-ear headphones do. 
+This has the benefit of giving the user access to ambient sounds and noise, so a person with limited vision can still rely on sound to detect, for example, oncoming vehicles and people~\cite{lichtenstein2012headphone}.
 Alternative solutions, such as open-back headphones, were also considered.
 These allow ambient noise through, but they still filter the incoming sound.
-The AfterShockz headphones (\cref{fig:tango-headphone}) were ultimately selected since they do not interfere on other sounds and are also more discreet than the larger over-ear headphones.
+The AfterShockz headphones (\cref{fig:tango-headphone}) were ultimately selected since they do not interfere with any other sounds and are also more discreet than the larger over-ear headphones.
 
@@
 
-Humans localise a sound source in 3 dimensions by considering cues recorded in one ear (monaural cues) and comparing cues received at both ears (binaural cues)~\cite{blauert1997spatial,blauert1969sound}.
-The binaural cues include inter-aural time and level differences (ITD and ILD respectively) that help to determine a source's location on the horizontal plane.
-Monaural cues are taken from the interaction of the sound with the human anatomy, e.g.\ head, shoulders, outer ear, before it enters the ear canal.
-%The user's anatomy acts as a filter and the direction the sound wave hits the body activates different filter resonances, attenuating or accentuating different frequency bands.
+Humans localise sound in 3 dimensions by considering cues recorded in one ear (monaural cues) and comparing cues received at both ears (binaural cues)~\cite{blauert1997spatial,blauert1969sound}.
+Binaural cues include inter-aural time and level differences (ITD and ILD respectively) that help to determine a source's location on the horizontal plane, while monaural cues are taken from the interaction of the sound with the human anatomy, e.g.\ head, shoulders, outer ear, before it enters the ear canal.
 When the modified audio signal enters the inner ear canal, the human brain is able to analyse the frequency response and accurately determine the position of the sound source on the median plane. 
@@
 
-When an audio signal is transmitted via a set of speakers or headphones, it can be transformed with an HRTF to mimic the characteristics of a natural sound source before it is transmitted, tricking the brain into believing a sound is located at some arbitrary position.
-An HRTF is a mathematical function that simulates the response signal of a human head and is derived by capturing key characteristics that affect the monaural and binaural responses, such as the user's hearing levels and head size.
+When an audio signal is transmitted via a set of speakers or headphones, it can be transformed with an HRTF to mimic the characteristics of a natural sound source, tricking the brain into believing a sound is located at some arbitrary position.
+An HRTF is a mathematical function that simulates the response signal of the human head and is derived by capturing key characteristics that affect the monaural and binaural responses, such as the user's hearing levels, ear shape and head size.
 Since hearing responses are unique amongst different users, the best results would be observed if each user had their own customised HRTF.\
-However, given the complicated process involved to capture the required user characteristics, making unique HRTFs is often an untenable solution and using average values (e.g.\ head measurements, height, etc.) have shown to produce acceptable results~\cite{gardner1995hrtf}.
+However, given the complicated process involved to capture the required user characteristics, making unique HRTFs is often an untenable solution and using average values (e.g.\ head measurements, height, etc.) produces acceptable results~\cite{gardner1995hrtf}.
 
@@
 
-The guidance information is presented to the user in terms of pan and elevation angles, indicating the angular adjustments required to point the device camera at the target location, as shown in \cref{fig:cam-coords}.
+Guidance information is presented to the user in terms of pan and elevation angles, indicating the angular adjustments required to point the device camera at the target location, as shown in \cref{fig:cam-coords}.
 Spatialised audio signals are well-suited to the task, displaying similar levels of performance to vocal feedback, but with less cognitive load and higher resolution~\cite{klatzky2006cognitive}.
 However, given the previously discussed limitations of bone-conduction and spatialised audio, we propose a simple linear adjustment to the signal's pitch as a function of the elevation angle. 
-The pan angle, instead, can be conveyed by transforming the audio signal with an HRTF, and indeed it has been found that this dimension is unaffected by using bone-conducting headphones\cite{schonstein2008comparison,macdonald2006spatial,stanley2006lateralization}.
+The pan angle, instead, can be conveyed by transforming the audio signal with an HRTF, and indeed it has been found that this dimension is unaffected by bone-conducting headphones~\cite{schonstein2008comparison,macdonald2006spatial,stanley2006lateralization}.
 
@@
 
-In this work, a pure sinusoidal wave was used. 
-People typically have trouble localising a pure tone without a sufficiently rich spectral signature.
+In this work, a pure sinusoidal wave was used, which people typically have trouble localising given its lack of a sufficiently rich spectral signature.
 However, the ITD and ILD are independent of the tone's spectral make-up, while the elevation angle is given through a different mechanism. 
 Therefore, a pure sine wave is suitable to convey the target's pan angle.
-To transform and spatialise the audio signal, we used OpenAL's default HRTF, based on the MIT's KEMAR dataset~\cite{hiebert2005openal}, which uses the person and targets' positions as input, and outputs a transformed audio signal.
+To transform and spatialise the audio signal, we used OpenAL's default HRTF, based on the MIT's KEMAR dataset~\cite{hiebert2005openal} which uses the person and targets' positions as input and outputs a transformed audio signal.
 
@@
 
-A generic HRTF implementation with bone-conduction headphones is not very effective in conveying the elevation angle of a sound source~\cite{macdonald2006spatial,schonstein2008comparison}.
-To compensate for this, we communicate the target's elevation angle by adjusting the tone's pitch (i.e.\ the sine wave's frequency) as a function of the elevation. 
-When the camera vector is at the correct elevation, the tone pitch is set to neutral.
-When the target is above or below the camera vector, the pitch is increased or decreased, respectively.
+A generic HRTF implementation with bone-conducting headphones is not very effective in conveying the elevation angle of a sound source~\cite{macdonald2006spatial,schonstein2008comparison}.
+To compensate for this, the target's elevation angle is conveyed by adjusting the tone's pitch (i.e.\ the sine wave's frequency) as a function of the elevation. 
+When the camera vector is at the correct elevation, the tone pitch is set to neutral and when the target is above or below the camera vector, the pitch is increased or decreased, respectively.
 This high/low association scheme is motivated by humans' natural association of high-pitched sounds with elevated sound sources, and low-pitched sounds with source's below the individual's earline~\cite{pratt1930spatial,blauert1997spatial}.
-An octave- and semitone-based function is used to adjust the tone's pitch to ensure perceptible changes, while keeping the timbre roughly constant~\cite{shepard1964circularity}.
+An octave- and semitone-based function is used to adjust the tone's pitch to ensure perceptible changes while keeping the timbre roughly constant~\cite{shepard1964circularity}.
 The pitch is updated at a rate of \SI{10}{\hertz} and changes as the user moves the device.
@@
 After practical tests with the interface, we set the neutral pitch to \SI{512}{\hertz}, which is comfortably audible and allows for a large number of suitable octave limits to be selected.
-We set the pitch limits to 2 octaves away from the neutral pitch, giving frequency limits of [\SI{128}{\hertz}, \SI{2048}{\hertz}].
+We set the pitch limits to 2 octaves from the neutral pitch, giving frequency limits of [\SI{128}{\hertz}, \SI{2048}{\hertz}].
 The linear function is visualised in \cref{fig:pitch-gain}.
+
+\subsection{Implementation}
+
+A diagram of the experimental system pipeline is shown in \cref{fig:pipeline}, where the arrows indicate the direction of the information flow.
+When the user taps the device screen, a new virtual target is generated and its coordinates are sent to the audio generation module, along with the device's current position and orientation.
+The audio generator then produces a tone based on the difference between the device and the target's positions.
+The tone is sent to the audio output channel, which plays it back to the user.
+A WiFi recording module is constantly monitoring the different values of the device's parameters and of the target's position, as well as the system's output, recording everything to a remotely stored datafile.
 
@@
 
-\subsection{Implementation}
-
-A diagram of the experimental system pipeline is shown in \cref{fig:pipeline}, where the arrows indicate the direction of the information flow.
-When the user taps the Tango's screen, a new virtual target is generated and its coordinates are sent to the audio generation module, along with the device's current position and orientation.
-The audio generator then produces a tone based on the difference between the device and the target's positions. The tone is sent to the audio output channel, which plays it back to the user.
-A WiFi recording module is constantly monitoring the different values of the device's parameters and of the target's position, as well as the system's output, recording everything to a remotely stored datafile.
-
 \section{Experiments and Results}\label{sec:experiment-and-results}
@@
 
-To test the interface's effectiveness at guiding the user in a pointing task, a set of experiments were conducted to capture the difference between the targets' actual direction and the directions the participants' perceived them to be.
-The participants are given a Tango device running an app written for the experiment that generates a set of virtual targets and presents them to each participants, one at a time. 
+To test the interface's effectiveness at guiding the user in a pointing task, a set of experiments were conducted to capture the difference between the targets' actual directions and the directions the participants perceived them to be.
+The participants were given a Tango device running an app written for the experiment that generates a set of virtual targets and presents them to the participants one at a time. 
 The targets are generated at a constant distance from the participant and their pan and elevation angles are uniformly generated across the four quadrants of the pan-elevation plane to avoid clustering.
 Each target's angular position is communicated to the participant through the audio interface, the output of which is adjusted in real-time as the participant points the device around. 
-When a participant was confident that the device was on-target, i.e.~hearing the audio front-on at \SI{512}{\hertz}, they tapped the screen, marking the location and generating the next target.
+When a participant was confident that the device was on-target, i.e.\ hearing the audio front-on at \SI{512}{\hertz}, they tapped the screen, marking the location and generating the next target.
 The targets' positions are all set relative to the device's coordinate system, which is tracked using the Tango hardware and localisation API.\
@@
 Two groups of participants were recruited for the experiments. 
-Group \textit{G1} consisted of 10 young adults with normal eyesight who were blindfolded for the experiments, and group \textit{G2} contained 4 people with severe visual impairments. 
-Both groups were given some time before the experiments to familiarise with the system, the audio signal's behaviour and the \SI{512}{\hertz} on-level tone. 
-To minimise the speed/accuracy biases in the results, we asked the participants to focus on finding the targets, without worrying about the time it took.
+Group \textit{G1} consisted of 10 young adults with normal eyesight who were blindfolded for the experiments, and group \textit{G2} contained 4 legally blind participants. 
+Both groups were given some time before the experiments to familiarise themselves with the system, the audio signal's behaviour and the \SI{512}{\hertz} on-level tone. 
+To minimise the speed/accuracy biases in the results, we asked the participants to focus on finding the targets without worrying about the time it took.
 
@@
 The data collected during the experiment were clustered together by participant groups and analysed.
-Similar to the results from~\cite{macdonald2006spatial,schonstein2008comparison}, we use the absolute mean errors for each dimension, which allows us to compare our results to the aforementioned works'.
+Similar to the results from~\cite{macdonald2006spatial,schonstein2008comparison}, we use the absolute mean errors for each dimension which allows us to compare our results to the aforementioned works'.
 These results are presented in the boxplots in \cref{fig:err-results-fig} and summarised in \cref{tab:results}.
@@
 0.45 radians, Kruskal-Wallis test $p<0.001$).
-The results in~\cite{katz2011spatial,zwiers2001spatial} reported a similar trend, although an investigation with a larger sample size for \textit{G2} should be considered before definitive conclusions can be made.
+The results in~\cite{katz2011spatial,zwiers2001spatial} reported a similar trend, although an investigation with a larger sample size for \textit{G2} should be considered before definitive conclusions can be drawn.
 
@@
 The mean absolute errors of the participants are \SI{0.25}{\radian} and \SI{0.36}{\radian} for groups \textit{G1} and \textit{G2}, respectively.
-As with the pan estimation errors, the participants from \textit{G1} seemed more adept at estimating the correct elevation.
-%However, the difference is not as significant as for the pan, with \textit{G1} displaying a 36\% improvement over \textit{G2} in their elevation estimation vs. 42.5\% in the pan dimension. 
+As with the pan estimation errors, the participants from \textit{G1} seem more adept at estimating the correct elevation.
 Comparing these results to methods that convey elevation through an HRTF, we see an improvement of approximately 63\% and 47\% for each group~\cite{schonstein2008comparison}. 
-Indeed, the performance is comparable to that of open-back and expensive in-ear headphones.
+Indeed, the performance is comparable to that of more expensive open-back and in-ear headphones.
 There is a more modest improvement of 43\% in \textit{G1} and 16\% for \textit{G2} when compared to the individualised and adjusted HRTFs in~\cite{stanley2006lateralization}.
-However, all the participants in~\cite{stanley2006lateralization} had healthy eyesight, making the latter comparison invalid. 
-Interestingly, the elevation errors for both groups are centred around the negative angle, as shown in \cref{fig:err-results}, indicating a possible underlying bias either in the participants or the interface.
-A similar trend was observed in~\cite{stanley2006lateralization}. Further investigation with a larger participant pool is needed to determine this underlying bias, if any.
+However, all the participants in~\cite{stanley2006lateralization} had healthy eyesight, making the latter comparison less valid. 
+Interestingly, the elevation errors for both groups are centred around a negative angle, as shown in \cref{fig:err-results}, indicating a possible underlying bias either in the participants or the interface.
+A similar trend was observed in~\cite{stanley2006lateralization}. 
+Further investigation with a larger participant pool is needed to determine this underlying bias, if any.
 
-The error data was used to generate scatter plots that show the correlation between the target's direction presented by the guidance system to the participants and the one they thought the targets were.
+The error data was used to generate scatter plots that show the correlation between the targets' actual positions and their positions as perceived by the participants.
 These plots are given in \cref{fig:correlation-results} and show a reasonably positive correlation between the guidance and the selected targets, indicating that the participants correctly interpreted the interface's output in most of the cases.
-The Spearman correlation scores for each dataset, listed in \cref{tab:results}, confirm this, as well as the negative angle bias in the elevation dimensions.
+The Spearman correlation scores for each dataset, listed in \cref{tab:results}, confirm this as well as the negative angle bias in the elevation dimensions.
 
@@
 
-In this work, we proposed a new audio interface that conveys a target's pan and elevation angles through a spatialised audio tone.
-The tone's pitch is adjusted to convey the elevation and compensate for the shortcomings of bone-conduction headsets in transmitting the elevation component of a spatialised signal.
-The results obtained from a simple pointing experiment are encouraging as they show improved performance over purely spatialised signals transmitted via bone-conduction headphones.
-Furthermore, the limitations of the latter were largely overcome by communicating the target's elevation with the tone's pitch, achieving pan and elevation errors comparable to that of normal over-ear headphones.
+In this work, we propose an audio interface that conveys a target's pan and elevation angles through a spatialised audio tone and bone-conducting headphones.
+The tone's pitch is adjusted to convey the elevation and compensate for the shortcomings of bone-conducting headsets in transmitting the elevation component of a spatialised signal.
+The results obtained from a simple pointing experiment are encouraging as they show improved performance over purely spatialised signals transmitted via bone-conducting headphones.
+Furthermore, the limitations of the latter were largely overcome by communicating the target's elevation with the tone's pitch, achieving pan and elevation errors comparable to that of over-ear headphones.
 
 Future work should consider a more in-depth investigation with a larger participant pool and a more detailed statistical analysis.
-Interesting questions include whether adjusting the gradient of the elevation function would make it easier for the user to guess the actual target's elevation, and to what degree the performance of participants with limited and healthy eyesight differs.
+Interesting questions include whether adjusting the gradient of the elevation function would make it easier for the user to guess the actual target's elevation and to what degree the performance of participants with limited and healthy eyesight differs.
 Furthermore, experiments with a large group of people with visual impairments would provide an opportunity to gauge their opinion on the interface and identify possible avenues of improvements.
 Finally, it would be interesting to incorporate a co-adaptation mechanism between human and assistive device, as suggested in~\cite{gallina2015progressive}, to improve its usability and performance over time.
+
+\paragraph{{\bf Acknowledgements.}}\label{sec:acknowledge}
+This research is partly supported by a Google Faculty Research Award.
+We would like to thank the Voluntary Centre Services UK for their help in facilitating the experiments with people with limited vision.
 

### 2019-09-11 paper.tex
Moved figure

--- 2019-07-15 paper.tex
@@
   The ActiVis\footnote{\url{https://lcas.github.io/ActiVis/}} project's aim is to build a mobile guidance aid to help people with limited vision find objects in an unknown environment.
-  This system uses bone-conduction headphones to transmit audio signals to the user and requires an effective non-visual interface.
+  This system uses bone-conducting headphones to transmit audio signals to the user and requires an effective non-visual interface.
   To this end, we propose a new audio-based interface that uses a spatialised signal to convey a target's position on the horizontal plane. 
-  The vertical position on the median plan is given by adjusting the tone's pitch to overcome the audio localisation limitations of bone-conduction headphones. 
+  The vertical position on the median plan is given by adjusting the tone's pitch to overcome the audio localisation limitations of bone-conducting headphones. 
   This interface is validated through a set of experiments with blindfolded and visually impaired participants. 
@@
 It also provides access to Android's full range of interface tools and I/O options. 
-Furthermore, a set of bone-conduction headphones (pictured in \cref{fig:tango-headphone}), which are placed on a user's cheekbones instead of their ears and do not interfere with normal hearing, are used to transmit the audio signals to the user.
+Furthermore, a set of bone-conducting headphones (pictured in \cref{fig:tango-headphone}), which are placed on a user's cheekbones instead of their ears and do not interfere with normal hearing, are used to transmit the audio signals to the user.
 
@@
 ~
-  \subfloat[]{\label{fig:tango-headphone}\includegraphics[width=0.35\columnwidth]{figures/tango_headphone.png}}
+  \subfloat[]{\label{fig:tango-headphone}\includegraphics[width=0.3\columnwidth]{figures/tango_headphone.png}}
 ~
@@
 In our case, only the pan and elevation positions are transmitted to the user to point the camera towards a target object or visual feature.
-However, since bone-conduction headphones bypass the outer ear structure, their spectral signature cannot properly be interpreted and we therefore convey the target's elevation angle by adjusting the tone's pitch.
+However, since bone-conducting headphones bypass the outer ear structure, their spectral signature cannot properly be interpreted and we therefore convey the target's elevation angle by adjusting the tone's pitch.
 A similar approach was used in~\cite{durette2008visuo}, but the study did not focus on or investigate the interface's efficacy.
-Indeed, the work presented in this paper serves as an initial study into an interface that can effectively communicate the pan and elevation angles of a target location through a set of bone-conduction headphones using standard audio libraries and audio manipulations.
+Indeed, the work presented in this paper serves as an initial study into an interface that can effectively communicate the pan and elevation angles of a target location through a set of bone-conducting headphones using standard audio libraries and audio manipulations.
 Furthermore, we demonstrate this approach's effectiveness through a simple experiment.
@@
 However, other authors have found that the choice of audio transmission medium can have a significant effect on performance.
-Cheaper headphones and bone-conduction headphones generally report unfavourable results when compared to over-ear or more expensive alternatives~\cite{schonstein2008comparison,macdonald2006spatial,stanley2006lateralization}.
-This seems to be limited to the elevation dimension, however, and can be improved with HRTFs adjusted for the bone-conduction pathway~\cite{stanley2006lateralization}.
+Cheaper headphones and bone-conducting headphones generally report unfavourable results when compared to over-ear or more expensive alternatives~\cite{schonstein2008comparison,macdonald2006spatial,stanley2006lateralization}.
+This seems to be limited to the elevation dimension, however, and can be improved with HRTFs adjusted for the bone-conducting pathway~\cite{stanley2006lateralization}.
 In~\cite{durette2008visuo}, they transmit a target's elevation angle by adjusting the signal's pitch and report favourable results. 
@@
 
-A set of bone-conduction headphones is used as the audio transmission medium.
+A set of bone-conducting headphones is used as the audio transmission medium.
 These headphones sit on a user's cheekbones and conduct the audio signals through the skull into the inner ear, instead of through the outer part like typical over-ear headphones. 
@@
 
-A generic HRTF implementation with bone-conduction headphones is not very effective in conveying the elevation angle of a sound source~\cite{macdonald2006spatial,schonstein2008comparison}.
+A generic HRTF implementation with bone-conducting headphones is not very effective in conveying the elevation angle of a sound source~\cite{macdonald2006spatial,schonstein2008comparison}.
 To compensate for this, we communicate the target's elevation angle by adjusting the tone's pitch (i.e.\ the sine wave's frequency) as a function of the elevation. 
@@
 
-\begin{figure}[t]
-  \centering
-  \includegraphics[clip=true, trim=0 120 80 50, width=0.9\columnwidth]{figures/pipeline.pdf}
-  \caption{A diagram of the individual system components and their communication pipelines. $F$ indicates a feedback signal and $P$ a pose signal. }\label{fig:pipeline}
-\end{figure}
-
 \subsection{Implementation}
@@
 A WiFi recording module is constantly monitoring the different values of the device's parameters and of the target's position, as well as the system's output, recording everything to a remotely stored datafile.
+
+\begin{figure}[t]
+  \centering
+  \includegraphics[clip=true, trim=0 120 80 50, width=0.9\columnwidth]{figures/pipeline.pdf}
+  \caption{A diagram of the individual system components and their communication pipelines. $F$ indicates a feedback signal and $P$ a pose signal. }\label{fig:pipeline}
+\end{figure}
 
@@
 
-The error data was used to generate scatter plots that show the correlation between the target's direction presented by the guidance system to the participants and the one they thought the targets were.
-These plots are given in \cref{fig:correlation-results} and show a reasonably positive correlation between the guidance and the selected targets, indicating that the participants correctly interpreted the interface's output in most of the cases.
-The Spearman correlation scores for each dataset, listed in \cref{tab:results}, confirm this, as well as the negative angle bias in the elevation dimensions.
-
-\begin{figure}[t]
+\begin{figure}
   \centering
@@
 
+The error data was used to generate scatter plots that show the correlation between the target's direction presented by the guidance system to the participants and the one they thought the targets were.
+These plots are given in \cref{fig:correlation-results} and show a reasonably positive correlation between the guidance and the selected targets, indicating that the participants correctly interpreted the interface's output in most of the cases.
+The Spearman correlation scores for each dataset, listed in \cref{tab:results}, confirm this, as well as the negative angle bias in the elevation dimensions.
+
 \section{Conclusion}\label{sec:conclusion}
@@
 In this work, we proposed a new audio interface that conveys a target's pan and elevation angles through a spatialised audio tone.
-The tone's pitch is adjusted to convey the elevation and compensate for the shortcomings of bone-conduction headsets in transmitting the elevation component of a spatialised signal.
-The results obtained from a simple pointing experiment are encouraging as they show improved performance over purely spatialised signals transmitted via bone-conduction headphones.
+The tone's pitch is adjusted to convey the elevation and compensate for the shortcomings of bone-conducting headsets in transmitting the elevation component of a spatialised signal.
+The results obtained from a simple pointing experiment are encouraging as they show improved performance over purely spatialised signals transmitted via bone-conducting headphones.
 Furthermore, the limitations of the latter were largely overcome by communicating the target's elevation with the tone's pitch, achieving pan and elevation errors comparable to that of normal over-ear headphones.
@@
 
+\paragraph{{\bf Acknowledgements.}}\label{sec:acknowledge}
+This research is partly supported by a Google Faculty Research Award.
+We would like to thank the Voluntary Centre Services UK for their help in facilitating the experiments with people with limited vision.
+
 \bibliographystyle{splncs04}

### 2019-09-11 paper.tex
moved ActiVis footnote ref

@@
 \begin{abstract}
-  The ActiVis\footnote{\url{https://lcas.github.io/ActiVis/}} project's aim is to build a mobile guidance aid to help people with limited vision find objects in an unknown environment.
+  The ActiVis project's aim is to build a mobile guidance aid to help people with limited vision find objects in an unknown environment.
   This system uses bone-conducting headphones to transmit audio signals to the user and requires an effective non-visual interface.
@@
 
-The ActiVis project aims to deliver a mobile guidance system that will ultimately be able to guide a user with vision impairments on the last leg of their journey, i.e.\ the so-called `last 10-yard problem'.
+The ActiVis\footnote{\url{https://lcas.github.io/ActiVis/}} project aims to deliver a mobile guidance system that will ultimately be able to guide a user with vision impairments on the last leg of their journey, i.e.\ the so-called `last 10-yard problem'.
 The prototype platform is based on a Google Project Tango\footnote{\url{https://en.wikipedia.org/wiki/Tango\_(platform)}} device, pictured in \cref{fig:tango} and \cref{fig:tango-headphone}, that embeds a colour camera and provides access to powerful real-time localisation (through IMU measurements and landmark tracking) and image-processing facilities.

### 2019-09-12 paper.tex
Reverted to conduction

--- 2019-09-11 paper.tex
@@
   The ActiVis\footnote{\url{https://lcas.github.io/ActiVis/}} project's aim is to build a mobile guidance aid to help people with limited vision find objects in an unknown environment.
-  This system uses bone-conducting headphones to transmit audio signals to the user and requires an effective non-visual interface.
+  This system uses bone-conduction headphones to transmit audio signals to the user and requires an effective non-visual interface.
   To this end, we propose a new audio-based interface that uses a spatialised signal to convey a target's position on the horizontal plane. 
-  The vertical position on the median plan is given by adjusting the tone's pitch to overcome the audio localisation limitations of bone-conducting headphones. 
+  The vertical position on the median plan is given by adjusting the tone's pitch to overcome the audio localisation limitations of bone-conduction headphones. 
   This interface is validated through a set of experiments with blindfolded and visually impaired participants. 
@@
 It also provides access to Android's full range of interface tools and I/O options. 
-Furthermore, a set of bone-conducting headphones (pictured in \cref{fig:tango-headphone}), which are placed on a user's cheekbones instead of their ears and do not interfere with normal hearing, are used to transmit the audio signals to the user.
+Furthermore, a set of bone-conduction headphones (pictured in \cref{fig:tango-headphone}), which are placed on a user's cheekbones instead of their ears and do not interfere with normal hearing, are used to transmit the audio signals to the user.
 
@@
   \subfloat[]{\label{fig:participant}\includegraphics[width=0.2\columnwidth]{figures/vi_participant.pdf}}
-  \caption{Pictures of the latest Tango mobile phone (left), the former Tango tablet and the bone-conducting headphones used in this work (centre), and a participant with visual impairments during an experiment (right).}
+  \caption{Pictures of the latest Tango mobile phone (left), the former Tango tablet and the bone-conduction headphones used in this work (centre), and a participant with visual impairments during an experiment (right).}
 \end{figure}
@@
 In our case, only the pan and elevation positions are transmitted to the user to point the camera towards a target object or visual feature.
-However, since bone-conducting headphones bypass the outer ear structure, their spectral signature cannot properly be interpreted and we therefore convey the target's elevation angle by adjusting the tone's pitch.
+However, since bone-conduction headphones bypass the outer ear structure, their spectral signature cannot properly be interpreted and we therefore convey the target's elevation angle by adjusting the tone's pitch.
 A similar approach was used in~\cite{durette2008visuo}, but the study did not focus on or investigate the interface's efficacy.
-Indeed, the work presented in this paper serves as an initial study into an interface that can effectively communicate the pan and elevation angles of a target location through a set of bone-conducting headphones using standard audio libraries and audio manipulations.
+Indeed, the work presented in this paper serves as an initial study into an interface that can effectively communicate the pan and elevation angles of a target location through a set of bone-conduction headphones using standard audio libraries and audio manipulations.
 Furthermore, we demonstrate this approach's effectiveness through a simple experiment.
@@
 However, other authors have found that the choice of audio transmission medium can have a significant effect on performance.
-Cheaper headphones and bone-conducting headphones generally report unfavourable results when compared to over-ear or more expensive alternatives~\cite{schonstein2008comparison,macdonald2006spatial,stanley2006lateralization}.
-This seems to be limited to the elevation dimension, however, and can be improved with HRTFs adjusted for the bone-conducting pathway~\cite{stanley2006lateralization}.
+Cheaper headphones and bone-conduction headphones generally report unfavourable results when compared to over-ear or more expensive alternatives~\cite{schonstein2008comparison,macdonald2006spatial,stanley2006lateralization}.
+This seems to be limited to the elevation dimension, however, and can be improved with HRTFs adjusted for the bone-conduction pathway~\cite{stanley2006lateralization}.
 In~\cite{durette2008visuo}, they transmit a target's elevation angle by adjusting the signal's pitch and report favourable results. 
@@
 
-A set of bone-conducting headphones is used as the audio transmission medium.
+A set of bone-conduction headphones is used as the audio transmission medium.
 These headphones sit on a user's cheekbones and conduct the audio signals through the skull into the inner ear, instead of through the outer part like typical over-ear headphones. 
@@
 However, given the previously discussed limitations of bone-conduction and spatialised audio, we propose a simple linear adjustment to the signal's pitch as a function of the elevation angle. 
-The pan angle, instead, can be conveyed by transforming the audio signal with an HRTF, and indeed it has been found that this dimension is unaffected by using bone-conducting headphones\cite{schonstein2008comparison,macdonald2006spatial,stanley2006lateralization}.
+The pan angle, instead, can be conveyed by transforming the audio signal with an HRTF, and indeed it has been found that this dimension is unaffected by using bone-conduction headphones\cite{schonstein2008comparison,macdonald2006spatial,stanley2006lateralization}.
 
@@
 
-A generic HRTF implementation with bone-conducting headphones is not very effective in conveying the elevation angle of a sound source~\cite{macdonald2006spatial,schonstein2008comparison}.
+A generic HRTF implementation with bone-conduction headphones is not very effective in conveying the elevation angle of a sound source~\cite{macdonald2006spatial,schonstein2008comparison}.
 To compensate for this, we communicate the target's elevation angle by adjusting the tone's pitch (i.e.\ the sine wave's frequency) as a function of the elevation. 
@@
 In this work, we proposed a new audio interface that conveys a target's pan and elevation angles through a spatialised audio tone.
-The tone's pitch is adjusted to convey the elevation and compensate for the shortcomings of bone-conducting headsets in transmitting the elevation component of a spatialised signal.
-The results obtained from a simple pointing experiment are encouraging as they show improved performance over purely spatialised signals transmitted via bone-conducting headphones.
+The tone's pitch is adjusted to convey the elevation and compensate for the shortcomings of bone-conduction headsets in transmitting the elevation component of a spatialised signal.
+The results obtained from a simple pointing experiment are encouraging as they show improved performance over purely spatialised signals transmitted via bone-conduction headphones.
 Furthermore, the limitations of the latter were largely overcome by communicating the target's elevation with the tone's pitch, achieving pan and elevation errors comparable to that of normal over-ear headphones.

### 2019-09-24 paper.tex
Made minor changes to coform to editor's requests

@@
 \usepackage{booktabs}
-\usepackage[ruled]{algorithm2e} % For algorithms
+\usepackage[ruled]{algorithm2e}
 
@@
 \author{Jacobus C. Lock\inst{1} \and 
-Iain Gilchrist\inst{2} \and
+Iain D. Gilchrist\inst{2} \and
 Grzegorz Cielniak\inst{1} \and
@@
 
-\institute{University of Lincoln \email{\{jlock, gcielniak, nbellotto\}@lincoln.ac.uk} \and
-	   University of Bristol \email{i.d.gilchrist@bristol.ac.uk}}
+\institute{University of Lincoln, Lincoln, LN6 7TS, UK \email{\{jlock, gcielniak, nbellotto\}@lincoln.ac.uk} \and
+	   University of Bristol, Bristol, BS8 1TH, UK \email{i.d.gilchrist@bristol.ac.uk}}
 
 \authorrunning{Jacobus C. Lock et.\ al.}
+\titlerunning{Bone-conduction Interface For People with Visual Impairments}
 
@@
 \begin{abstract}
-  The ActiVis\footnote{\url{https://lcas.github.io/ActiVis/}} project's aim is to build a mobile guidance aid to help people with limited vision find objects in an unknown environment.
+  The ActiVis project's aim is to build a mobile guidance aid to help people with limited vision find objects in an unknown environment.
   This system uses bone-conduction headphones to transmit audio signals to the user and requires an effective non-visual interface.
@@
 Monaural cues are taken from the interaction of the sound with the human anatomy, e.g.\ head, shoulders, outer ear, before it enters the ear canal.
-%The user's anatomy acts as a filter and the direction the sound wave hits the body activates different filter resonances, attenuating or accentuating different frequency bands.
 When the modified audio signal enters the inner ear canal, the human brain is able to analyse the frequency response and accurately determine the position of the sound source on the median plane. 
@@
 As with the pan estimation errors, the participants from \textit{G1} seemed more adept at estimating the correct elevation.
-%However, the difference is not as significant as for the pan, with \textit{G1} displaying a 36\% improvement over \textit{G2} in their elevation estimation vs. 42.5\% in the pan dimension. 
 Comparing these results to methods that convey elevation through an HRTF, we see an improvement of approximately 63\% and 47\% for each group~\cite{schonstein2008comparison}. 
